### 2024-08-01 weeks/week1/week1a.Rmd
updated week 1 - intro

---
title:  |
    | Environmental and Development Economics
    | Week 1 - Introduction
author:
  |
    | Raahil Madhok
    | UMN Applied Economics
date: "`r Sys.Date()`"

# Output type and options (no TOC and yes fig captions)
output: 
  beamer_presentation:
    theme: Montpellier
classoption: "aspectratio=169"

# This includes latex arguments
header-includes:
  - \AtBeginDocument{\title[Week 1 - Introduction]{\textbf{Environmental and Development Economics} \\ Week 1 - Introduction}}
  - \input{header.tex}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "png") # NOTE: switched to png instead of pdf to decrease size of the resulting pdf
```

# Introduction

## Introduce yourself

- First, lets do introductions
\vfill

- Name, year, memorable summer activity, **research interests**
\vfill

- Why are you taking this class?

## Today

\vfill
- Course overview
\vfill
- Detailed outline
\vfill
- Grade breakdown
\vfill
- Course website on github: https://github.com/rmadhok/enviro-dev-grad
  - lecture slides, assignments, syllabus

\vfill

##

\centering
\huge Course Overview

## Course Overview: There is no textbook

Instead, I am organizing around **\textcolor{orange}{FOUR}** key questions:

1. Why is environmental quality so bad in developing countries?

2. What are the costs of poor environmental quality in developing countries?

3. Why is WTP for environmental quality low in developing countries?

4. What are the political economy barriers to environmental protection?

## Course Approach

I will:

- Frame (almost) each topic with some theory

- Teach applied papers
  - research design, identification strategy, estimation techniques

- Emphasize recent JMPs

\vfill

I will NOT:

- Teach econometrics

- Teach coding

- Teach every topic in environment/development

## Course Goals

For students specializing in environment/development:

1. Show you research frontier

2. Inspire your thesis/JMP ideas

\vfill

For other students:

3. Advance your training as applied microeconomists

4. Show you what makes a top-tier research question

## Course Structure

- This is a brand new class, so I give myself leeway to make changes
\vfill
- You have the unique opportunity to determine direction of the course
  - Think about what topics do and don't interest you
  - And let me know!
\vfill
- Please check the course website regularly for updates

##

\centering
\huge Course Outline and Topics

## Weeks 1 and 2

**Week 1: Theory of Environment/Development**

- How to use theory to ask the right questions (lecture 2)

\vfill

**Week 2: The effect of development on the environment**

- air quality, water quality (lecture 3)
- forests, biodiversity (lecture 4)

## Weeks 3 and 4

**Week 3: The effect of the environment on development**

- human capital (lecture 5)
- firms (lecture 6)

\vfill

**Week 4: WTP for Environmental Quality**

- TBD (lecture 7)
- TBD (lecture 8)

## Weeks 5, 6, and 7

**Week 5: Environmental policy design**

- Spillovers (lecture 9)
- Pollution markets in developing countries (lecture 10)

\vfill

**Week 6: Political economy of the environment**

- Corruption and weak enforcement  (lecture 11)
- Environmental justice (lecture 12)

\vfill

**Week 7: Research presentations**

- I will provide more details throughout the semester

##

\centering
\huge Grade Breakdown

## Broad Breakdown

\Large

\begin{table}[]
\begin{tabular}{lllll}
In-class presentations      & 10\% &  &  &  \\
Problem Set           & 20\% &  &  &  \\
Research Proposal & 60 \% &  &  &  \\
Participation                     & 10\%     &  &  & 
\end{tabular}
\end{table}

## In-class presentations (10%)

- At start of each lecture (from week 2), you will present a paper for 15 minutes

- Each student reads paper and submits **seven** slides (5% of grade)
  - motivation, research question, methods, results
  - 10 mins presentation + 5 mins Q&A (5% of grade)

- I will select presenter **randomly** with replacement**

\vfill

** \footnotesize If you are never chosen, your grade is based on slides.

## Problem Set (20%)

- You will replicate an environment/development paper
  - You will also **extend** the results

- You will become familiar with coding in publication-quality papers

- You will use R or Stata

## Research Proposal (60%)

\begin{table}[]
\begin{tabular}{lllll}
Written Proposal      & 30\% &  &  &  \\
Peer Review           & 20\% &  &  &  \\
Proposal Presentation & 10\% &  &  &  \\
                      &      &  &  & 
\end{tabular}
\end{table}

- You will develop a research proposal for an original idea
  - You are NOT expected to actually do the analysis  
  - I will provide small deadlines (outline, first draft, etc.) along the way

- You will peer review each others proposals

- You will present the proposal at the end of the semester (30 mins)

## Participation (10%)

- I take this seriously

- Not enough to just show up to class

- Quality of questions/discussion count

##

\centering
\Huge Questions?

## Next class

\vfill

- Introduce conceptual framework of environment/development economics

### 2024-08-01 weeks/week1/script.R

@@ -1,5 +1,0 @@
-# Week 1 - Introduction to R
-library(tidyverse)
-
-# load data
-data <- read_csv("data.csv")

### 2024-08-02 lectures/week1/week1a.Rmd
update slides

--- 2024-08-01 weeks/week1/week1a.Rmd
@@
 \vfill
-- Course website on github: https://github.com/rmadhok/enviro-dev-grad
-  - lecture slides, assignments, syllabus
+- Course website: https://github.com/rmadhok/enviro-dev-grad
+  - lectures, assignments, syllabus
 
@@
 
-For students specializing in environment/development:
-
-1. Show you research frontier
+1. Show you environment/development research frontier
 
 2. Inspire your thesis/JMP ideas
-
-\vfill
-
-For other students:
 
@@
 
-## Broad Breakdown
+## Breakdown
 
@@
 
-- At start of each lecture (from week 2), you will present a paper for 15 minutes
+- At start of **first** lecture each week, you'll give a 15 min paper presentation
 
-- Each student reads paper and submits **seven** slides (5% of grade)
+- Each student submits **seven** summary slides (5% of grade)
   - motivation, research question, methods, results
@@
 
-- I will select presenter **randomly** with replacement**
+- I will select presenter on-the-spot 
+  - **randomly** with replacement**
 

### 2024-08-29 assignments/research_proposal/response_to_reviewers.Rmd
add lectures 1-5

---
title: |
        | Assignment: Response to Reviewers
subtitle: |
        | Ph.D. Environmental and Development Economics
        | UMN Applied Economics 2024
        | Instructor: Raahil Madhok
# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# This includes latex arguments
header-includes:
  - \input{header.tex}

---

**Due Date: October 10th, 2024**

# Overview
In this assignment, you will write a mock ``response to the reviewer'' based on the peer review you received for your research proposal. This process mimics the response letter you would write in a journal submission, where you address each comment made by the referee, explain how you revised the paper, and justify any decisions you made in response to the feedback.

The purpose of this exercise is to help you practice responding to critical feedback in a constructive and professional manner, while also improving your proposal based on your peer's suggestions.

# Structure of the Response

Your response should include the following components:

## Introduction Paragraph

- **Thank the reviewer**: Begin by thanking your reviewer for their thoughtful feedback
- **General summary**: Summarize the key points of the review and provide a brief overview of the changes you have made to your proposal as a result of the feedback.
- **Tone**: Keep your tone respectful, professional, and appreciative.

## Point-by-Point Response
For each comment made by the reviewer, provide a clear and detailed response. The unwritten rule is to copy and paste the peer review into a document, turn each comment into a bulletted list, and then respond underneath each comment in the list.

For each point:

- **Restate the reviewer’s comment**: Begin by restating the comment or suggestion. This ensures clarity and demonstrates that you fully understand the critique.

- **Explain your response**:
    - **If you have made changes**: Explain what specific changes you made to the proposal in response to the reviewer’s comment. Clearly identify where in the revised proposal these changes can be found (e.g., “We have revised Section 2, paragraphs 3 and 4, to clarify the theoretical framework and the role of environmental externalities.”).
    - **If you disagree with the comment**: Provide a reasoned explanation for why you have chosen not to make the suggested change. Be polite and respectful, offering evidence or justification for maintaining your original approach. For example, “While we appreciate this suggestion, we believe that keeping the focus on District-level data is more consistent with the existing literature.”

- **Detail the improvement**: Explain how the changes you made improve the overall quality and clarity of your proposal. This will help the reviewer see the value of the revisions.

## Conclusion Paragraph

- **Reaffirm gratitude**: Reiterate your appreciation for the reviewer’s input.
- **Summary of revisions**: Offer a final summary of the revisions made and express your confidence that the proposal has been strengthened as a result.

# Guidelines for Writing Your Response

- **Be thorough**: Ensure that every comment in the review, whether major or minor, receives a response. Do not skip any points.

- **Be specific**: When describing the changes you made, be as specific as possible. Reference the exact sections or pages of your proposal where you incorporated changes.

- **Be polite and constructive**: Even if you disagree with some of the reviewer’s suggestions, maintain a respectful tone. Use evidence and logical reasoning to support your decisions.

- **Reflect improvements**: Make sure the changes you discuss in your response are actually incorporated into the revised version of your proposal. Your response and your revised proposal should be fully aligned.

# Format

- **Length:** Responses (not including the original comment) should not exceed 4 pages (1000 words)

- **Format:** 12 point font, double spaced, 1 inch margins

- **Submission:** PDF document (latex or word)

# Rubric
\begin{table}[h!]
\centering
\begin{tabular}{|l|p{8cm}|c|c|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Score} \\
\hline
\multirow{2}{*}{\textbf{Comprehensive Response}} & Addresses all major and minor points raised by the reviewer, with no comments left unaddressed. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Clarity and Specificity}} & Clearly provides a detailed, specific response to each comment. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Incorporation of Changes}} & Demonstrates thoughtful incorporation of reviewer suggestions into the proposal, with clear references to specific changes made. & 30 & \\
\hline
\multirow{2}{*}{\textbf{Justification for Disagreements}} & Provides clear and well-reasoned explanations for why certain suggestions were not incorporated, using evidence and logical reasoning. & 30 & \\
\hline
\multirow{2}{*}{\textbf{Tone and Professionalism}} & Maintains a respectful, professional, and constructive tone throughout the response, even when disagreeing. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Writing Quality}} & The response document is well-organized, with clear writing, correct grammar, and appropriate length. & 10 & \\
\hline
\multicolumn{3}{|r|}{\textbf{Total Points Possible:}} & \textbf{100} \\
\hline
\end{tabular}
\end{table}

### 2024-08-29 assignments/research_proposal/response_to_reviewers.tex
add lectures 1-5

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\input{header.tex}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Assignment: Response to Reviewers}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ph.D.~Environmental and Development Economics\\
UMN Applied Economics 2024\\
Instructor: Raahil Madhok}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\textbf{Due Date: October 10th, 2024}

\hypertarget{overview}{%
\section{Overview}\label{overview}}

In this assignment, you will write a mock ``response to the reviewer'\,' based on the peer review you received for your research proposal. This process mimics the response letter you would write in a journal submission, where you address each comment made by the referee, explain how you revised the paper, and justify any decisions you made in response to the feedback.

The purpose of this exercise is to help you practice responding to critical feedback in a constructive and professional manner, while also improving your proposal based on your peer's suggestions.

\hypertarget{structure-of-the-response}{%
\section{Structure of the Response}\label{structure-of-the-response}}

Your response should include the following components:

\hypertarget{introduction-paragraph}{%
\subsection{Introduction Paragraph}\label{introduction-paragraph}}

\begin{itemize}
\tightlist
\item
  \textbf{Thank the reviewer}: Begin by thanking your reviewer for their thoughtful feedback
\item
  \textbf{General summary}: Summarize the key points of the review and provide a brief overview of the changes you have made to your proposal as a result of the feedback.
\item
  \textbf{Tone}: Keep your tone respectful, professional, and appreciative.
\end{itemize}

\hypertarget{point-by-point-response}{%
\subsection{Point-by-Point Response}\label{point-by-point-response}}

For each comment made by the reviewer, provide a clear and detailed response. The unwritten rule is to copy and paste the peer review into a document, turn each comment into a bulletted list, and then respond underneath each comment in the list.

For each point:

\begin{itemize}
\item
  \textbf{Restate the reviewer's comment}: Begin by restating the comment or suggestion. This ensures clarity and demonstrates that you fully understand the critique.
\item
  \textbf{Explain your response}:

  \begin{itemize}
  \tightlist
  \item
    \textbf{If you have made changes}: Explain what specific changes you made to the proposal in response to the reviewer's comment. Clearly identify where in the revised proposal these changes can be found (e.g., ``We have revised Section 2, paragraphs 3 and 4, to clarify the theoretical framework and the role of environmental externalities.'').
  \item
    \textbf{If you disagree with the comment}: Provide a reasoned explanation for why you have chosen not to make the suggested change. Be polite and respectful, offering evidence or justification for maintaining your original approach. For example, ``While we appreciate this suggestion, we believe that keeping the focus on District-level data is more consistent with the existing literature.''
  \end{itemize}
\item
  \textbf{Detail the improvement}: Explain how the changes you made improve the overall quality and clarity of your proposal. This will help the reviewer see the value of the revisions.
\end{itemize}

\hypertarget{conclusion-paragraph}{%
\subsection{Conclusion Paragraph}\label{conclusion-paragraph}}

\begin{itemize}
\tightlist
\item
  \textbf{Reaffirm gratitude}: Reiterate your appreciation for the reviewer's input.
\item
  \textbf{Summary of revisions}: Offer a final summary of the revisions made and express your confidence that the proposal has been strengthened as a result.
\end{itemize}

\hypertarget{guidelines-for-writing-your-response}{%
\section{Guidelines for Writing Your Response}\label{guidelines-for-writing-your-response}}

\begin{itemize}
\item
  \textbf{Be thorough}: Ensure that every comment in the review, whether major or minor, receives a response. Do not skip any points.
\item
  \textbf{Be specific}: When describing the changes you made, be as specific as possible. Reference the exact sections or pages of your proposal where you incorporated changes.
\item
  \textbf{Be polite and constructive}: Even if you disagree with some of the reviewer's suggestions, maintain a respectful tone. Use evidence and logical reasoning to support your decisions.
\item
  \textbf{Reflect improvements}: Make sure the changes you discuss in your response are actually incorporated into the revised version of your proposal. Your response and your revised proposal should be fully aligned.
\end{itemize}

\hypertarget{format}{%
\section{Format}\label{format}}

\begin{itemize}
\item
  \textbf{Length:} Responses (not including the original comment) should not exceed 4 pages (1000 words)
\item
  \textbf{Format:} 12 point font, double spaced, 1 inch margins
\item
  \textbf{Submission:} PDF document (latex or word)
\end{itemize}

\hypertarget{rubric}{%
\section{Rubric}\label{rubric}}

\begin{table}[h!]
\centering
\begin{tabular}{|l|p{8cm}|c|c|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Score} \\
\hline
\multirow{2}{*}{\textbf{Comprehensive Response}} & Addresses all major and minor points raised by the reviewer, with no comments left unaddressed. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Clarity and Specificity}} & Clearly provides a detailed, specific response to each comment. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Incorporation of Changes}} & Demonstrates thoughtful incorporation of reviewer suggestions into the proposal, with clear references to specific changes made. & 30 & \\
\hline
\multirow{2}{*}{\textbf{Justification for Disagreements}} & Provides clear and well-reasoned explanations for why certain suggestions were not incorporated, using evidence and logical reasoning. & 30 & \\
\hline
\multirow{2}{*}{\textbf{Tone and Professionalism}} & Maintains a respectful, professional, and constructive tone throughout the response, even when disagreeing. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Writing Quality}} & The response document is well-organized, with clear writing, correct grammar, and appropriate length. & 10 & \\
\hline
\multicolumn{3}{|r|}{\textbf{Total Points Possible:}} & \textbf{100} \\
\hline
\end{tabular}
\end{table}

\end{document}

### 2024-08-29 assignments/paper_presentations/presentation_tips.Rmd
add lectures 1-5

---
title: |
        | Assignment: Paper Presentations
subtitle: |
        | Ph.D. Environment and Development Economics
        | UMN Department of Applied Economics
        | Instructor: Raahil Madhok
date: "`r Sys.Date()`"
# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# This includes latex arguments
header-includes:
  - \input{header.tex}

---

# Overview
The goal of this assignment is to prepare you for presenting at economics conferences. As you proceed through your PhD, and then into your careers, you will inevitably attend and present at such conferences. This first few times can be nerve-wracking.

Through this assignment, you will development the ability to present a full length research paper in a clearly, concisely, and confidently. We will focus not just on content, but also public speaking skills, slide preparation, content organization, and delivery.

# Structure
One randomly-selected student will present a pre-assigned paper at the beginning of each lecture. All students will read **the same** paper, which can be found below and in the syllabus (highlighted in red). The selection procedure is as follows:

1. **Slides:** Before each lecture, you will each read the assigned paper and prepare **10 slides** summarizing the paper. These slides will be uploaded to Canvas. All students are required to submit summary slides each week regardless of whether you are chosen to present.

2. **The hat:** At the beginning of each lecture, I will put all your names in a hat

3. **Randomization:** I will select a student to present randomly **with replacement**. This means that your name goes back into the hat each week. Thus, you may present in more than once, **or not at all**.

4. **Presentation:** You will give a 10 minute "conference-style" presentation of the paper, with 5 minutes for Q&A. Presentation tips can be found below.

5. **Grades:** If you are never chosen to present, your grade will be based upon your slide submissions.

# Presentation Tips
Many of these tips were provided by \href{https://pages.uoregon.edu/cameron/vita/}{Trudy Cameron} to help students prepare for AERE sessions. \href{https://www.raahilmadhok.com/pdf/fra_slides.pdf}{Here} is an example of a recent presentation I gave (15 mins/15 slides) that follows these tips.

- Each presentation should have 10 slides (fewer is okay, but not more)
- Allocate about 1 minute per slide
- It helps to rehearse a couple of times to ensure it fits into the allotted time. It is considered poor form at conferences if you exceed your time.
- Do not present every detail. Your job is to summarize and entice people to read the full paper.
- Start your presentation with a single slide that succinctly bullets your: research question, setting, data, empirical strategy, and two key results. This previews the whole paper, and rest of the presentation is dedicating to elaborating each bullet. 
- Type and use one-line bullet points. Eye-tracking to a new line can interfere with quickly grasping a point. Don't leave just a single word on a wrapped line of a bullet point). Remove superfluous words.
- Since you only have 10 minutes, don't spend too much time on the literature review. Instead, state how your paper improves on the literature (i.e. contributions), but still include citations.
- Aim to show only one equation. Try to to talk ``about'' your equation rather than provide a detailed explanation of every term. You want to bring along the median audience member, not just dazzle the most sophisticated theorist/econometrician in the room.
- Try to not read your slides directly. The audience can read. The bullet points should guide what you say, which means rehearsal is important. Think carefully about how you will transition between slides.

# Rubric

\begin{tabular}{| p{2.1cm} | p{3.2cm} | p{3.2cm} | p{3.2cm} | p{3.2cm} | c |}
    \hline
    \textbf{Criteria} & \textbf{Exemplary (9-10 points)} & \textbf{Proficient (7-8 points)} & \textbf{Adequate (5-6 points)} & \textbf{Needs Improvement (1-4 points)} & \textbf{Score} \\
    \hline
    \textbf{1. Content} & Presentation is well-organized, provides a clear and concise summary of research, includes key points (question, setting, data, methods, results). & Presentation covers most aspects of the research but lacks some detail or clarity. & Presentation includes some key aspects but may be incomplete or unclear. & Key components of research (question, setting, data, methods, results) are missing or poorly presented. & /10 \\
    \hline
    \textbf{2. Slide Design \& Visuals} & Slides are well-designed, easy to follow, and visually appealing. Appropriate use of visuals enhances understanding. & Slides are clear but may have minor design issues. Visuals are used effectively but may lack impact. & Slides are functional but may be cluttered or unclear in some places. Visual aids are minimal. & Slides are difficult to follow, overcrowded, or unappealing. Minimal or inappropriate use of visuals. & /10 \\
    \hline
    \textbf{3. Time Management} & Presentation stays within the 10-minute limit and allocates appropriate time to each slide. & Presentation slightly exceeds or falls short of time but still covers most key points. & Presentation noticeably over/under time. Some key points are rushed or not fully covered. & Presentation significantly exceeds/under time, leading to incomplete or poorly covered content. & /10 \\
    \hline
    \textbf{4. Clarity \& Communication} & Information is conveyed clearly and confidently. Speaker engages the audience and transitions smoothly between slides. & Communication is clear, but speaker may be less confident or less engaging at times. & Communication is uneven or unclear at times. Speaker may rely too much on reading slides. & Presentation is unclear or disorganized, and speaker has difficulty communicating effectively. & /10 \\
    \hline
    \textbf{5. Depth of Analysis} & Presentation demonstrates a deep understanding of the topic, with insightful analysis of data and results. Contributions to the literature are highlighted effectively. & Presentation demonstrates a solid understanding but may lack depth in analysis or critical insights. & Presentation shows basic understanding but lacks depth in analysis or discussion of contributions. & Presentation lacks understanding of the topic, analysis is superficial or absent, and contributions are unclear. & /10 \\
    \hline
    \textbf{6. Audience Engagement} & Speaker interacts well with the audience, maintains attention, and responds effectively to questions. & Speaker engages the audience but may be less interactive or responsive. & Audience engagement is minimal, and responses to questions may be limited or unclear. & Speaker fails to engage the audience, with little interaction or ineffective responses to questions. & /10 \\
    \hline
    \textbf{7. Professionalism} & Presentation is polished, well-rehearsed, and professional. Speaker appears confident and poised. & Presentation is generally professional, though there may be minor issues with delivery or polish. & Presentation shows some professionalism but may feel unpolished or under-rehearsed. & Presentation lacks professionalism, is disorganized, or appears poorly rehearsed. & /10 \\
    \hline
    \textbf{Total Score} & \multicolumn{5}{c|}{\textbf{/70}} \\
    \hline
\end{tabular}

### 2024-08-29 assignments/paper_presentations/presentation_tips.tex
add lectures 1-5

--- 2024-08-29 assignments/paper_presentations/presentation_tips.Rmd
@@
----
-title: |
-        | Assignment: Paper Presentations
-subtitle: |
-        | Ph.D. Environment and Development Economics
-        | UMN Department of Applied Economics
-        | Instructor: Raahil Madhok
-date: "`r Sys.Date()`"
-# Output type and options (no TOC and yes fig captions)
-output:
-  bookdown::pdf_document2: 
-    toc: false
-    fig_caption: yes
-    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
-    keep_tex: yes
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\input{header.tex}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-# This includes latex arguments
-header-includes:
-  - \input{header.tex}
+\title{Assignment: Paper Presentations}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Ph.D.~Environment and Development Economics\\
+UMN Department of Applied Economics}
+\author{}
+\date{\vspace{-2.5em}2024-08-25}
 
----
+\begin{document}
+\maketitle
 
-# Overview
+\hypertarget{overview}{%
+\section{Overview}\label{overview}}
+
 The goal of this assignment is to prepare you for presenting at economics conferences.
@@
 
-# Structure
+\hypertarget{structure}{%
+\section{Structure}\label{structure}}
+
 One randomly-selected student will present a pre-assigned paper at the beginning of each lecture.
-All students will read **the same** paper, which can be found below and in the syllabus (highlighted in red).
+All students will read \textbf{the same} paper, which can be found below and in the syllabus (highlighted in red).
 The selection procedure is as follows:
 
-1.
-**Slides:** Before each lecture, you will each read the assigned paper and prepare **10 slides** summarizing the paper.
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  \textbf{Slides:} Before each lecture, you will each read the assigned paper and prepare \textbf{10 slides} summarizing the paper.
 These slides will be uploaded to Canvas.
 All students are required to submit summary slides each week regardless of whether you are chosen to present.
+\item
+  \textbf{The hat:} At the beginning of each lecture, I will put all your names in a hat
+\item
+  \textbf{Randomization:} I will select a student to present randomly \textbf{with replacement}.
+This means that your name goes back into the hat each week.
+Thus, you may present in more than once, \textbf{or not at all}.
+\item
+  \textbf{Presentation:} You will give a 10 minute ``conference-style'' presentation of the paper, with 5 minutes for Q\&A. Presentation tips can be found below.
+\item
+  \textbf{Grades:} If you are never chosen to present, your grade will be based upon your slide submissions.
+\end{enumerate}
 
-2. **The hat:** At the beginning of each lecture, I will put all your names in a hat
+\hypertarget{presentation-tips}{%
+\section{Presentation Tips}\label{presentation-tips}}
 
-3.
-**Randomization:** I will select a student to present randomly **with replacement**.
-This means that your name goes back into the hat each week.
-Thus, you may present in more than once, **or not at all**.
-
-4. **Presentation:** You will give a 10 minute "conference-style" presentation of the paper, with 5 minutes for Q&A. Presentation tips can be found below.
-
-5. **Grades:** If you are never chosen to present, your grade will be based upon your slide submissions.
-
-# Presentation Tips
 Many of these tips were provided by \href{https://pages.uoregon.edu/cameron/vita/}{Trudy Cameron} to help students prepare for AERE sessions.
@@
 
-- Each presentation should have 10 slides (fewer is okay, but not more)
-- Allocate about 1 minute per slide
-- It helps to rehearse a couple of times to ensure it fits into the allotted time. It is considered poor form at conferences if you exceed your time.
-- Do not present every detail. Your job is to summarize and entice people to read the full paper.
-- Start your presentation with a single slide that succinctly bullets your: research question, setting, data, empirical strategy, and two key results.
+\begin{itemize}
+\tightlist
+\item
+  Each presentation should have 10 slides (fewer is okay, but not more)
+\item
+  Allocate about 1 minute per slide
+\item
+  It helps to rehearse a couple of times to ensure it fits into the allotted time. It is considered poor form at conferences if you exceed your time.
+\item
+  Do not present every detail. Your job is to summarize and entice people to read the full paper.
+\item
+  Start your presentation with a single slide that succinctly bullets your: research question, setting, data, empirical strategy, and two key results.
 This previews the whole paper, and rest of the presentation is dedicating to elaborating each bullet.
-- Type and use one-line bullet points.
+\item
+  Type and use one-line bullet points.
 Eye-tracking to a new line can interfere with quickly grasping a point.
@@
 Remove superfluous words.
-- Since you only have 10 minutes, don't spend too much time on the literature review. Instead, state how your paper improves on the literature (i.e. contributions), but still include citations.
-- Aim to show only one equation.
-Try to to talk ``about'' your equation rather than provide a detailed explanation of every term.
+\item
+  Since you only have 10 minutes, don't spend too much time on the literature review. Instead, state how your paper improves on the literature (i.e.~contributions), but still include citations.
+\item
+  Aim to show only one equation.
+Try to to talk ``about'\,' your equation rather than provide a detailed explanation of every term.
 You want to bring along the median audience member, not just dazzle the most sophisticated theorist/econometrician in the room.
-- Try to not read your slides directly.
+\item
+  Try to not read your slides directly.
 The audience can read.
@@
 Think carefully about how you will transition between slides.
+\end{itemize}
 
-# Rubric
+\hypertarget{rubric}{%
+\section{Rubric}\label{rubric}}
 
@@
 \end{tabular}
+
+\end{document}

### 2024-08-29 assignments/replication_exercise/replication_exercise.Rmd
add lectures 1-5

---
title: |
        | Assignment: Replication and Extension of Empirical Paper
subtitle: |
        | Ph.D. Environment and Development Economics
        | UMN Department of Applied Economics
        | Instructor: Raahil Madhok

# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# This includes latex arguments
header-includes:
  - \input{header.tex}

---

**Due Date: October 21st, 2024**

# Overview
The goal of this assignment is to conduct a replication and extension of a published empirical paper in the field of environmental and development economics. This exercise is designed to help you engage deeply with research in your field of interest by reproducing key findings and contributing your own insights through an extension of the analysis. The goal is to enhance your understanding of empirical methods, familiarize yourself with the challenges of coding, and encourage critical thinking about existing research.

You will select an empirical paper of your choice, replicate and extend it, and submit a short write-up explaining what you did. You will also submit your code.

# Assignment Components

## Select a paper

- Most papers on the syllabus have replication files. 
- Choose a paper that aligns with your research interests
- You may choose a different peer-reviewed paper in environmental & development economics that has publicly available data 
- Note: you will benefit most from selecting a paper with methods you intend to use in your own research.

## Replication Code

- Replicate **2-3 of the main tables or figures**. These should represent the core empirical findings
- You can use Stata or R
- Write your own code for the analysis. Try to do this yourself before consulting the authors' code. Doing so will help you identify potential sources of divergence between your results and the published findings.
- Extensively comment your code to explain exactly what each part does
- Output all tables to LaTeX. If you do not know how, use ChatGPT to help you.

## Extension

- After replicating the main tables, you will **extend the analysis**
- The extension is up to you, but should be carefully thought out and yield a new insight
  - I reserve the right to determine what is ``carefully thought out''
  - E.g. if you drop the five largest outcome values and re-run the analysis, this **does not count**
  - But, if you plot a histogram, find outliers, and truncate at the 95th percentile, **that counts**

- Some ideas for an extension are:
    - **Enhance the research question**: this could mean studying heterogeneity or 
    - **Robustness checks**: Verify the original results under different assumptions or specifications.
    - **Error identification and correction**: If you identify any mistakes in the original analysis, provide a corrected version with clear explanations.

# Write-up (6 pages double spaced)
After conducting the replication and extension, you will explain what you did in a short write-up. The report should contain the following:

1. **Summary:** A one paragraph summary of the paper's research question, setting, methodology, and main findings
2. **Replication Process**: Describe which tables/figures you opted to replicate, the steps you took to replicate the results, and any issues you encountered. Include a detailed explanation of your data processing and analysis.
3. **Results**: Present your replicated tables/figures and compare them with the original paper’s results. Discuss any differences.
4. **Extension**: Explain the extension you undertook, your motivation for doing so, and the results you obtained. Provide any new tables or figures related to your extension.

## Submission Format

- Submit your write up as a PDF in Canvas
- Submit your code: either link to repository in your write-up or upload through Canvas
- Format:
  - **Length:** Max 6 pages double spaced
  - **Format:** 12 point font, 1 inch margins
  - **Code:** Clear, concise, and commented

## Grading Criteria

You will be evaluated based on:

- **Replication Quality**: Completeness, accuracy, and clarity of your replication.
- **Extension**: Insightfulness and rigor of the extension.
- **Writing**: Clear, well-structured, and concise writing that follows academic conventions.
- **Code**: Organization, clarity, and documentation of the code, including reproducibility of results.

\begin{table}[ht]
\centering
\begin{tabular}{|p{4cm}|p{6cm}|p{1.5cm}|p{4cm}|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Description} \\
\hline
\textbf{Written Report (60 Points)} & \textbf{Replication Process (15 Points)} & & \\
\hline
& Description of Methods and Process & 10 & Clear explanation of replication steps, including data cleaning and methodology. \\
\hline
& Data and Documentation Review & 5 & Thoughtful review of data quality and documentation. \\
\hline
\textbf{Replication Results (15 Points)} & Presentation of Results & 10 & Clear presentation and formatting of replicated tables/figures. \\
\hline
& Comparison with Original & 5 & Detailed comparison of replicated results with original findings. \\
\hline
\textbf{Extension (20 Points)} & Description and Justification & 10 & Clear explanation of extension and rationale for its connection to the original paper. \\
\hline
& Results of the Extension & 10 & Presentation of findings from the extension, including new tables/figures. \\
\hline
\textbf{Writing Quality (10 Points)} & Clarity and Structure & 5 & Report is logically structured and clearly written. \\
\hline
& Grammar and Style & 5 & Proper grammar, spelling, and adherence to scientific writing conventions. \\
\hline
\textbf{Code (40 Points)} & \textbf{Code Quality and Organization (15 Points)} & & \\
\hline
& Clarity and Commenting & 10 & Code is clearly written with sufficient comments. \\
\hline
& Organization & 5 & Code is well-organized with logical structure and consistent formatting. \\
\hline
\textbf{Correctness and Functionality (15 Points)} & Accuracy of Replication & 10 & Code accurately replicates the main results. \\
\hline
& Extension Implementation & 5 & Correct implementation of the extension in the code. \\
\hline
\textbf{Reproducibility and Efficiency (10 Points)} & Reproducibility & 5 & Code can be easily reproduced with clear instructions. \\
\hline
& Efficiency & 5 & Code is efficient, avoiding unnecessary complexity or overhead. \\
\hline
& & \textbf{Total} & \textbf{100 Points} \\
\hline
\end{tabular}
\end{table}

### 2024-08-29 assignments/replication_exercise/replication_exercise.tex
add lectures 1-5

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\input{header.tex}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Assignment: Replication and Extension of Empirical Paper}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ph.D.~Environment and Development Economics\\
UMN Department of Applied Economics\\
Instructor: Raahil Madhok}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\textbf{Due Date: October 21st, 2024}

\hypertarget{overview}{%
\section{Overview}\label{overview}}

The goal of this assignment is to conduct a replication and extension of a published empirical paper in the field of environmental and development economics. This exercise is designed to help you engage deeply with research in your field of interest by reproducing key findings and contributing your own insights through an extension of the analysis. The goal is to enhance your understanding of empirical methods, familiarize yourself with the challenges of coding, and encourage critical thinking about existing research.

You will select an empirical paper of your choice, replicate and extend it, and submit a short write-up explaining what you did. You will also submit your code.

\hypertarget{assignment-components}{%
\section{Assignment Components}\label{assignment-components}}

\hypertarget{select-a-paper}{%
\subsection{Select a paper}\label{select-a-paper}}

\begin{itemize}
\tightlist
\item
  Most papers on the syllabus have replication files.
\item
  Choose a paper that aligns with your research interests
\item
  You may choose a different peer-reviewed paper in environmental \& development economics that has publicly available data
\item
  Note: you will benefit most from selecting a paper with methods you intend to use in your own research.
\end{itemize}

\hypertarget{replication-code}{%
\subsection{Replication Code}\label{replication-code}}

\begin{itemize}
\tightlist
\item
  Replicate \textbf{2-3 of the main tables or figures}. These should represent the core empirical findings
\item
  You can use Stata or R
\item
  Write your own code for the analysis. Try to do this yourself before consulting the authors' code. Doing so will help you identify potential sources of divergence between your results and the published findings.
\item
  Extensively comment your code to explain exactly what each part does
\item
  Output all tables to LaTeX. If you do not know how, use ChatGPT to help you.
\end{itemize}

\hypertarget{extension}{%
\subsection{Extension}\label{extension}}

\begin{itemize}
\tightlist
\item
  After replicating the main tables, you will \textbf{extend the analysis}
\item
  The extension is up to you, but should be carefully thought out and yield a new insight

  \begin{itemize}
  \tightlist
  \item
    I reserve the right to determine what is ``carefully thought out'\,'
  \item
    E.g. if you drop the five largest outcome values and re-run the analysis, this \textbf{does not count}
  \item
    But, if you plot a histogram, find outliers, and truncate at the 95th percentile, \textbf{that counts}
  \end{itemize}
\item
  Some ideas for an extension are:

  \begin{itemize}
  \tightlist
  \item
    \textbf{Enhance the research question}: this could mean studying heterogeneity or
  \item
    \textbf{Robustness checks}: Verify the original results under different assumptions or specifications.
  \item
    \textbf{Error identification and correction}: If you identify any mistakes in the original analysis, provide a corrected version with clear explanations.
  \end{itemize}
\end{itemize}

\hypertarget{write-up-6-pages-double-spaced}{%
\section{Write-up (6 pages double spaced)}\label{write-up-6-pages-double-spaced}}

After conducting the replication and extension, you will explain what you did in a short write-up. The report should contain the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Summary:} A one paragraph summary of the paper's research question, setting, methodology, and main findings
\item
  \textbf{Replication Process}: Describe which tables/figures you opted to replicate, the steps you took to replicate the results, and any issues you encountered. Include a detailed explanation of your data processing and analysis.
\item
  \textbf{Results}: Present your replicated tables/figures and compare them with the original paper's results. Discuss any differences.
\item
  \textbf{Extension}: Explain the extension you undertook, your motivation for doing so, and the results you obtained. Provide any new tables or figures related to your extension.
\end{enumerate}

\hypertarget{submission-format}{%
\subsection{Submission Format}\label{submission-format}}

\begin{itemize}
\tightlist
\item
  Submit your write up as a PDF in Canvas
\item
  Submit your code: either link to repository in your write-up or upload through Canvas
\item
  Format:

  \begin{itemize}
  \tightlist
  \item
    \textbf{Length:} Max 6 pages double spaced
  \item
    \textbf{Format:} 12 point font, 1 inch margins
  \item
    \textbf{Code:} Clear, concise, and commented
  \end{itemize}
\end{itemize}

\hypertarget{grading-criteria}{%
\subsection{Grading Criteria}\label{grading-criteria}}

You will be evaluated based on:

\begin{itemize}
\tightlist
\item
  \textbf{Replication Quality}: Completeness, accuracy, and clarity of your replication.
\item
  \textbf{Extension}: Insightfulness and rigor of the extension.
\item
  \textbf{Writing}: Clear, well-structured, and concise writing that follows academic conventions.
\item
  \textbf{Code}: Organization, clarity, and documentation of the code, including reproducibility of results.
\end{itemize}

\begin{table}[ht]
\centering
\begin{tabular}{|p{4cm}|p{6cm}|p{1.5cm}|p{4cm}|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Description} \\
\hline
\textbf{Written Report (60 Points)} & \textbf{Replication Process (15 Points)} & & \\
\hline
& Description of Methods and Process & 10 & Clear explanation of replication steps, including data cleaning and methodology. \\
\hline
& Data and Documentation Review & 5 & Thoughtful review of data quality and documentation. \\
\hline
\textbf{Replication Results (15 Points)} & Presentation of Results & 10 & Clear presentation and formatting of replicated tables/figures. \\
\hline
& Comparison with Original & 5 & Detailed comparison of replicated results with original findings. \\
\hline
\textbf{Extension (20 Points)} & Description and Justification & 10 & Clear explanation of extension and rationale for its connection to the original paper. \\
\hline
& Results of the Extension & 10 & Presentation of findings from the extension, including new tables/figures. \\
\hline
\textbf{Writing Quality (10 Points)} & Clarity and Structure & 5 & Report is logically structured and clearly written. \\
\hline
& Grammar and Style & 5 & Proper grammar, spelling, and adherence to scientific writing conventions. \\
\hline
\textbf{Code (40 Points)} & \textbf{Code Quality and Organization (15 Points)} & & \\
\hline
& Clarity and Commenting & 10 & Code is clearly written with sufficient comments. \\
\hline
& Organization & 5 & Code is well-organized with logical structure and consistent formatting. \\
\hline
\textbf{Correctness and Functionality (15 Points)} & Accuracy of Replication & 10 & Code accurately replicates the main results. \\
\hline
& Extension Implementation & 5 & Correct implementation of the extension in the code. \\
\hline
\textbf{Reproducibility and Efficiency (10 Points)} & Reproducibility & 5 & Code can be easily reproduced with clear instructions. \\
\hline
& Efficiency & 5 & Code is efficient, avoiding unnecessary complexity or overhead. \\
\hline
& & \textbf{Total} & \textbf{100 Points} \\
\hline
\end{tabular}
\end{table}

\end{document}

### 2024-08-29 assignments/research_proposal/peer_review.Rmd
add lectures 1-5

---
title: |
        | Assignment: Referee Report
subtitle: |
        | Ph.D. Environmental and Development Economics
        | UMN Applied Economics 2024
        | Instructor: Raahil Madhok
# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# This includes latex arguments
header-includes:
  - \input{header.tex}

---

**Due Date: October 10th, 2024**

# Overview
In this assignment, you will conduct a peer review of a classmate's research proposal. Peer reviewing is a critical skill in the academic community, where you assess the quality, clarity, and originality of others' work while providing constructive feedback.

On the evening that you submit your first draft (Oct. 3rd), you will receive a classmate's proposal to review. This will be a double-blind process so that neither of you will know who's proposal you are reviewing.

Your peer review should be divided into two sections: major points and minor points. Major points will address significant issues that need more attention in the research question, conceptual framework, or research design. Minor points will address more detailed aspects that can improve the proposal, such as writing style or even suggestions for better data.

Your review should be written in a formal and constructive tone, ensuring your comments are helpful for the writer's improvement.

# Structure of the Peer Review

## Major Points (Minimum 5 items)
The major points should highlight the most critical issues in the proposal. Below are some questions to think about as your identify your major points. You do not necessarily need to come up with criticism under each heading. You may, for example, identify two major criticisms in the theory and none in the research design.

- **The Research Question:** Is the question clear, well-motivated, and significant within the field of environmental and development economics? Does it have a clear connection to both theory and real-world policy relevance?

- **Theoretical Framework:** Is the theoretical framework well-developed and logically consistent? Does it directly relate to the research questions in a meaningful way? Are the assumptions justified? If the author is adapting an existing model, does the adaptation generate new and meaningful insights?

- **Research Design:** Is the empirical strategy clearly stated? Does the design convincingly address the research question? Are all of the identification assumptions addressed? Is there a different research design that may have yielded better identification?

- **Data Sources:** Does the proposal rely on real, available data, and is the data appropriate for the research question? Are there any concerns about data access, quality, or relevance that could hinder the feasibility of the research?

- **Policy Implications:** Does the proposal clearly discuss how the results of the research could influence policy decisions? Are the potential implications grounded in the real-world context?

## Minor Points (Minumum 5 items)
The minor points are smaller issues that, while less critical than the major points, can still enhance the quality and clarity of the proposal. These might include:

- Clarity of the writing (grammar, organization, and structure),
- Minor inconsistencies in the theoretical argument,
- Specific technical clarifications needed in the research design or data analysis,
- Suggestions for additional literature that could strengthen the theoretical framework,
- Formatting issues or citation inaccuracies.
- For each minor point, include:

# Writing Style
When writing your peer review, keep in mind the following guidelines:

- Be constructive. Frame your comments in a way that helps the author improve. For example, instead of saying "The research question is unclear," say, "The research question could be clarified by focusing more on the causal relationship between X and Y."

- Be specific. Provide clear examples and references to specific sections of the proposal. Vague comments like "Improve the theoretical framework" are less helpful than "On page 3, the author should elaborate on how the proposed model incorporates environmental externalities."

- Be respectful. Even if you find serious flaws in the proposal, present your feedback respectfully and professionally. Academic critique should be rigorous but fair.

- Balance criticism with praise. Highlight what the author has done well. This not only encourages them but also makes your review more balanced.

# Format

- **Style:** Each point (major or minor) should be clearly numbered or bulletted.

- **Length:** 4 pages max (1000 words)

- **Format:** 12 point font, double spaced, 1 inch margins

- **Submission:** PDF document (latex or word)

\newpage
# Rubric

\begin{table}[h!]
\centering
\begin{tabular}{|l|p{8cm}|c|c|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Score} \\
\hline
\multirow{2}{*}{\textbf{Understanding of the Proposal}} & Demonstrates a clear understanding of the peer's research question, theoretical framework, and research design. & 10 & \\
\hline
\multirow{3}{*}{\textbf{Major Points}} & Each of the major points identifies key issues in the proposal, explains their importance, and provides constructive suggestions for improvement. & 30 & \\
\cline{2-4}
& Relevance and Depth: The major points address significant elements like research question, theoretical framework, data, design, or policy implications. & 10 & \\
\cline{2-4}
& Explanation of Issues: Major points clearly explain what needs improvement and why these issues matter to the overall proposal. & 10 & \\
\cline{2-4}
& Constructive Suggestions: Provides specific and constructive suggestions to improve the proposal. & 10 & \\
\hline
\multirow{3}{*}{\textbf{Minor Points}} & Each of the minor points identifies smaller issues (clarity, grammar, technical details) and suggests minor improvements. & 20 & \\
\cline{2-4}
& Identification of Minor Issues: The minor points identify relevant and helpful smaller issues. & 10 & \\
\cline{2-4}
& Clarity and Constructiveness: Provides clear, specific feedback on how to address the minor issues. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Tone and Constructiveness}} & The review is respectful, professional, and balanced. The reviewer frames critiques in a positive and constructive manner. & 20 & \\
\hline
\multirow{2}{*}{\textbf{Specificity of Feedback}} & Feedback is specific, well-documented, and refers directly to the text of the proposal. Avoids vague comments. & 10 & \\
\hline
\multirow{2}{*}{\textbf{Writing Quality}} & The peer review is well-written, clear, and organized. Grammar and syntax are correct. & 10 & \\
\hline
\multicolumn{3}{|r|}{\textbf{Total Points Possible:}} & \textbf{150} \\
\hline
\end{tabular}
\end{table}

### 2024-08-29 assignments/research_proposal/peer_review.tex
add lectures 1-5

--- 2024-08-29 assignments/research_proposal/peer_review.Rmd
@@
----
-title: |
-        | Assignment: Referee Report
-subtitle: |
-        | Ph.D. Environmental and Development Economics
-        | UMN Applied Economics 2024
-        | Instructor: Raahil Madhok
-# Output type and options (no TOC and yes fig captions)
-output:
-  bookdown::pdf_document2: 
-    toc: false
-    fig_caption: yes
-    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
-    keep_tex: yes
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\input{header.tex}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-# This includes latex arguments
-header-includes:
-  - \input{header.tex}
+\title{Assignment: Referee Report}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Ph.D.~Environmental and Development Economics\\
+UMN Applied Economics 2024\\
+Instructor: Raahil Madhok}
+\author{}
+\date{\vspace{-2.5em}}
 
----
+\begin{document}
+\maketitle
 
-**Due Date: October 10th, 2024**
+\textbf{Due Date: October 10th, 2024}
 
-# Overview
+\hypertarget{overview}{%
+\section{Overview}\label{overview}}
+
 In this assignment, you will conduct a peer review of a classmate's research proposal.
@@
 
-On the evening that you submit your first draft (Oct.
-3rd), you will receive a classmate's proposal to review.
+On the evening that you submit your first draft (Oct.~3rd), you will receive a classmate's proposal to review.
 This will be a double-blind process so that neither of you will know who's proposal you are reviewing.
@@
 
-# Structure of the Peer Review
+\hypertarget{structure-of-the-peer-review}{%
+\section{Structure of the Peer Review}\label{structure-of-the-peer-review}}
 
-## Major Points (Minimum 5 items)
+\hypertarget{major-points-minimum-5-items}{%
+\subsection{Major Points (Minimum 5 items)}\label{major-points-minimum-5-items}}
+
 The major points should highlight the most critical issues in the proposal.
@@
 
-- **The Research Question:** Is the question clear, well-motivated, and significant within the field of environmental and development economics?
+\begin{itemize}
+\item
+  \textbf{The Research Question:} Is the question clear, well-motivated, and significant within the field of environmental and development economics?
 Does it have a clear connection to both theory and real-world policy relevance?
-
-- **Theoretical Framework:** Is the theoretical framework well-developed and logically consistent?
+\item
+  \textbf{Theoretical Framework:} Is the theoretical framework well-developed and logically consistent?
 Does it directly relate to the research questions in a meaningful way?
@@
 If the author is adapting an existing model, does the adaptation generate new and meaningful insights?
-
-- **Research Design:** Is the empirical strategy clearly stated?
+\item
+  \textbf{Research Design:} Is the empirical strategy clearly stated?
 Does the design convincingly address the research question?
@@
 Is there a different research design that may have yielded better identification?
+\item
+  \textbf{Data Sources:} Does the proposal rely on real, available data, and is the data appropriate for the research question?
+Are there any concerns about data access, quality, or relevance that could hinder the feasibility of the research?
+\item
+  \textbf{Policy Implications:} Does the proposal clearly discuss how the results of the research could influence policy decisions? Are the potential implications grounded in the real-world context?
+\end{itemize}
 
-- **Data Sources:** Does the proposal rely on real, available data, and is the data appropriate for the research question?
-Are there any concerns about data access, quality, or relevance that could hinder the feasibility of the research?
+\hypertarget{minor-points-minumum-5-items}{%
+\subsection{Minor Points (Minumum 5 items)}\label{minor-points-minumum-5-items}}
 
-- **Policy Implications:** Does the proposal clearly discuss how the results of the research could influence policy decisions? Are the potential implications grounded in the real-world context?
-
-## Minor Points (Minumum 5 items)
 The minor points are smaller issues that, while less critical than the major points, can still enhance the quality and clarity of the proposal. These might include:
 
-- Clarity of the writing (grammar, organization, and structure),
-- Minor inconsistencies in the theoretical argument,
-- Specific technical clarifications needed in the research design or data analysis,
-- Suggestions for additional literature that could strengthen the theoretical framework,
-- Formatting issues or citation inaccuracies.
-- For each minor point, include:
+\begin{itemize}
+\tightlist
+\item
+  Clarity of the writing (grammar, organization, and structure),
+\item
+  Minor inconsistencies in the theoretical argument,
+\item
+  Specific technical clarifications needed in the research design or data analysis,
+\item
+  Suggestions for additional literature that could strengthen the theoretical framework,
+\item
+  Formatting issues or citation inaccuracies.
+\item
+  For each minor point, include:
+\end{itemize}
 
-# Writing Style
+\hypertarget{writing-style}{%
+\section{Writing Style}\label{writing-style}}
+
 When writing your peer review, keep in mind the following guidelines:
 
-- Be constructive.
+\begin{itemize}
+\item
+  Be constructive.
 Frame your comments in a way that helps the author improve.
-For example, instead of saying "The research question is unclear," say, "The research question could be clarified by focusing more on the causal relationship between X and Y."
+For example, instead of saying ``The research question is unclear,'' say, ``The research question could be clarified by focusing more on the causal relationship between X and Y.''
+\item
+  Be specific.
+Provide clear examples and references to specific sections of the proposal.
+Vague comments like ``Improve the theoretical framework'' are less helpful than ``On page 3, the author should elaborate on how the proposed model incorporates environmental externalities.''
+\item
+  Be respectful. Even if you find serious flaws in the proposal, present your feedback respectfully and professionally. Academic critique should be rigorous but fair.
+\item
+  Balance criticism with praise. Highlight what the author has done well. This not only encourages them but also makes your review more balanced.
+\end{itemize}
 
-- Be specific.
-Provide clear examples and references to specific sections of the proposal.
-Vague comments like "Improve the theoretical framework" are less helpful than "On page 3, the author should elaborate on how the proposed model incorporates environmental externalities."
+\hypertarget{format}{%
+\section{Format}\label{format}}
 
-- Be respectful. Even if you find serious flaws in the proposal, present your feedback respectfully and professionally. Academic critique should be rigorous but fair.
-
-- Balance criticism with praise. Highlight what the author has done well. This not only encourages them but also makes your review more balanced.
-
-# Format
-
-- **Style:** Each point (major or minor) should be clearly numbered or bulletted.
-
-- **Length:** 4 pages max (1000 words)
-
-- **Format:** 12 point font, double spaced, 1 inch margins
-
-- **Submission:** PDF document (latex or word)
+\begin{itemize}
+\item
+  \textbf{Style:} Each point (major or minor) should be clearly numbered or bulletted.
+\item
+  \textbf{Length:} 4 pages max (1000 words)
+\item
+  \textbf{Format:} 12 point font, double spaced, 1 inch margins
+\item
+  \textbf{Submission:} PDF document (latex or word)
+\end{itemize}
 
 \newpage
-# Rubric
+
+\hypertarget{rubric}{%
+\section{Rubric}\label{rubric}}
 
@@
 \end{table}
+
+\end{document}

### 2024-08-29 assignments/research_proposal/research_proposal.Rmd
add lectures 1-5

---
title: |
        | Research Proposal
subtitle: |
        | Ph.D. Environmental and Development Economics
        | UMN Applied Economics 2024
        | Instructor: Raahil Madhok
# Output type and options (no TOC and yes fig captions)
output:
  bookdown::pdf_document2: 
    toc: false
    fig_caption: yes
    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
    keep_tex: yes

# This includes latex arguments
header-includes:
  - \input{header.tex}

---

# Assignment Overview
The goal of this project is to take you through a mock process of publishing in economics, including writing a paper (proposal, in this case) as well as the revise and resubmit process. You will: 1) draft a research proposal, 2) send your proposal to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, and 4) submit the final draft for ``publication'' (i.e. a final grade...).

You **are not** allowed to ``double dip''. This means the idea must be original and you cannot use your second-year paper, or a proposal from another class.

The objective of this assignment is to help prepare you for writing high-quality economics research papers. Throughout the assignment, we will develop your ability to design a clear and focused research project in the field of environmental and development economics. You are not expected to carry out the analysis, although I hope that you will pursue the project to the end and use it as a dissertation chapter or separate publication.

Broadly, you will identify an original and specific research question, explain why it is important, and outline the research design that you will use to address it. After receiving an ``R\&R'' on the paper, you will improve it and submit a final draft at the end of the semester.

## Timeline
\begin{table}[ht]
\begin{tabular}{|c|c|}
\hline
\textbf{Item} & \textbf{Deadline} \\ \hline
First draft     & October 3rd     \\ \hline
Referee report    & October 10th     \\ \hline
Final draft    & October 31st     \\ \hline
\end{tabular}
\end{table}

## Helpful Resources

- Marc Bellemare's guide on \href{https://marcfbellemare.com/wordpress/wp-content/uploads/2020/09/BellemareHowToPaperSeptember2020.pdf}{how to write applied economics papers}
- Keith Head's \href{https://blogs.ubc.ca/khead/research/research-advice/formula}{introduction formula}
- Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}
- Read papers!

## Format

- **Length:** 15 pages, double spaced (approx 4000 words)

- **Format:** 12 point font, double spaced, 1-inch margins

- **Submission:** PDF document (latex or word)

\newpage
# Proposal Structure

## Title
The title should be concise and informative, reflecting the focus of your research. Please do not use one word titles (e.g. "Forests") or three word titles ("Forests and Development").

## Abstract (250 words)
The abstract should be no more than 10 sentences (about 250 words) and provide a clear and succinct summary of the paper. Describe the central question and why it is important. Then state your methods, research design, and data. Conclude with a statement of policy implications.

## Introduction (3-4 pages)

The introduction is essentially a short summary of the paper. I should be able to read your introduction and know exactly what the proposal contains. In fact, many people only read the introduction and nothing else, so it is important to be clear about what you do in each section of the paper.

- **Contextualize the issue:** Provide a brief background on the topic within environmental and development economics. What are some general descriptive statistics about the issue at hand? Is there a puzzle, or something counterintuitive that your research may shed some light on?

- **Research Question**: Clearly state your research question. It should be specific, well-defined, and narrow enough to be addressed within the scope of the proposal. At the same time, it should ideally speak to a broader issue in environmental and development economics such that it contributes meaningfully to the literature.

- **Importance**: Explain why answering the research question is important. Think about its relevance for policy, environmental conservation, and economic development. Even if it is not directly policy-relevant, state what the intellectual importance is.

- **Summary:** Summarize your proposal for the rest of the introduction. Describe your data, measurement, and why your data is suitable for the research question. Describe your research design, why you chose it, and any identification assumptions you make.

- **Literature Review:** Provide an overview of key literature on your topic. Identify gaps or unresolved questions that your research will address. Avoid providing a detailed summary of each paper, but rather focus on a key takeaway and how it relates to your topic. Also avoid ``punching down'' on the literature; instead, point out how your work will advance our understanding of the topic.

## Conceptual Framework (2 page)
You are required to include a theoretical framework to guide your research design. The goal is to guide my thinking, pin down potential mechanisms, and generate empirically testable hypotheses.

The model represents a framework to fix ideas and help think about the issue at hand. It **DOES NOT** need to be a mathematical model, although I will be happy if you did construct one. If you do write a mathematical model, you are welcome to adapt any of the ones we study in class, or replicate one from the literature (adapted to your context).

Otherwise, your model can be written in words, or even graphs. If your question is about how $x$ affects $y$, think about why you expect a relationship in the first place. Is there an underlying behavioural assumption? Do you expect a positive or negative relationship, or is it ambiguous? Are you interested in recovering a direct effect, or an indirect effect via some other channel?

## Data (1-2 page)
You are expected to choose a research question where **real** data is available to answer it. The data description need not be dry and boring; it can be an opportunity to present an interesting narrative. You can discuss **why** you chose a dataset. Remind the reader about what kind of data would be needed to answer your research question, and any barriers to obtaining that data. You can also discuss what data have been used in the past, and any limitations. If data was a limiting factor preventing your research question from being answered in the past, then say so. Then finally introduce your data and highlight how it enables you to overcome previous barriers:

- **Data sources**: List the data source for each variable. You are welcome to propose scraping the data, as long as the raw data exist (in pdf or online tables) and I can access it.

- **Measurement**: Describe how you will measure each variable using your data. If you are using a household survey, if the variable directly provided? Are you constructing an index? If you are using remote sensing data, briefly describe how the data product captures the variable you want to measure.

- **Unit of analysis**: Desribe your unit of analysis. If your raw data is at a different level (e.g. gridded), then think about how you will aggregate the data.

- **Final Dataset**: Describe your (hypothesized) final dataseta and the panel dimensions.

## Empirical Strategy (4 pages)
In this section you will describe your research design, estimation strategy and your identification strategy.

- **Estimation Strategy:** The estimation strategy comprises the equations you will use to answer your research question. Clearly state which estimator you will use to recover your coefficient(s) of interest. If relevant, describe how your coefficient connects to your theoretical framework.

- **Identification Strategy:** Clearly describe how your research design is set up to identify a causal relationship. State the variation that is exploited and how to interpret your coefficient of interest. Clearly state any assumptions needed when claiming causality. It may help to frame this as a discussion of threats to identification, and how your research design overcomes those threats. If a test is needed to improve confidence in your identification strategy (e.g. a balance test, or test of pre-trends), then write and describe the corresponding equations.

In writing this section, it is important to be honest about what your empirical strategy can and cannot do. Avoid making bold claims about causality that are not fully backed up by your research design.

## Conclusion (1 page)
- Summarize your research question again and why it is important. Try not to re-state it in exactly the same way as in your introduction.

- Discuss any limitations of your research design. Aim to think deeply beyond standard ones such as lack of external validity.

- Discuss real world implications of having answered your research question. Think about which policy-makers would be interested in your results.

- Discuss implications for future research

\clearpage
# Grading Rubric

\begin{table}[h!]
\centering
\begin{tabular}{|l|p{8cm}|c|c|}
\hline
\textbf{Category} & \textbf{Criteria} & \textbf{Points} & \textbf{Score} \\
\hline
\multirow{2}{*}{\textbf{Title and Abstract}} & Title is concise and informative. Abstract clearly summarizes the research question, methods, data, and policy implications in no more than 250 words. & 10 & \\
\hline
\multirow{4}{*}{\textbf{Introduction}} & Provides a clear background and context for the research. The research question is specific, well-defined, and important. Effectively motivates the importance of the research within environmental and development economics. & 20 & \\
\cline{2-4}
& Includes a summary of data, methods, and a brief literature review. & & \\
\hline
\multirow{2}{*}{\textbf{Conceptual Framework}} & Presents a clear theoretical framework, either in words, graphs, or mathematically. Demonstrates an understanding of the mechanisms at play and generates testable hypotheses. & 15 & \\
\hline
\multirow{4}{*}{\textbf{Data Description}} & Describes the data sources accurately and clearly. Provides a thoughtful narrative about data selection, barriers to obtaining data, and why the data is suitable for the research question. & 15 & \\
\cline{2-4}
& Discusses measurement and unit of analysis effectively. & & \\
\hline
\multirow{4}{*}{\textbf{Empirical Strategy}} & Presents a well-defined research design and estimation strategy. Clearly explains the identification strategy, the variation exploited, and any necessary assumptions for causal inference. & 20 & \\
\cline{2-4}
& Addresses potential threats to identification and how they will be mitigated. & & \\
\hline
\multirow{2}{*}{\textbf{Conclusion}} & Restates the research question and its importance. Discusses the limitations of the research design and real-world policy implications. Suggests avenues for future research. & 10 & \\
\hline
\multirow{3}{*}{\textbf{Writing and Organization}} & The proposal is well-written, clear, and logically organized. There are no significant grammatical or typographical errors. & 10 & \\
\cline{2-4}
& Arguments flow logically from one section to the next. & & \\
\hline
\multicolumn{3}{|r|}{\textbf{Total Points Possible:}} & \textbf{100} \\
\hline
\end{tabular}
\end{table}

### 2024-08-29 assignments/research_proposal/research_proposal.tex
add lectures 1-5

--- 2024-08-29 assignments/research_proposal/research_proposal.Rmd
@@
----
-title: |
-        | Research Proposal
-subtitle: |
-        | Ph.D. Environmental and Development Economics
-        | UMN Applied Economics 2024
-        | Instructor: Raahil Madhok
-# Output type and options (no TOC and yes fig captions)
-output:
-  bookdown::pdf_document2: 
-    toc: false
-    fig_caption: yes
-    # This also keeps the .tex file that is used to make the output (we will eventually need this for publication)
-    keep_tex: yes
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\input{header.tex}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-# This includes latex arguments
-header-includes:
-  - \input{header.tex}
+\title{Research Proposal}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Ph.D.~Environmental and Development Economics\\
+UMN Applied Economics 2024\\
+Instructor: Raahil Madhok}
+\author{}
+\date{\vspace{-2.5em}}
 
----
+\begin{document}
+\maketitle
 
-# Assignment Overview
+\hypertarget{assignment-overview}{%
+\section{Assignment Overview}\label{assignment-overview}}
+
 The goal of this project is to take you through a mock process of publishing in economics, including writing a paper (proposal, in this case) as well as the revise and resubmit process.
-You will: 1) draft a research proposal, 2) send your proposal to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, and 4) submit the final draft for ``publication'' (i.e.
-a final grade...).
+You will: 1) draft a research proposal, 2) send your proposal to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, and 4) submit the final draft for ``publication'\,' (i.e.~a final grade\ldots).
 
-You **are not** allowed to ``double dip''. This means the idea must be original and you cannot use your second-year paper, or a proposal from another class.
+You \textbf{are not} allowed to ``double dip'\,'. This means the idea must be original and you cannot use your second-year paper, or a proposal from another class.
 
@@
 Broadly, you will identify an original and specific research question, explain why it is important, and outline the research design that you will use to address it.
-After receiving an ``R\&R'' on the paper, you will improve it and submit a final draft at the end of the semester.
+After receiving an ``R\&R'\,' on the paper, you will improve it and submit a final draft at the end of the semester.
 
-## Timeline
+\hypertarget{timeline}{%
+\subsection{Timeline}\label{timeline}}
+
 \begin{table}[ht]
@@
 
-## Helpful Resources
+\hypertarget{helpful-resources}{%
+\subsection{Helpful Resources}\label{helpful-resources}}
 
-- Marc Bellemare's guide on \href{https://marcfbellemare.com/wordpress/wp-content/uploads/2020/09/BellemareHowToPaperSeptember2020.pdf}{how to write applied economics papers}
-- Keith Head's \href{https://blogs.ubc.ca/khead/research/research-advice/formula}{introduction formula}
-- Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}
-- Read papers!
+\begin{itemize}
+\tightlist
+\item
+  Marc Bellemare's guide on \href{https://marcfbellemare.com/wordpress/wp-content/uploads/2020/09/BellemareHowToPaperSeptember2020.pdf}{how to write applied economics papers}
+\item
+  Keith Head's \href{https://blogs.ubc.ca/khead/research/research-advice/formula}{introduction formula}
+\item
+  Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}
+\item
+  Read papers!
+\end{itemize}
 
-## Format
+\hypertarget{format}{%
+\subsection{Format}\label{format}}
 
-- **Length:** 15 pages, double spaced (approx 4000 words)
-
-- **Format:** 12 point font, double spaced, 1-inch margins
-
-- **Submission:** PDF document (latex or word)
+\begin{itemize}
+\item
+  \textbf{Length:} 15 pages, double spaced (approx 4000 words)
+\item
+  \textbf{Format:} 12 point font, double spaced, 1-inch margins
+\item
+  \textbf{Submission:} PDF document (latex or word)
+\end{itemize}
 
 \newpage
-# Proposal Structure
 
-## Title
-The title should be concise and informative, reflecting the focus of your research. Please do not use one word titles (e.g. "Forests") or three word titles ("Forests and Development").
+\hypertarget{proposal-structure}{%
+\section{Proposal Structure}\label{proposal-structure}}
 
-## Abstract (250 words)
+\hypertarget{title}{%
+\subsection{Title}\label{title}}
+
+The title should be concise and informative, reflecting the focus of your research. Please do not use one word titles (e.g.~``Forests'') or three word titles (``Forests and Development'').
+
+\hypertarget{abstract-250-words}{%
+\subsection{Abstract (250 words)}\label{abstract-250-words}}
+
 The abstract should be no more than 10 sentences (about 250 words) and provide a clear and succinct summary of the paper.
@@
 
-## Introduction (3-4 pages)
+\hypertarget{introduction-3-4-pages}{%
+\subsection{Introduction (3-4 pages)}\label{introduction-3-4-pages}}
 
@@
 
-- **Contextualize the issue:** Provide a brief background on the topic within environmental and development economics.
+\begin{itemize}
+\item
+  \textbf{Contextualize the issue:} Provide a brief background on the topic within environmental and development economics.
 What are some general descriptive statistics about the issue at hand?
 Is there a puzzle, or something counterintuitive that your research may shed some light on?
-
-- **Research Question**: Clearly state your research question.
+\item
+  \textbf{Research Question}: Clearly state your research question.
 It should be specific, well-defined, and narrow enough to be addressed within the scope of the proposal.
 At the same time, it should ideally speak to a broader issue in environmental and development economics such that it contributes meaningfully to the literature.
-
-- **Importance**: Explain why answering the research question is important.
+\item
+  \textbf{Importance}: Explain why answering the research question is important.
 Think about its relevance for policy, environmental conservation, and economic development.
 Even if it is not directly policy-relevant, state what the intellectual importance is.
-
-- **Summary:** Summarize your proposal for the rest of the introduction.
+\item
+  \textbf{Summary:} Summarize your proposal for the rest of the introduction.
 Describe your data, measurement, and why your data is suitable for the research question.
 Describe your research design, why you chose it, and any identification assumptions you make.
-
-- **Literature Review:** Provide an overview of key literature on your topic.
+\item
+  \textbf{Literature Review:} Provide an overview of key literature on your topic.
 Identify gaps or unresolved questions that your research will address.
 Avoid providing a detailed summary of each paper, but rather focus on a key takeaway and how it relates to your topic.
-Also avoid ``punching down'' on the literature; instead, point out how your work will advance our understanding of the topic.
+Also avoid ``punching down'\,' on the literature; instead, point out how your work will advance our understanding of the topic.
+\end{itemize}
 
-## Conceptual Framework (2 page)
+\hypertarget{conceptual-framework-2-page}{%
+\subsection{Conceptual Framework (2 page)}\label{conceptual-framework-2-page}}
+
 You are required to include a theoretical framework to guide your research design. The goal is to guide my thinking, pin down potential mechanisms, and generate empirically testable hypotheses.
@@
 The model represents a framework to fix ideas and help think about the issue at hand.
-It **DOES NOT** need to be a mathematical model, although I will be happy if you did construct one.
+It \textbf{DOES NOT} need to be a mathematical model, although I will be happy if you did construct one.
 If you do write a mathematical model, you are welcome to adapt any of the ones we study in class, or replicate one from the literature (adapted to your context).
@@
 Otherwise, your model can be written in words, or even graphs.
-If your question is about how $x$ affects $y$, think about why you expect a relationship in the first place.
+If your question is about how \(x\) affects \(y\), think about why you expect a relationship in the first place.
 Is there an underlying behavioural assumption?
@@
 
-## Data (1-2 page)
-You are expected to choose a research question where **real** data is available to answer it.
+\hypertarget{data-1-2-page}{%
+\subsection{Data (1-2 page)}\label{data-1-2-page}}
+
+You are expected to choose a research question where \textbf{real} data is available to answer it.
 The data description need not be dry and boring; it can be an opportunity to present an interesting narrative.
-You can discuss **why** you chose a dataset.
+You can discuss \textbf{why} you chose a dataset.
 Remind the reader about what kind of data would be needed to answer your research question, and any barriers to obtaining that data.
@@
 
-- **Data sources**: List the data source for each variable. You are welcome to propose scraping the data, as long as the raw data exist (in pdf or online tables) and I can access it.
-
-- **Measurement**: Describe how you will measure each variable using your data.
+\begin{itemize}
+\item
+  \textbf{Data sources}: List the data source for each variable. You are welcome to propose scraping the data, as long as the raw data exist (in pdf or online tables) and I can access it.
+\item
+  \textbf{Measurement}: Describe how you will measure each variable using your data.
 If you are using a household survey, if the variable directly provided?
@@
 If you are using remote sensing data, briefly describe how the data product captures the variable you want to measure.
+\item
+  \textbf{Unit of analysis}: Desribe your unit of analysis. If your raw data is at a different level (e.g.~gridded), then think about how you will aggregate the data.
+\item
+  \textbf{Final Dataset}: Describe your (hypothesized) final dataseta and the panel dimensions.
+\end{itemize}
 
-- **Unit of analysis**: Desribe your unit of analysis. If your raw data is at a different level (e.g. gridded), then think about how you will aggregate the data.
+\hypertarget{empirical-strategy-4-pages}{%
+\subsection{Empirical Strategy (4 pages)}\label{empirical-strategy-4-pages}}
 
-- **Final Dataset**: Describe your (hypothesized) final dataseta and the panel dimensions.
-
-## Empirical Strategy (4 pages)
 In this section you will describe your research design, estimation strategy and your identification strategy.
 
-- **Estimation Strategy:** The estimation strategy comprises the equations you will use to answer your research question.
+\begin{itemize}
+\item
+  \textbf{Estimation Strategy:} The estimation strategy comprises the equations you will use to answer your research question.
 Clearly state which estimator you will use to recover your coefficient(s) of interest.
 If relevant, describe how your coefficient connects to your theoretical framework.
-
-- **Identification Strategy:** Clearly describe how your research design is set up to identify a causal relationship.
+\item
+  \textbf{Identification Strategy:} Clearly describe how your research design is set up to identify a causal relationship.
 State the variation that is exploited and how to interpret your coefficient of interest.
@@
 It may help to frame this as a discussion of threats to identification, and how your research design overcomes those threats.
-If a test is needed to improve confidence in your identification strategy (e.g.
-a balance test, or test of pre-trends), then write and describe the corresponding equations.
+If a test is needed to improve confidence in your identification strategy (e.g.~a balance test, or test of pre-trends), then write and describe the corresponding equations.
+\end{itemize}
 
@@
 
-## Conclusion (1 page)
-- Summarize your research question again and why it is important. Try not to re-state it in exactly the same way as in your introduction.
+\hypertarget{conclusion-1-page}{%
+\subsection{Conclusion (1 page)}\label{conclusion-1-page}}
 
-- Discuss any limitations of your research design. Aim to think deeply beyond standard ones such as lack of external validity.
-
-- Discuss real world implications of having answered your research question. Think about which policy-makers would be interested in your results.
-
-- Discuss implications for future research
+\begin{itemize}
+\item
+  Summarize your research question again and why it is important. Try not to re-state it in exactly the same way as in your introduction.
+\item
+  Discuss any limitations of your research design. Aim to think deeply beyond standard ones such as lack of external validity.
+\item
+  Discuss real world implications of having answered your research question. Think about which policy-makers would be interested in your results.
+\item
+  Discuss implications for future research
+\end{itemize}
 
 \clearpage
-# Grading Rubric
+
+\hypertarget{grading-rubric}{%
+\section{Grading Rubric}\label{grading-rubric}}
 

### 2024-08-29 lectures/week1/week1.Rmd
add lectures 1-5

---
title:  |
    | Environmental and Development Economics
    | Week 1 - Introduction
author:
  |
    | Raahil Madhok
    | UMN Applied Economics
date: "`r Sys.Date()`"

# Output type and options (no TOC and yes fig captions)
output: 
  beamer_presentation:
    theme: Montpellier
classoption: "aspectratio=169"

# This includes latex arguments
header-includes:
  - \AtBeginDocument{\title[Week 1 - Introduction]{\textbf{Environmental and Development Economics} \\ Week 1 - Introduction}}
  - \input{header.tex}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dev = "png") # NOTE: switched to png instead of pdf to decrease size of the resulting pdf
```

## Introduce yourself

- First, lets do introductions
\vfill

- Name, year, memorable summer activity, **research interests**
\vfill

- Why are you taking this class?

## Today

- Why study environmental economics in LMICs?
\vfill
- Course overview + detailed outline
\vfill
- Grade breakdown
\vfill
- Course website: https://github.com/rmadhok/enviro-dev-grad
  - lectures, assignments, syllabus

\vfill

##

\centering
\huge Why study environmental economics in LMICs?

## Why study environmental economics in LMICs?

- Environmental quality is worse and has worse consequences in LMICs
  - Highest pollution, highest deforestation

- New field: room for applied theory, empirical innovation

- Data breakthroughs
  - Measurement: remote sensing, DHS, etc.
  - Access: lower barriers to government access and experimentation

- Evidence needed; big implications for poverty alleviation

## Environmental quality worse in LMICs

```{r rstudio1, echo = FALSE, message = FALSE, warning = FALSE, out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/pm_lmic.png",
                        auto_pdf = TRUE
                        )

```

## Disease burden higher in LMICs

```{r rstudio2, echo = FALSE, message = FALSE, warning = FALSE, out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/disease_lmic.png",
                        auto_pdf = TRUE
                        )

```

## Is environmental economics different in LMICs?

My answer: sometimes...
\vfill

- Magnitudes
  - Same questions, but costs and benefits different

- **Local** environmental quality is more important

- Different topics
  - cookstoves, enforcement/corruption, ethnic favoritism

- Institutions and state capacity

##

\centering
\huge Course Overview

## Course Overview: There is no textbook

Instead, I am organizing around **\textcolor{orange}{FOUR}** key questions:

1. Why is environmental quality so bad in developing countries?

2. What are the costs of poor environmental quality in developing countries?

3. Why is WTP for environmental quality low in developing countries?

4. What are the political economy barriers to environmental protection?

## Course Approach

I will:

- Frame (almost) each topic with some theory

- Teach applied papers
  - research design, identification strategy, estimation techniques

- Emphasize recent JMPs

\vfill

I will NOT:

- Teach econometrics

- Teach coding

- Teach every topic in environment/development

## Course Goals

1. Show you environment/development research frontier

2. Inspire your thesis/JMP ideas

3. Advance your training as applied microeconomists

4. Show you what makes a top-tier research question

## Course Structure

- This is a brand new class, so I give myself leeway to make changes
\vfill
- You have the unique opportunity to determine direction of the course
  - Think about what topics do and don't interest you
  - And let me know!
\vfill
- Please check the course website regularly for updates

##

\centering
\huge Course Outline and Topics

## Weeks 1 and 2

**Week 1: Theory of Environment/Development**

- How to use theory to ask the right questions (lecture 2)

\vfill

**Week 2: The effect of development on the environment**

- air quality, water quality (lecture 3)
- forests, biodiversity (lecture 4)

## Weeks 3 and 4

**Week 3: The effect of the environment on development**

- human capital (lecture 5)
- firms (lecture 6)

\vfill

**Week 4: WTP for Environmental Quality**

- TBD (lecture 7)
- TBD (lecture 8)

## Weeks 5, 6, and 7

**Week 5: Environmental policy design**

- Spillovers (lecture 9)
- Pollution markets in developing countries (lecture 10)

\vfill

**Week 6: Political economy of the environment**

- Corruption and weak enforcement  (lecture 11)
- Environmental justice (lecture 12)

\vfill

**Week 7: Research presentations**

- I will provide more details throughout the semester

##

\centering
\huge Grade Breakdown

## Breakdown

\Large

\begin{table}[]
\begin{tabular}{lllll}
In-class presentations      & 10\% &  &  &  \\
Problem Set           & 20\% &  &  &  \\
Research Proposal & 60 \% &  &  &  \\
Participation                     & 10\%     &  &  & 
\end{tabular}
\end{table}

## In-class presentations (10%)

- At start of **first** lecture each week, you'll give a 15 min paper presentation

- Each student submits **seven** summary slides (5% of grade)
  - motivation, research question, methods, results
  - 10 mins presentation + 5 mins Q&A (5% of grade)

- I will select presenter on-the-spot 
  - **randomly** with replacement**

\vfill

** \footnotesize If you are never chosen, your grade is based on slides.

## Problem Set (20%)

- You will replicate an environment/development paper
  - You will also **extend** the results

- You will become familiar with coding in publication-quality papers

- You will use R or Stata

## Research Proposal (60%)

\begin{table}[]
\begin{tabular}{lllll}
Written Proposal      & 30\% &  &  &  \\
Peer Review           & 20\% &  &  &  \\
Proposal Presentation & 10\% &  &  &  \\
                      &      &  &  & 
\end{tabular}
\end{table}

- You will develop a research proposal for an original idea
  - You are NOT expected to actually do the analysis  
  - I will provide small deadlines (outline, first draft, etc.) along the way

- You will peer review each others proposals

- You will present the proposal at the end of the semester (30 mins)

## Participation (10%)

- I take this seriously

- Not enough to just show up to class

- Quality of questions/discussion count

##

\centering
\Huge Questions?

## Today

- **Guiding question:** Why is environmental quality so low in LMICs?

\vfill

- Your explanations

\vfill
- **Main goal:** Conceptual framework
  - Four *theory-informed explanations*
  - Set the stage for rest of class

\vfill

## Remember from last time

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/pm_lmic.png",
                        auto_pdf = TRUE
                        )

```

## Remember from last time

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/disease_lmic.png",
                        auto_pdf = TRUE
                        )

```

## Why is environmental quality low in LMICs?

- MWTP is low (paradox)
  - Berkouwer and Dean (2022): $12 for clean air
  - Kremer et al. (2013): ~ $4 for clean water
    - Imply VSL \$USD 860 vs \$USD 8.6 million for USA

- Do we take this as given? Perhaps status quo is optimal
  - is bad environmental quality another dimension of poverty?

- Is welfare loss from pollution greater in rich countries, even though they're cleaner?

- **What are your explanations?**

##

\centering
\Huge Theory-informed Explanations

\large Greenstone and Jack (2013)

## Aside: why is applied theory important?

- Builds structure for answering big (and small) questions

- Generates potentially unexpected insights w/ testable predictions

- In reverse: helps rationalize results

- Gets you into better journals (and better jobs)

- Field is headed that way (from my recent experience)

## Conceptual Framework of Environmental and Development Economics

- Social planner chooses optimal EQ where social $MWTP_{e} = MC$
  - Need to know MWTP for representative agent

\vfill

Set up:

- $n$ identical agents with utility from consumption, EQ, and health
- Initial wealth $y_{0}$, health $h_{0}$, environmental equality $e_{0}$
- health depends on self-protection, $s$, and $e$
- Assume perfect markets (i.e. no externalities)

## First Best

- Agent chooses $c$, $\Delta e$, and $s$ to maximize:

$$ U(e, h(s,e), c) \quad \quad \text{s.t.} \quad \quad y \geq c_{e}(\Delta e) + c_{s}(s) + c$$

- where wealth (endowment + income) and experienced EQ are:
$$y = y_{0} + \Delta y(e, h(s, e))$$
$$e = e_{0} + \Delta e + a(c, s)$$

- where $a(c, s)$ captures impact of $c$ and $s$ on EQ

## Model Particulars

- EQ affects utility directly through existence value
\vfill
- EQ affects utility indirectly via health (which also affects income)
  - e.g. pollution exposure affects productivity
  - This can be mitigated by self-protection, $s$ (e.g. mask, air purifier)
\vfill
- EQ affects income, which in turn affects utility via budget constraint
  - e.g. agricultural productivity
\vfill
- Experienced EQ depends directly on $\Delta e$, and indirectly via $c$, $s$
  - $a(c,s)$: defensive investments i.e. clean cookstove, bottled water, etc.

## MWTP for improving environmental quality

- Let $\lambda_{e} = \frac{\partial u}{\partial \Delta e}$, $\lambda_{y} = \frac{\partial u}{\partial c}$
- Set up lagrangian and solve for $MWTP_{e}$:
\begin{align}
MWTP_{e} = \frac{\lambda_{e}}{\lambda_{y}} 
 = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e} + \frac{\partial u}{\partial h}\frac{\partial h}{\partial e}\right) + \frac{\partial \Delta y}{\partial e} + \frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e} \notag
\end{align}

- aesthetic benefit from improved EQ (converted to dollars)
- indirect benefit of EQ for health (converted to dollars)
- direct impact of EQ on income and indirect impact via health
\vfill
**Note: if $U''(c)<0$, low $y \rightarrow$ high MUC ($\lambda_{y}$) and low $MWTP_{e}$**

## MWTP for self-protection

- Set up lagrangian and solve for $MWTP_{s}$

\begin{align}
MWTP_{s} & = \frac{\lambda_{s}}{\lambda_{y}} \notag \\ & = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e}\frac{\partial a}{\partial s} + \frac{\partial u}{\partial h} \left(\frac{\partial h}{\partial s} + \frac{\partial h}{\partial e}\frac{\partial a}{\partial s} \right)\right) + \frac{\partial \Delta y}{\partial e}\frac{\partial a}{\partial s} + \frac{\partial \Delta y}{\partial h}\left(\frac{\partial h}{\partial s} + \frac{\partial h}{\partial e}\frac{\partial a}{\partial s}\right) \notag
\end{align}

- indirect effect of $s$ on EQ and health (converted to dollars)
- indirect effect of $s$ on income via productivity and health
\vfill
**Note: if $U''(c)<0$, high $y \rightarrow$ low MUC ($\lambda_{y}$) and high $MWTP_{s}$**

## The Social Planner

- In first best, social planner sets MB = MC
  - where $MC_{e} = \frac{\partial c_{e}}{\partial \Delta e}$ and $MC_{s} = \frac{\partial c_{s}}{\partial \Delta s}$
\vfill
- But to aggregate over $n$, we must assume:
  - No preferences of her own
  - No market failures
  - Can observe true MWTP
  - Anything else?
\vfill
- Do these hold in LMICs?
<!-- political economy and market failures -->
<!-- what about externalities? this doesn't differ between HICs and LMICs. But they may be exacerbated by or interact with market failures -->

## Course Structure

- Set the stage: 
  - how does environment affect development ($\frac{\partial h}{\partial e}$) (week 2)
  - how does development affect the environment (week 3)

\vfill

- Bulk of course:
  - Explain why environmental quality low in LMICs
  - Identify as many parameters of the social planner problem as possible

\vfill

- Goal: where can you make a contribution?

## Why is environmental quality so low in LMICs?
Four explanations informed by the model:

1. High marginal utility of consumption

2. High marginal abatement costs -- includes state capacity

3. Political economy distortions (first best violation)

4. Market failures (first best violation)
    - frictions cause revealed MWTP $\neq$ true MWTP

##

\centering
\huge Preview of Answers

## 1. High marginal utility of consumption

- Intuitively, poor people care more about meeting basic consumption needs
\vfill
- Economically, agent trades off $c$ and $e$ by setting $u'(c) = u'(e)$
  - If $u''(c)<0$, prefer $c$ at lower levels of $y$
  - even if health benefits of $e$ are large!
\vfill
- **Very** few revealed preference studies on $MWTP_{e}$
  - Kremer et al. (2013) randomly clean up springs in Kenya
  - WTP USD 11/year for clean water; VSL of USD 860
\vfill
- Larger literature on $u'(h)$ also suggests low valuation (Berkouwer and Dean, 2022)
\vfill

## 2. High MC

- High MAC suggests sub-optimal environmental quality. Why?
    - Upward sloping MAC suggests low MC in poor countries

\vfill

- MC not only driven by MAC; also reflects weak state capacity
  - Enforcement (Duflo et al., 2013)
  - Incentives (Jagnani and Mahadevan, 2024; Gulzaar and Dipoppa, 2024)
  - Spillovers (Viera et al. 2024)

\vfill
- High MC **does not** mean deviation from first best

## 3. Political economy

- Social planner includes own utility weights social welfare function
  - i.e. corruption
\vfill
- Many examples from LMICs
  - pollution (Duflo et al., 2013)
  - deforestation (Burgess et al., 2012; Viera et al., 2024)
  - human-wildlife conflict (Madhok et al., 2024)
\vfill
- Leads to second best policy (inefficient)
\vfill

## 4. Market Failures

- This is partially a couse on development economics
  - About market failures: land, labor, credit, etc.
\vfill
- Implication for us: revealed $MWTP_{e} \neq$ first best $MWTP_{e}$
\vfill
- Example: weak property rights $\rightarrow$ underinvestment in $e$
  - Underestimate $MWTP_{e}$ from observed data
  - RCT evidence from crop-burning PES contracts: Jack et al. (2024)
\vfill

## Lots of room for research

- Environment and development economics is new
  - Challenge: find something unique about LMICs
\vfill
- Goal: identify model parameters 
\vfill
- Evidence on many parameters are absent
\vfill
- Barriers to research in LMICs are falling
  - remote sensing, administrtive/survey data, webscraping
\vfill

## Next week

- In-class presentations

- Impact of development on the environment (air, water)

- Impact of development on the environment (forests, biodiversity)

### 2024-08-29 lectures/week2/week2.Rmd
add lectures 1-5

---
title: "| Environmental and Development Economics\n| Week 2 - How does development affect the environment?
  \n"
author: |
  | Raahil Madhok
  | UMN Apllied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Week 2 - development and environment]{Environmental and Development
  Economics \\\\ Week 2 - How does development affect the environment?}}"
- \input{header.tex}
---

## 
\centering
\Huge Lecture 2

\large How does development affect the environment?

## Housekeeping

- State things that are due.

## Today

- **Guiding question:** how does economic development affect the environment?
\vfill
- Descriptive overview
\vfill
- Channel I: Income effects
  - Changes in consumption
  - Energy
  - Diet
\vfill
- Channel II: Technology and Infrastructure
  - Agricultural productivity
  - Infrastructure
\vfill
- Channel III: Institutions
  - Later in the course

\vfill

## How does economic development affect the environment?

- There is no one answer

- Choices that maximize economic growth often degrade environment

- But development expands our choice set (e.g. clean energy)

- Development also lowers $u'(c)$ (model from last week)

- Hence: the Environmental Kuznets Curve (EKC)

## Environmental Kuznets Curve

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "95%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ekc.png",
                        auto_pdf = TRUE
                        )

```

## Discussion: Do you believe EKC?

- Mechanisms?
\vfill

- Causality?
\vfill
- Robustness?
\vfill

## GDP and CO2
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_co2.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## GDP and PM2.5
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_pm.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## GDP and Ozone
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_ozone.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## My view

- Macro correlations give limited answer
\vfill
- Not the right question
  - What do we mean by development?
  - e.g. manufacturing $\uparrow$ pollution, but services may not

\vfill
- Narrow, micro/empirical papers more helpful
  - even if they don't find consensus

## Discussion: What is the ideal (quasi-) experiment?

- Can we randomize GDP?

- What variation in "development" can approximate the experiment?
  - what proxies development?
  - what does the proxy miss?
  - what (non-random) variation isolates causal relationship?
  - what does that variation miss?

## Discussion: What about mechanisms?

- **Why and how** does economic development affect the environment?

- Theory is your friend: especially total derivatives...
  - Sometimes enough to rationalize mechanisms
  - Otherwise generate testable predictions

- Heterogeneity is your friend
  - But have a reason **why** you choose one channel over another

## Alix-Garcia et al. (2013): Ecological footprint of poverty alleviation

**Question:** What is the effect of income on environmental  quality?
\vfill

- Poverty alleviation may $\uparrow$ demand for resource-intensive good
\vfill
- Or, it may $\uparrow$ demand for conservation

\vfill
- Empirical challenge: household income is endogenous
\vfill
- **This paper:** Exploit RDD eligibility to study impact of CCT on forests

## Why can't we regress forests on income?

- Income correlated with other things that affect environment

\vfill
- Like what?
  <!-- population agricultural returns, forest product prices, capital -->

\vfill
- If we control for OVB, will panel fixed effects work? 
  - No: small year-to-year income changes do not reflect ``development''
  - Predictable short-term income fluctuations different than permanent change

## Mexico's Oportunidades Program

- **Goal:** Ambitious program to build human capital

\vfill
- **Scope:** \$2.6 billion/year (half of anti-poverty budget)
  - Transfers are 1/3 of total income
  - Represents shift to new income path

\vfill
- **Size:** 4 million households treated
\vfill

- **Design:** Randomized rollout from 1998-2000
  - Afterwards, enrollment is non-random and based on **marginality index**
  - Two step: 1) village selected on index; 2) households selected in eligible villages

**Results Preview:** conditional cash transfer increases deforestation

## Data

- **Main variables:** village lat/lon, eligibility index, forest cover
\vfill
- **Sample size:** 105,749 villages
\vfill
- **Unit of analysis:** only village centroids provided?
  - How do we estimate village-level impacts? Thiessen polygons
  - Area around a point where every location closer to the point than to all others
  - Problems?

## Thiessen/Voronoi Polygons
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Data Visualization
Eligibility cutoff: -1.2
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "49%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ag2013_fig2.png","assets/ag2013_fig3.png"),
                        auto_pdf = TRUE
                        )

```

## Empirical Strategy

- OLS:
$$\Delta f_{i} = \alpha + \delta E_{i} + \beta'X_{i} + \epsilon_{i}$$
where $E_{i}=1$ if index $I_{i} > -1.2$

\vfill
- RDD: Sample window $-2 < I < -0.2$
\vfill
- Fuzzy RDD/IV:
$$\Delta f_{i} = \alpha + \delta T_{i} + \gamma I_{i} + \beta'X_{i} + \epsilon_{i}$$
$$\Delta T_{i} = \omega + \tau_{1}E_{i} + \tau_{2}E_{i}I_{i} + \tau_{3}M_{i} + \tau_{4}M_{i}I_{i} + \mu I_{i} + \Gamma' X_{i} + \epsilon_{i} $$ 
where $T_{i}$ is enrollment dummy, $E_{i}$ is eligibility cutoff dummy, $I_{i}$ is index, $M_{i}$ is dummy for region where enrollment increases rapidly

## Identification Assumptions

## Results: OLS and Sharp RDD
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Results: Fuzzy RDD / Instrumental Variables
- Very strong first stage: eligiblity $\uparrow$ probability of enrollment ($\tau_{1}=0.8$)
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Mechanisms

- RD results show that CCT increases deforestation
\vfill
- Why? What changes at the household level?
\vfill
- Approach 1: Use experimental sample (Progresa) with household survey data
\vfill
- Approach 2: Heterogeneity by road density to study role of market access

## Mechanisms: Consumption channel
- Before (1997-1998) and after (2000) data on consumption
- 506 villages, 320 treated, 186 control
- Treatment at the village level

Difference in differences:
$$ y_{it} = \gamma_{0} + \gamma_{1}Treat_{i} + \gamma_{2}Post_{t} + \gamma_{3} Treat_{i} \times Post_{t} + \epsilon_{it}$$

- $y_{it}$ = consumption
- cluster at locality level 
\vfill
**Hypothesis: $\gamma_{3}>0$**

## Mechanisms: Market channel
- Demand shock must be met by supply, which drives land use change
  - e.g. $\uparrow$ consumption of milk met by $\uparrow$ in grazing land
\vfill

- But $\gamma_{3}$ captures partial equilibrium
  - Part of demand shock supplied locally. What about rest?
\vfill

- Hypothesis: If consumption is driving deforestation
  - Then effect should be larger when infrastructure quality is low
  - Low market access: demand met by local supply
  - High market access: demand shock propagates across markets

\vfill

## Triple Differences:

Heterogeneity by infrastructure quality:
\begin{align*} 
y_{it} & = \beta_{0} + \beta_{1}Treat_{i} + \beta_{2}Post_{t} + \beta_{3}Treat_{i} \times Post_{i} \\ & + \beta_{4}Road_{i} + \beta_{5} Road_{i} \times Treat_{i} + \beta_{6} Road_{i} \times Post_{t} \\ & + \beta_{7} Road_{i} \times Treat_{i} \times Post_{i} + \epsilon_{it}
\end{align*}

- $Road_{i}$ = inverse road density (km of road w/n 10km of village/polygon)

\vfill

**Hypothesis: $\beta_{7} > 0$**
\vfill

## Results: Increase in consumption of land intensive goods
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig6.png",
                        auto_pdf = TRUE
                        )

```
Higher beef and milk demand (land intensive products)

Demand-side impacts do not vary with market access (we did't expect it to)

## Results: No increase in local production
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig7.png",
                        auto_pdf = TRUE
                        )

```
Supply-side impacts do not vary with market access

## Results: Deforestation higher in places with poor market access
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/ag2013_fig8.png",
                        auto_pdf = TRUE
                        )

```
\vfill
Bigger impact where supply response is localized

Supports mechanism of increased demand for land-intensive goods

## Thoughts?
- Form discussion group to think of alternate story

- Do we believe the story about increased demand for land-intensive goods?
  - Increased income relieves credit constraints
  - Expands capital, increases ag productivity on extensive margin b/c malfunctioning factor markets.
  - Market access may improve access to inputs, and reduce deforestation. 
  - Consistent with papers results.

- Only focused on one mechanism
  - returns to off-farm labor
  - income may increase aggregate production and lead to deforestation?
  - market access may lead to migration, less population pressure and deforestation
  - also consistent story

- This all points to shortcomings of RCTs in general
  - we cannot answer why

- But where is supply response coming from?
  - Need much more details spatial analysis
  - SLX model

- Road density can be capturing other dimensions of underdevelopment, not only market access

- How do we know that number of plots, and farm size increases are encroaching into forest? what if they displace other land types

- Policy implication limited by thiessen polygon
  - What is the natural unit for policy implementation?

## Same results in Colombia! Malerba (2020)

- Question: What is the impacts of CCT on consumption, energy, and deforestation?

- Context: Familias en Accion CCT project (2001-2005)

- Design: Matched Difference in Differences

- Result: Increased beef and milk (land intensive) consumption
  - Mediated by markets
  - Negligible $\uparrow$ in deforestation (counterintuitive)
  - No impact on CO2

## Research Design
:::::::::::::: {.columns}
::: {.column width="45%"}
- CCT program launched in 2000
  - Non-random: Municipalities selected on amenities (banks, education, health)
  - 721 eligible: slow phase-in

\vspace{0.3cm}
- Study sample: 5,477 households
\vspace{0.3cm}
- Variation: compare outcome in treated hh before/after enrollment relative to control hh
\vspace{0.3cm}
- **Identification assumptions**

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/malerba2020_fig2.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Results: Increased beed and milk consumption
Household DiD: $$ \Delta Y_{i} = \alpha + \delta T_{i} + X_{i} + \epsilon_{i} $$

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/malerba2020_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Effect muted with low market access (very weak)
- Why is sample size declining from 5,477?

## Design: Environmental Impacts

- Unit of analysis: Municipality
\vfill
- Design: municipalities enrolled gradually over time
  - Many were not enrolled by 2004
\vfill
- Define $Treat_{i} = 1$ if enrolled by 2004
- $Post_{t} = 1$ after treatment, zero for 2000

$$y_{it} + \beta_{0} + \beta_{1} Treat_{i} + \beta_{2} Post_{t} + \beta_{3} (Treat_{i} \times Post_{t}) + \beta_{4} X_{it} + \epsilon_{it} $$
\vfill
- Variation: $\Delta$ forest b/w 2000/05 for municipalities enrolled by 2004 rel. to non-enrolled

## Results: No environmental impact
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/malerba2020_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Weakly positive deforestation (counterintuitive)
- **Why?**

## Back to the question
How does development affect the environment?
\vfill

- Income 
  - Land intensive consumption
  - **Energy**

\vfill
- Next time: Technology and Infrastructure

## Development, Energy, and the Environment

- Energy ladder: as income $\uparrow$, move from solid fuel $\rightarrow$ gas $\rightarrow$ electricity
  - Electricity may displace dirty energy (wood)

- But, electricity also increases total energy (fridge, AC)
  - energy footprint increases with income

- Unless generated from solar, wind, etc

- People also buy cars as they get richer...

## GDP and Car Ownership
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "45%", fig.align = "center"}

knitr::include_graphics(
                        path = c("assets/ekc_cars.png","assets/ekc_legend.png"),
                        auto_pdf = TRUE
                        )

```

## Gertler et al. (2016): Demand for energy-using assets among middle class
:::::::::::::: {.columns}
::: {.column width="45%"}
- **Question:** How does income affect energy demand?
\vspace{0.3cm}
- **Motivation:** 1.3 billion people live without electricity
\vspace{0.3cm}
- Should we expect linear climb up the energy ladder as incomes rise?
  - what about credit markets?

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig1.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Non-linear relationship between income and fridges
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "65%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig2.png",
                        auto_pdf = TRUE
                        )

```
- Are fridges representative of energy-intensive assets?
- Can aggregating non-linearities $\rightarrow$ linearity?

## S-shape appears robust
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "50%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig3.png",
                        auto_pdf = TRUE
                        )

```

## How can we explain the S-shape?

**Set up:**

- Two periods with no discounting
- Agent $i$ can consume two goods
  - non-durable gives per period utility: $u'(\cdot)>0$, $u''(\cdot)<0$
  - lumpy durable that gives static per period utility R, if owned

- Durable price = P; non-durable is numeraire

- Let $Y_{1}$, $Y_{2}$ be per period 1, 2 income

- Let $\bar{Y} = 1/2(Y_{1}+Y_{2})$ be average income

## No credit constraints

Without credit constraints, if don't buy, total utility is:
$$u(\bar{Y}) + u(\bar{Y}) = 2u(\bar{Y})$$
If buy durable, spend $P$ and spread cost equally across periods:
$$2u(\bar{Y}-\frac{P}{2}) + 2R$$
Purchase durable iff:
$$ u(\bar{Y}) - u(\bar{Y}-\frac{P}{2}) \leq R$$
**$u''(\cdot)<0$ implies acquisition increases in income**

## 1) With credit constraints: Buy in period 1

If don't purchase, total utility is:
$$ u(Y_{1}) + u(Y_{2}) $$

If purchase in period 1: 
$$ u(Y_{1} - P) + 2R + u(Y_{2}) $$

Purchase in period 1 iff:

$$ \frac{u(Y_{1}) - u(Y_{1}-P)}{2} \leq R $$

## 2) With credit constraints: Wait to buy in period 2

If don't purchase, total utility is: 
$$ u(Y_{1}) + u(Y_{2}) $$
If save in period 1, and buy in period 2:
$$ 2u(\bar{Y} - \frac{P}{2}) + R$$
Wait to buy in period 2 if:
$$u(Y_{1})+u(Y_{2}) - 2u(\bar{Y} - \frac{P}{2}) \leq R $$

## Intuition of Model

- $u''(\cdot)<0$ means consumers gain from smoothing consumption

- Unconstrained household will buy in period 1
  - Use period 2 income (loan) to smooth consumption

- Credit constrained households cannot do this
  - If buy now, magnify consumption inequality across periods
  - Or, wait and buy in period 2, but delay utility gain

- First period buyers respond only to first period income (lemma 1)

- But increase in *cumulative* income increases buying through delay/saving (lemma 2)

## Testable Predictions

- Prediction 0: Acquisition increasing in income
- Prediction 1: S-shaped curve in acquisition
- Prediction 2: Faster income growth leads to more period 2 adoption
- Prediction 3: Period 1 ownership depends on interaction of income and growth

## Empirical Setting: Oportunidades (again)

- 320 randomly selected communities given early treatment (April 1998)
\vfill
- 186 randomly selected communities given late treatment (October 1999)
\vfill
- This paper: N=506 communities, 10,000 households surveyed (1997-2007)
\vfill
- Compare early and late households
  - Similar b/c both groups selected on ``vulnerability'' characteristics
  - Strongly balanced on covariates

## Empirical Design

- **Prediction 0:** Probability of asset purchase $\uparrow$ in income

$$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \beta X_{i} + \beta_{t}F_{i} + R_{rt} + \epsilon_{it}$$
where h(a) is the prob. that $i$ buys $a$ in time $t$ conditional on not having it in $t-1$. 
\vfill

- **Prediction 1:** S-shaped curve in acquisition
$$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{2} CI^{2}_{it} + \beta X_{i} + R_{rt} + \epsilon_{it}$$
Hypothesis: $\alpha_{2} >0$

## Empirical Design

- **Prediction 2 and 3:** conditional on having same level of cumulative income, households which accumulated income slower are less likely to acquire asset; this effect is increasing in cumulative income

$$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{3} Early_{i} + \alpha_{4} (Early_{i} \times CI_{it}) + \beta X_{i} + R_{rt} + \epsilon_{it}$$
- where $Early_{i}=1$ if $i$ began receiving transfers 18 months before control households

## Measurement issues

- Cumulative Income is sum of wage, farm, business income, CCT transfers
  - measurement error in wage data
  - cumulation requires interpolating between survey waves
\vfill
- Double counting problem if households invest CCT money into business
\vfill
- **Solution:** use cumulative CCT **transfers** to instrument cumulative **income**
  - Assumes transfers affect consumption only through income
  - Is this reasonable?

## Identifying Variation

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig4.png",
                        auto_pdf = TRUE
                        )

```

Use maximum cumulative transfer as **instrument** for actual (cumulative) transfer

Plus variation from early/late enrollment

## Results: Prediction 0
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig5.png",
                        auto_pdf = TRUE
                        )

```
- For every 10,000 pesos, probability of buy fridge $\uparrow$ by 4.7 pp.

## Results: Prediction 1
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gert2016_fig6.png",
                        auto_pdf = TRUE
                        )

```
- $\alpha_{2}>0$ implies convex relationship b/w cumulative income and asset purchase
- Since all households are poor, don't expect any to be past second inflexion point

## Results: Prediction 2 and 3
- include only if time

## Thoughts on paper?

<!-- We learned about effect of income on energy. What about environment? -->

## Lessons from income/consumption channel?

- What do we learn about how development affects the environment?

- What are shortcomings of research designs? Measurement?

- External validity?

## Research gaps

- Does earned income have different effect than transfers?
\vfill
- Which externalities have highest income elasticities?
\vfill
- Can the effect of income on $MWTP_{e}$ be separated from changes in consumption?

## Next week

- How does development affect the environment?
  - **Technology and Infrastructure**

##

\centering
\Huge Lecture 3

\large How does development affect the environment?

## Housekeeping

## Recap

\vfill
- Higher income leads to land-intensive **consumption**
  - Beef, milk, etc.
  - This requires clearing land for agriculture
  - Effect depends on market access
\vfill  
- Higher income leads to energy-intensive **consumption**
  - But the relationship is non-linear
  - Role of credit constraints

## Today

- Development increases access to **productive** capital
  - How do we think about environmental implications?
\vfill
- Case I: Development increases agricultural productivity
  - Abman et al. (2023)
\vfill
- Case II: Development spurs infrastructure development
  - Asher et al. (2020)

## Technology and infrastructure

- As economies develop, technology and infrastructure improve
\vfill
- Productivity boost means firms produce at lower cost
  - Outward shift in supply curve $\rightarrow$ \textcolor{brown}{bad for environment}
  - But, technological progress reduces resource intensity $\rightarrow$ \textcolor{green}{good for environment} 
\vfill
- Most active areas of research: 
  - agricultural productivity
  - infrastructure

## Agricultural productivity and the environment

- **Boserup's hypothesis **: extensive margin   
  - improvements in land productivity $\uparrow$ ag land value
  - leads to pressure on forests to clear land for ag
\vfill
- **Borluag's hypothesis**: intensive margin
  - *Under factor market constraints*, $\rightarrow$ productivity spurs intensification
  - ``spares'' land for nature and puts less pressure on forests
\vfill
- **Jevon's paradox**: general equilibrium
  - technology for increasing yield also lowers MC of producing given amount
  - Ag supply curve shifts out and may dominate gains in yield
  - Leads to net increase in land demand

**Takeaway: this is an empirical question!**

## Conceptual framework

- Continuum of agents with heterogeneous outside options
  - Changes in returns to ag affect number of agents who select into farming

- Farmers allocate labor b/w farming on existing land, or cultivate new land

- Let $\gamma$ denote household labor allocated to new land

- Production function for new and existing land are $f(\cdot)$ and $g(\cdot)$

- Exogenous productivity shock, $\Omega$, improves productivity of existing land more

$$ \frac{\partial g(l)}{\partial \Omega} \geq \frac{\partial f(l)}{\partial \Omega} \geq 0 \quad \quad \forall l$$

## Conceptual framework

- Farmers are factor market constrained (Conning and Udry, 2007)
  - Cannot hire labor from market

- Total deforestation given by new land cleared for agriculture:
$$ D^{*} = n^{*} \cdot \gamma^{*} $$

- Differentiating with respect to $\Omega$ gives:
$$\frac{\partial D^{*}}{\partial \Omega} = \underbrace{n^{*}\frac{\partial \gamma^{*}}{\partial \Omega}}_\text{$\leq 0$} + \underbrace{\gamma^{*}\frac{\partial n^{*}}{\partial \Omega}}_\text{$\geq 0$}$$

- What does each term represent and what is the sign?

## Abman et al. (2023): Agricultural Productivity and Deforestation

- **Goal:** To test $\frac{\partial D^{*}}{\partial \Omega}$, we need random variation in $\Omega$

- **Question:** what is the impact of ag productivity on deforestation?

- **Context:** Large scale ag extension program in Uganda (2008-2013)

- **Design:** Leverage Spatial discontinuity in village eligibility

- **Results:** improvement in agricultural productivity; reduction in deforestation

## Background: Why Uganda?

- Forest cover shrunk from 24\% to 9\% from 1990-2015

- Most forest land is **privately owned**
  - Land owners allowed to covert forest (Land Act, 2010)
  - Cannot use state powers to limit deforestation

- Need an indirect mechanism to incentivize conservation

## Background: BRAC Extension Program

- Launched in 2008

- Aims to extend modern cultivation techniques to smallholders

- Two step treatment:
  - Training: ``model farmers'' trained in modern cropping, irrigation, weeding, pest control. Then, set up demo plot and train others
  - Extension: Promoters selected to sell subsidized HYV seeds in their villages

- **Key feature: Program limited to villages within 6km of each BRAC branch**

## Data

- Deforestation from GFC (Hansen et al., 2013)
  - Pro: high resolution (30m)
  - Con: 1) only captures loss not gain, 2) everything relative to 2000

- Lat/lon of villages within 12km of BRAC office (N=807 villages)

- Attribute forest pixel to village if w/n 400m of village centroid
  - Rationale: median household distance to village center = 400m
  - Problems?

- Agricultural survey on 7,781 households (451 villages) in 2011
  - Detailed cropping practices from last two seasons

## Empirical Design: Spatial Regression Discontinuity

- Estimate IIT since we do not know true treatment designation 
\vfill
- Local linear regression to estimate left and right side of discontinuity

$$ \beta = lim_{z \uparrow 0} E[Y|z_{i} = z] - lim_{z \downarrow 0} E[Y|z_{i} = z]$$

- where running variable, $z_{i}$, is distance of village from 6km cutoff
  - $z \leq 0$ means village had access to extension
\vfill
- For estimation, use 2km bandwidth 
  - Higher regression weights on points nearer to cutoff (triangle kernel)

## Aside: RD with local linear regressions

- Left side of cutoff
$$ Y_{i} = \alpha_{L} + \beta_{L}(z_{i}-c), \quad \quad \text{for $z_{i} \leq$ c} $$
- Right side of cutoff
$$ Y_{i} = \alpha_{R} + \beta_{R}(z_{i}-c), \quad \quad \text{for $z_{i} \geq$ c} $$
- Treatment effect at threshold:
$$ \beta = \alpha_{R} - \alpha_{L}$$
since $\alpha$ are the intercepts on each side of cutoff

## Identification Assumptions?

<!-- eligibility uncorrelated with unobserved factors that drive deforestation and vary at the cutoff -->

<!-- what if baseline forest cover increases around 6km, and more degraded areas selected for program? -->

## Identification Assumptions
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Identification Assumptions
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Results: 14\% less deforestation in treated villages
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Mechanisms: Intensification
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "97%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Techniques to address nutrient depletion **on existing land**
- No evidence of increased fertilizer, pesticides, or HYV seed adoption

## Intensification vs. Extensification
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/abman2020_fig5.png",
                        auto_pdf = TRUE
                        )

```
- Revenue/acre is a proxy for yields
- No extensification $\rightarrow$ binding constraint on land clearing

## Discussion

- Empirical issues
- Alternative explanations
- External validity (recall forest is private in Uganda)

## Back to the Question
How does development affect the environment?
\vfill

- Income 
  - Land intensive consumption
  - Energy

\vfill
- Access to capital
  - Technology 
  - **Infrastructure**

## Infrastructure and the Environment
- Roads can increase deforestation by: 
  - $\downarrow$ MC (transport costs) of forest products, increasing supply
  - $\uparrow$ market access for agricultural products, causing extensification
  - $\uparrow$ land value, leading to agglomeration effects
\vfill
- Roads can increase forest cover by:
  - $\uparrow$ access to forest product substitutes (i.e. clean fuel vs. firewood)
  - $\uparrow$ access to external labor/capital markets, lowering need for land-clearing
\vfill
- Net effect is ambiguous

## What about Infrastructure and Air Pollution?

- Lets build a conceptual framework
- Talk to your neighbor for 10 mins

## Garg et al. (2023): Rural Roads, Labor Exits, and Crop Fires

- **Question:** what is the impact of roads on air quality?
  - what is the role of labor reallocation as a mechanism?

- **Context:** India builds last-mile roads using population cutoff

- **Design:** Regression discontinuity with crop fires as outcome

- **Results:** Roads $\uparrow$ crop fires and PM2.5 levels in a village

- **Mechanism:** Roads $\rightarrow$ labor exit $\rightarrow$ higher wages $\rightarrow$ labor-saving technology

## Background

- Labor is misallocated in agriculture in developing countries
  - $MP_{labor}$ lower in agriculture than other sectors

- Sources of misallocation is a big question in dev econ

- Many policies aimed at reducing misallocation
  - e.g. road-building (improve market access)

- Do these have externalities?
  - **innovation**: where do the externalities come from?

## Context: India

- 14 of the 20 most polluted cities in the world are in India

- Important pollution source: crop burning

- Pros:
  - Clear rice straw residue \& undergrowth b/w cropping seasons
  - Cheaper than labor/capital alternative

- Cons:
  - Carries civil/criminal penalty (weakly enforced)
  - Reduces soil nutrients and decreases ag productivity
  - Damages human health

- Roads may facilitate agricultural labor loss
  - Will farmers turn to crop burning, despit the costs?

## Research Design

- PMGSY aims to provide all-weather roads to all unconnected villages in India
\vfill
- Prioritize larger villages first
  - Connect villages with population > 1000 by 2003
  - Connect villages with population > 500 by 2007
  - Connect villages with population > 250 afterwards
\vfill
- State-by-state basis
  - If connected all large villages, can proceed to smaller ones immediately

## Data
- Roads: SHRUG database (N=11,151 villages w/o paved road at baseline)
\vfill
- Fires: NASA EODIS 
  -\# fires w/n 10km of village polygon
\vfill
- Pollution: Van donkelaar et al. (2016) measure PM2.5 at 10km resolution
\vfill
- Infant mortality: NFHS-IV (2015-16)
  - child-birthyear panel from 2001-2013
  - Identify survey clusters (villages) within 50km of each PMGSY village

## Empirical Strategy: Fuzzy RDD
First Stage:
\begin{align*}
  Road_{vdst} & = \gamma_{0} + \gamma_{1} 1(pop_{vds} \geq T) + \gamma_{2}(pop_{vds} - T) \\ & + \gamma_{3}(pop_{vds}-T) \times 1(pop_{vds} \geq T) + \theta X_{vds} + \mu_{d,h} + \rho_{t} + \epsilon_{vdst}
\end{align*}

Second Stage:
\begin{align*}
  Y_{vdst} & = \beta_{0} + \beta_{1} Roads_{vdst} + \beta_{2}(pop_{vds} - T) \\ & + \beta_{3}(pop_{vds}-T) \times 1(pop_{vds} \geq T) + \delta X_{vds} + \eta_{d,h} + \omega_{t} + \epsilon_{vdst}
\end{align*}

- $Road_{vdst}=1$ if village $v$ received road by 2013
- $Y_{vdst}$ is outcome (fires, pollution, mortality)
- $pop_{vds}$ is village population in 2001
- $T$ is treatment threshold (500 or 1000, depending on state)

## Covariate Balance
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig2.png",
                        auto_pdf = TRUE
                        )

```
- Road access increases annual crop fires by 60\%
- Road access increase PM2.5 level by 0.5 $\mu g/m^{3}$ (1.1\%)

## Does pollution result also capture increased driving on new roads?
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Mechanism: Roads increase ag wage rate
- Data: REDS Survey (1999 and 2006)
- Design: Diff-in-Diff (221 villages, 100 districts, 17 states)
  - Treat $=1$ if $v$ received road b/w 1999-2006; Post=1 in 2006
$$ Y_{vt} = \alpha + \beta_{1}(Treat_{v} \times Post_{t}) + \gamma_{v} + \theta_{t} + \epsilon_{vt}$$
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Mechanism: Roads induce labor exit
Back to RDD design:
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig6.png",
                        auto_pdf = TRUE
                        )

```
- More labor exit in districts with lower baseline ag wages

## Mechanism: places with more labor exit burn more biomass
```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/garg2023_fig5.png",
                        auto_pdf = TRUE
                        )

```
- Fire/pollution impact greater in places with lower baseline ag wage
- Road generates more labor exit in these places

## Discussion

- Do you believe the story?
- Alternative explanations

## Next week

- Presentations
- How does the environment affect development?
  - **Health**
  - Productivity

### 2024-08-29 lectures/week3/week3.Rmd
add lectures 1-5

---
title: "| Environmental and Development Economics\n| Week 3 - How does the environment affect development?
  \n"
author: |
  | Raahil Madhok
  | UMN Apllied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Week 3 - environment and development]{Environmental and Development
  Economics \\\\ Week 3 - How does the environment affect development?}}"
- \input{header.tex}
- \usepackage{gensymb}
---

## 
\centering
\Huge Lecture 4

\large How does the environment affect economic development?

## Housekeeping

- State things that are due.

## Today

- **Guiding question:** how does the environment affect economic development?
\vfill
- Today's focus: health effects

\vfill
- Next time: productivity and direct income effects

## Recall our model of environment and development
\begin{align}
MWTP_{e} = \frac{\lambda_{e}}{\lambda_{y}} 
 = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e} + \frac{\partial u}{\partial h}\frac{\partial h}{\partial e}\right) + \frac{\partial \Delta y}{\partial e} + \frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e} \notag
\end{align}

- Most research in environment/development economics seeks to identify $\frac{\partial h}{\partial e}$
  - At this point, $\frac{\partial h}{\partial e}>0$ is unambiguous
  - Focus is on **magnitudes**
\vfill
- What about indirect effects via productivity and health ($\frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}$)? 
  - This is the research frontier
  - Next lecture

## Lets formalize our intuition
- Weather can harm health directly (physiological stress)
  - Or indirectly, e.g. by lowering incomes of those in weather-dependent (outdoor) jobs

- There is a market for defensive goods (e.g. medicine, AC, etc.) which $\uparrow$ P(survival)

- But in developing countries, credit constraints restrict consumption smoothing

- Lets incorporate this into a model and study the comparative statics
  - Goal: exploain how weather affects health, considering financial constraints
  - Health is endogenously determined by consumption and health investments

## Setup
- **Agents:** value consumption $c_{it}$ and health $h_{it}$.
- **Objective:** Maximize lifetime utility
$$\mathbb{E}_{0} \left[\sum_{t=0}^{\infty} \beta^{t}u_{i}(c_{it}, h_{it}) \right]$$
\vfill
- **Health Status**: Determined by health investments ($q_{it}$) + exogenous weather ($z_{it}$)
$$h_{it} = h_{i}(\underbrace{q_{it}}_{(+)}, \underbrace{z_{it}}_{(-)})$$

## Setup

- **Production**: $c_{it}$ produced by $f(\cdot)$, and $q_{it}$ traded at price $P_{t}$
  - Agent endowed with labor $a_{it}$, sold at wage $\omega_{t}$
  - Labor can be affected by weather ($z_{it}$)
$$a_{it} = a_{i}(z_{it})$$

\vfill

- **Financial Constraints**: Savings ($x_{it}$) and borrowing are limited by:
  - Minimum savings: $x_{it} \geq -b_i(r_t)$
  - Maximum savings: $x_{it} \leq b_i(r_t)$
\vfill

- **Budget Constraint**: Income is used for consumption and health investments:
  $$x_{i,t+1} \leq r_t x_{it} + w_t a_i(z_{it}) - P_t q_{it} - c_{it}$$

## Setup

- **Market Equilibrium**: set optimal demand for labor/capital = aggregate supply  
  - where $K_{t} = \int x_{it}di$
  - wages $w_t$ and interest rates $r_t$ are determined in equilibrium as:
  $$w_t = w(K_t), \quad r_t = r(K_t) $$

## Mechanisms relating weather to health

- **Direct health effect:** random shock $z_{it}$ enters health directly: $h_{it} = h_{i}(q_{it}, z_{it})$
  - Physiological channel: extreme temperature stresses cadiovascular/respiratory system
  - Impact can depend on region, or proximity to water (where pathogens thrive in hot weather), denoted by agent-specific function $h_{i}(\cdot)$
\vfill
- **Indirect health effect:** through productivity $a_{i}(z_{it})$
  - Likely in rural areas, where ag workers are outside
\vfill  
- **Mitigation:** agent minimizes neg. utility shock from $z_{it}$ by purchasing $q_{it}$.
  - This could also be opp cost of working inside (at lower $w_{it}$) to avoid outdoors
  - Effect of mitigation limited by $\frac{\partial h_{i}(q_{it}, z_{it})}{\partial q_{it}}$
\vfill
- **Smoothing:** agent can also mitigate by reallocating resources b/w periods
  - Agents with near-binding borrowing constraints experience worse shocks

## Comparative Statics
- **Object of interest**: aggregate correlation b/w health \( h_{it} \) and weather \( z_{it} \):
    $$H_t = \left| \frac{\int h_{it} z_{it}^{(T)} di}{\int h_{it} di} \right|$$
\vfill
- **Rural Areas**: Consider a set of changes that:
  - increases responsiveness of $a_{i}(z_{it})$ to $z_{it}^{T}$.
  - raises prices for health goods $P_t$ above that of urban 
  - lower incomes by reducing $a_{i}(z_{it})$ for any $z_{it}$
  - Any one of these leads to increase in $H_{t}$
\vfill
**Prediction:** rural areas will have stronger correlation b/w weather and health

## Effect of Bank Access

- **Relaxing financial constraints** increases either or both of $(-b_{i}(r_{t}), b_{i}(r_{t}))$
  - Partial credit modeled as increase in $b_{i}(r_{t})$ only (i.e. can save more)
<!-- loans typically for production only -->
  - Allows agents to smooth consumption; draw down savings to mitigate weather shock
  - Reduces the correlation between weather shocks \( z_{it} \) and health outcomes \( h_{it} \).
\vfill

- **Prediction**: $\text{Prediction: } H_t \downarrow \text{ as } b_i(r_t) \uparrow$
  - Areas with better financial access have smaller effect of $z_{it}$ on health

## What is the effect on weather on health?

- What is the ideal experiment?

- What variation in environmental quality can approximate the experiment?
  - How would you measure environmental quality?
  - What might this measure miss?
  - What variation isolates a causal relationship?
  - What does that variation miss?

## Climate change versus weather

- Climate change is long term, weather is short term

- Studying impacts of weather is easier:
  - Usually use high dimensional fixed effects
  - E.g. with district and state-year FEs, exploit monthly deviations from average

- Studying impacts of climate is harder:
  - There is no good counterfactual for climate change
  - New method: long differences (Burke and Emerick, 2016)

## Deschenes et al. (2013): Weather and Death in India

- **Research question:** how do temperature shocks affect mortality?
  - How does that different between urban/rural areas?
  - What is the role of income as an intermediary?

- **Setting:** Indian districts from 1957-2000

- **Data:** daily gridded weather + annual infant mortality

- **Design:** panel fixed effects
  - Difference-in-differences with rural bank expansion

- **Result:** 1) heat $\uparrow$ mortality, 2) driven by rural populations, 3) credit access mitigates effect by facilitating consumption smoothing

<!-- ## Conceptual Model -->
<!-- - Goal: distinguish two channels thru which weather affects health -->
<!--   - Direct: physiological response -->
<!--   - Indirect: lower income (less health spending) -->

<!-- - Idea: More weather-dependent production $\rightarrow$ worse income shocks from high temps -->
<!--   - Implies rural people suffer more b/c cannot smooth consumption -->

<!-- - Extend classic incomplete market macro model in two ways: -->
<!--   - health, which depends on shocks and choices -->
<!--   - savings constraints -->

<!-- - In response to shocks -->
<!--   - reallocate resources b/w health and other consumption -->
<!--   - reallocation resources across periods (smoothing) -->

## Empirical Implications

- Remember permanent income hypothesis? 
  - Income shocks in $t=0$ caused by weather shock only affects contemporaneous survival via extent to which income shock reduces NPV lifetime income 
\vfill
- From model: if two agents face same shock at $t=0$, and one has income stream that depends on weather
  - Then impact of shock on his death is stronger
\vfill
- What if income shocks are at high frequency?
  - Will weather shock be enough to affect remaining lifetime income (and survival)?
  - This is why studying climate vs. weather is important

## Data

Rural

- Yields: annual, district level output/cultivated area (27 crops) (World Bank)

- Ag price index: mean crop price weighted by value in district (World Bank)

- Ag wages: daily wage rate (World Bank)

Urban

- Manufacturing productivity: output per worker (ASI)

- Urban CPI: NSS survey

- Manufacturing wages: ASI (nominal wage / CPI)

## Empirical Strategy
Flexible specification to model effect of daily temperature:
$$ Y_{dt} = \sum_{j=1}^{11} \theta_{j}TMEAN_{dtj} + \sum_{k} \delta_{k}1\{RAIN_{dt} \text{ in tercile k\}} + \alpha_{d} + \gamma_{t} + \lambda_{r}^{1}t + \lambda_{r}^{2}t^{2} + \epsilon_{dt} $$

- $r$ is a climatic region
- $TMEAN$ is days in year $t$ where: 
  - daily mean temp $<70\degree$, $\geq 97\degree$, or in the nine 3$\degree$ bins in between 
- $\theta_{j}$'s capture non-linearities, despite having annual unit of analysis
- $RAIN$ accounts for correlation b/w temperature and rainfall

## Identifying Assumptions?

- Impact governed by daily mean alone

- Impact of days' mean temperature on annual mortality is constant within 3$\degree$ intervals
  - This is reasonable

- By using number of days in each bin, assume that the sequence of hot and cold days is irrelevant.

- Annual deviation from district average is random

- Anything else? Reverse causality?

## Secondary Approach
- consolidates same information
- workhorse model nowadays because more parsimonious

$$Y_{dt} = \beta CDD 80_{dt} + \sum_{k=1}^{3} + \sum_{k} \delta_{k}1\{RAIN_{dt} \text{ in tercile k\}} + \alpha_{d} + \gamma_{t} + \lambda_{r}^{1}t + \lambda_{r}^{2}t^{2} + \epsilon_{dt}$$

- where CDD80 is cumulative degree days in district that exceeded 80$\degree$. 
- All bins $j < 80\degree$ are zero 
  - The three above $80 \degree$ restricted to be linearly increasing in average temp  
- Assumption: days when mean daily temp below 80 irrelevant for determining Y

## Results: Yields
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig1.png",
                        auto_pdf = TRUE
                        )

```
A single day with temperature over 85$\degree$ reduces yields by 0.5\%

## Results: Wages
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig2.png",
                        auto_pdf = TRUE
                        )

```
A single day with temperature over 85$\degree$ reduces ag wages by 0.25-0.5\%

## Results: Urban vs. Rural
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig4.png",
                        auto_pdf = TRUE
                        )

```
What is happening? Direct (agonomical) effect should not vary

## Results: Mortality
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/des2013_fig3.png",
                        auto_pdf = TRUE
                        )

```
Rural: years with more days with exceeding 87$\degree$ have 0.5\% higher mortality

## Bank Access
- If income $\rightarrow$ consumption smoothing, then it should mute relationship
  - Mechanism: lifetime utility maximized when $u'(c)$ equated across $t$ (Euler) 
\vfill
- Test: Between 1977-1990, central bank extends banking to the poor
  - Existing banks required to open four branches in unbanked areas
\vfill
- Method: compare weather-death relationship in banked/unbanked areas
  - Difference in differences
  - Identification problem?
\vfill
- Instrument: SEE BURGESS AND PANDE (2005)

## Equation

- See BURGESS AND PANDE (2005)

## Results: Credit Access

- On full sample, one more bank has no effect on weather-death relationship
  - Is single bank opening enough to enable smoothing?
\vfill
- Make dummy equal to one for 75th pctile of banks, zero for 25th pctile
  - Moving from 25th to 75th pctile of banks eliminates effect of temperature!
\vfill
- Mitigating effect of banks concentrated within growing season
  - Smoothing by farmers key mechanism
\vfill
- **Takeaway: credit enables poor to withstand environmental shocks**

## Discussion
- aggregated data (district) with satellite temperature makes it hard to distinguish atmospheric/ambient means from experienced exposure
- How can we then measure direct effect through physiological channel
  - Leave it to the epidemiologists...

## Back to the question
- How does environmental quality affect economic development?
\vfill
- Most reduced form studies identify $\frac{\partial h}{\partial e}$
\vfill
- Recall, this is an input into estimating benefits (MTWP) of $\uparrow e$
  - which we need to know in order to set optimal environmental policy
  - optimal in the sense that agents decisions are first-best
\vfill
- Is physical health the only thing that matters for estimating $\frac{\partial h}{\partial e}$?

## Chen et al. (2024): Air Pollution and Mental Health

- **Research question:** What is the impact of air pollution on mental health?
  - Direct mechanism: oxidative stress
  - Indirect mechanism: labor productivity, income, physical activity

- **Setting:** China in 2014-2015

- **Data:** Kessler Physchological Distress Scale
  - PM2.5 from China National Environmental Monitoring Center

- **Design:** instrument pollution with thermal inversions

- **Results:** Pollution $\uparrow$ depression
  - Mechanisms: lower exercise, worse physical health
  - Total cost = \$22.88 billion, similar to mortality costs!

## Background

- Why do we care about mental health?
  - Poverty associated with life satisfaction and **stress**
  - These mechanisms affect productivity, which perpetuates poverty
  - See Haushofer and Salicath (2023); Haushofer and Fehr (2014)
\vfill
- Frontier in development economics: RCTs on income and stress
  - Cash transfers $\uparrow$ subjective well-being (McGuire et al., 2022)
  - Cash transfers $\downarrow$ suicide in Indonesia (Christian et al., 2019)
  - Cash transfers $\rightarrow$ depression (Haushofer and Shapiro, 2016)
\vfill
- Toxicology literature: pollution $\uparrow$ oxidative stress (Calderon-Garciduenas et al., 2003; Sørensen et al., 2003; MohanKumar et al., 2008, Power et al., 2015)

## Background
- Theory of change: **pollution $\rightarrow$ stress** $\rightarrow$ productivity
  - Poverty correlated with pollution and stress
  - Direct effect of pollution on productivity (Graff Zivin \& Niedell, 2012; California)
  - Many parts of the chain are open areas for research!

## Air Pollution and Mental Health

What are the empirical challenges?

1. OVB: pollution correlated with economic activity
    - Richer counties more polluted
    - Induces **negative** bias
    - Solution: ?

2. ?

3. ?

4. ?

## Instrumental Variable: Thermal Inversion

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/inversion2.png",
                        auto_pdf = TRUE
                        )

```

## Instrumental Variable: Thermal Inversion
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/inversion.png",
                        auto_pdf = TRUE
                        )

```

## Instrumental Variable: Thermal Inversion

- Air is warmer near surface, becomes cooler at higher altitudes

- Thermal inversion: temperature-altitude gradient is reversed
  - Rare phenomenon
  - Depends on **random** meteorological factors

- Cool air is denser and does not rise

- Pollution becomes trapped near surface

## What do you think?

- Pros and cons

- Inversions are short lived (last for a few hours)
  - cannot estimate impact of sustained exposure
  - Is mental health affected by rare short term events?

- correlated with weather, which can have direct effect on mental health
  - control for ground level temperature/rain

- If mechanism is through chronic condition, then TI may isolate wrong type of variation

- TI are very localized phenomena. Is this what we want to study?

## Data

- Mental health: 2014 China Family Panel Study Survey
  - 16,000 rural and 13,000 urban residents (162 counties out of 1,355)
  - Six questions about mental health on 0-24 Scale
  - Depression, nervousness, restlessness, hopelessness, effort, worthlessness
\vfill
- Pollution: PM2.5
  - Webscraped
  - Validate with satellite data?
\vfill
- Thermal inversion (MERRA-2 Air Temperature Layers)
  - Inversion = 1 if temperature of first layer (110m) $<$ second layer (320m)
  - IV: number of inversions in month prior to each interview

## Inversions and Pollution
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig1.png",
                        auto_pdf = TRUE
                        )

```
Do thermal inversions seem rare and random?

## Empirical Strategy

First Stage:

$$ P_{ict} = \alpha_{0} + \alpha_{1}I_{ict} + f(W_{ict}) + \gamma_{c} + \sigma_{t} + \mu_{ict} $$

Second Stage:
$$ H_{ict} = \beta_{0} + \beta_{1}P_{ict} + f(W_{ict}) + \gamma_{c} + \sigma_{t} + \epsilon_{ict} $$

- $P_{ict}$ is PM2.5 in month prior to interview
- $I_{ict}$ is instrument, number of thermal inversions in past month
- $H_{ict}$ is total mental illness score (0-24)
- $f(W_{ict})$ are flexible weather controls
- $\sigma_{t}$ are year-month FEs (**important**!)

## Main Results: Mental Health
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig2.png",
                        auto_pdf = TRUE
                        )

```
- 1 $\mu/m^{3}$ PM2.5 $\uparrow$ mental illness by 0.1164 units (4\% of mean; 0.4$\sigma$)
- 1 $\mu/m^{3}$ PM2.5 $\uparrow$ prob. of several mental illness by 0.42pp (9.5\% of mean)

## Main Results: Symptoms
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Not driven by any one particular symptom

## Main Results: Persistence
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Effect of pollution exposure on mental health dissipates over time

## Mechanisms
- Focus on indirect effects (physical health, productivity, behavior)

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig5.png",
                        auto_pdf = TRUE
                        )

```
- Driven by reduction in physical exercise, not other illness
- Very small impact compared to pollution effect (0.002$\sigma$ vs. 0.4$\sigma$)

## Heterogeneity
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/oliva2024_fig6.png",
                        auto_pdf = TRUE
                        )

```
- Why are effects larger for men?

## Economic Cost? Back of the Envelope

- 1$\sigma$ $\uparrow$ in PM2.5 leads to 9.5\% $\uparrow$ in prob. of severe mental illness
\vfill
  - About 0.095 $\times$ 1.4 billion = 13.3 million additional people 
- Annual cost per patient of mental illness in China is \$USD 3,665 (Xu et. al 2016)
\vfill
- Assume 8\% of patients with mental illness get treated (Phillips et al., 2009)
\vfill
- **Economic cost of mental illness:** $13.3 \times 3665 \times 0.008 \times 12 = 46.7$ billion USD
\vfill
- Mortality cost of $1\sigma$ PM2.5 is \$USD 30 billion in USA (Deryugina et al., 2016)

\vfill
**Issues with the back of the envelope calculation?**
<!-- 1) for severe mental illness; 2) what about non-medial costs? productivity
3) hard to interpret in sd. do a counterfactual with policy-relevant change in pollution -->

## Discussion

- small sample
- few counties
- Small time window
  - cannot include year FEs or control for annual macro trends
- exploit short term variation (daily)
- mechanism unclear

## 
\centering
\huge Lecture 5

\large How does the environment affect economic development?

## Housekeeping

- State things that are due

## Today

- **Guiding question:** how does the environment affect economic development?
\vfill
- Today's focus: productivity and income
  - worker productivity
  - absenteeism

## Pollution and Productivity

- We know temperature is neg. associated w/ output (Dell, Jones, and Olken 2012)
\vfill
- We also know high temperatures are associated with lower crop yields
  - Mendelsohn and Dinar (1999), Auffhammer et al. (2006), Schlenker and Roberts (2009),
  Lobell et al. (2011), and Gupta et al. (2017)
\vfill
- What are the mechanims?
  - heat stress, mortality, conflict
\vfill
- Open and policy-relevant research area
  - If we know importance of each mechanism, we know where to target policy

## Somanathan et al. (2021): Temperature, Productivity, and Labor

- **Research question:** What is the effect of temperature on output?
  - How does heat stress mediate the temperature-output relationship

- **Setting:** India's manufacturing sector (cloth weaving, sewing, steel)

- **Data:** High-frequency (15 year panel, daily)
  - Worker-level, sewing line level, team level, plant-year, district-year

- **Design:** panel fixed effects

- **Result:** 1) heat $\downarrow$ plant-level annual output, 2) driven by output elasticity of labor, not capital 3) AC used in labor-intensive tasks (pareto efficient)

## Background

- Workers should respond quickly when working in uncomfy temperatures
  - Heat impacts on labor can be identified from daily or aggregated data
  - Distinguishes heat stress from alternative mechanisms which materialized over long time scales (i.e. conflict)
\vfill
- Two channels through which temperatures affect workers
  - Produce less (productivity)
  - Absenteeism
\vfill
- Can we expect lumpy investments in climate control in developing countries?
  - Depends on value-added per worker
  - If too low, then electricity cost not worth it

## Data: Worker + Cloth Weaving

- Worker data
  - Surveyed output and attendance from cloth-weaving, garment sewing, and steel firms
  - Three cloth-weaving factories in Surat, 
  - Eight garment plants (Delhi, Hyderabad, Chhindwara)
  - Steel plant (one large plant in Central India)
\vfill
- Cloth data (2012): 147 workers
  - Daily meters woven (for each worker)
  - Worker pay slips each day (output + attendance)

## Data: Garment Sewing
- Garment sewing (2012-2013): Line level
  - Lines of 10-20 workers
  - Productivity: hourly output relative to target
  - Daily production from 103 sewing lines
\vfill
- Climate control
  - Phased rollout of cooling equipment
  - 84 of 103 sewing lines in Delhi, of which 74 had cooling

## Data: Rail and Steel Plant
- Supplies rectangular blocks of steel to Indian Railways
- Three shifts per day
- Workers assigned to one of three teams, which rotate across shifts
- Data:
  - Team and number of blocks for each shift during 1999-2008
  - 9,172 shifts over 3,337 working days
  - Team-level absences over 857 days for 2000-2003

## Garment factory and cloth weaving
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig7.png",
                        auto_pdf = TRUE
                        )

```

## Rail/Steel Mill
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig8.png",
                        auto_pdf = TRUE
                        )

```

## Data: Panel
- Annual Survey of Industries (1999-2013)
  - Census of large manufacturing plants ($>$ 100 workers)
  - Random sample of small plants
  - Output, fixed assets, input expenses, employment
  - Unbalanced panel of 58,377 plants
\vfill
- District Manufacturing GDP (1998-2009)
  - Planning Commission of India
  - ASI plants + unregistered plants
\vfill
- Weather: public weather stations + gridded data (IMD)
  - Daily maximum at closest station to factory
  - Gridded data for panel data

## Data: Diamond Firms
- Survey 150 diamond cutting plants in Surat
  - Random sample from 500 registered units
- Five operations: 1) grading 2) marking, 3) rounding, 4) cutting, 5) polish
- Ask about number of workers and use of AC in each operation
- Ask to rate on 1-5 scale importance of each operation to quality of final output
  - Why?

## Empirical Strategy
Temperature-output estimated with weekly data as follows:

$$y_{iw} = \alpha_{i} + \gamma_{M} + \gamma_{t} + \sum_{j=2}^{J} \beta_{j}T_{iw}^{j} + \theta R_{iw} + \lambda X_{iw} + \epsilon_{iw} $$

- $y_{iw}$ is physical output by worker $i$ in week $w$
- If absent, output=0 for that day before aggregating to week
- $\alpha_{i}$ is individual FE (what does it account for?)
- $\gamma_{M/t}$ are month/year FEs (what does it account for?)
- $T_{j}$ counts days falling into temperature bin $j$ (captures nonlinearities)

## Results: Garment Sewing
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig1.png",
                        auto_pdf = TRUE
                        )

```
- output is lower in hotter weeks **without climate control**
- Moving from lowest to highest temperature bin $\downarrow$ productivity by 8\%
- Note: these are 90\% CI's!

## Results: Cloth and Steel
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Lagged Effects
- Previously, $\beta_{j}$ combined productivity and absenteeism channel
- Now we use daily data
$$y_{id} = \alpha_{i} + \gamma_{M} + \gamma_{t} + \sum_{j} \beta_{j}T_{id}^{j} + \sum_{j} \omega_{j}L_{id}^{j} + \theta R_{id} + \lambda X_{id} + \epsilon_{id} $$ 
- where $T^{j}$ is now an indicator for day $d$ falling into temperature bin $j$
- $X_{id}$ includes day-of-week FE
- $L_{id}^{j}$ counts number of days in bin $j$ during six preceding days

## Discussion

- Is lag specification plausible?
  - Is six days a plausible window to detect cumulative exposure?

- What are the identification assumptions?

- What are the threats?

## Lagged Effects: Productivity
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Lower output on hot days in sites w/o climate control
- Weavers: additional day $>35\degree$ in past week $\downarrow$ daily output by 3\%

## Lagged Effects: Absenteeism
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig4.png",
                        auto_pdf = TRUE
                        )

```
- How do you interpret these results?

## 15-year plant-level panel results
$$y_{it} = \alpha_{i} + \gamma_{t} + \sum_{j=2}^{5}\beta_{j}T_{it}^{j} + \theta R_{it} + \epsilon_{it} $$
outcome is log value of plant $i$ output in year $t$

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "95%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Recap of Results

- Daily worker data: workers are less productive on hot days
\vfill
- Daily worker data: (kind of) more absenteeism on hot days
\vfill
- Plant panel: **aggregate** output is lower on hot days
  - This is the object we are interested in
  - Remember $\frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}$
\vfill
- **How much of the temperature-induced $\Delta Y$ is due to $\Delta MP_{L}$?**

## Lets make a Cobb-Douglas model

Consider a production function:
$$ Y = A \cdot K^{\omega} \cdot L^{\beta} $$
Now modify by letting elasticities be a function of temperature, $T$
$$ Y = A(T) \cdot K^{\omega(T)} \cdot L^{\beta(T)} $$
where:
\begin{align*}
A(T) &= A_{0} + \sum a_{j}T_{j} \\
\omega(T) &= \omega_{0} + \sum \omega_{j}T_{j} \\
\beta(T) &= \beta_{0} + \sum \beta_{j}T_{j}
\end{align*}

## Estimating Equation

- Log-linearizing, we get Equation (7) from the paper:

$$ y = \alpha_{0} + \sum_{j=2}^{5} \alpha_{j}T^{j} + \omega_{0} \cdot k + \sum_{j=2}^{5} \omega_{j}T^{j}k + \beta_{0} \cdot l + \sum_{j=2}^{5} \beta_{j}T^{j}l $$

- Data:
  - k: opening value of equipment and machinery at start of financial year
  - l: number of full-time workers

- **Results**
  - temperature effects on output elasticitiy of L is negative and significant
  - temperature effects on output elasticity of K are (weakly) positive
  - Controlling for interactions, residual temperature-output effect is zero
  - Since overall temperature-output relationship is negative, temperature-induced declines in labor productivity drive the effect

## Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "55%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/anant2021_fig6.png",
                        auto_pdf = TRUE
                        )

```

<!-- ## Adaptation -->
<!-- - Foregone output from high temperature may encourage adaptation -->
<!-- \vfill -->
<!-- - Short term decision: compare cooling costs with output loss -->
<!-- \vfill -->
<!-- - Long term decision: labor-capital substitution or move location -->
<!-- \vfill -->
<!-- - Efficient allocation: if labor productivity drives temperature-output relation, then labor-intensive processes with high value added should be protected -->
<!-- \vfill -->
<!-- - Diamond processes: 1) sorting/grading; 2) marking; 3) rounding; 4) cutting; 5) polish -->
<!--   - Measure labor intensity (share of workers engaged in each process) -->
<!--   - Mechanization (share of total machines used in each process) -->
<!--   - Importance rating (scale 1-5, from survey) -->
<!-- \vfill -->
<!-- - Estimate probability of AC with logit model  -->

## Discussion
- What do you think about the paper
- Critique
- How can they improve the identification?
- This data is as good as it gets: did they do everything they could?
  - What about long term effects/adaptation?
- Alternative explanations

## Back to the model
\begin{align}
MWTP_{e} = \frac{\lambda_{e}}{\lambda_{y}} 
 = \frac{1}{\lambda_{y}} \left(\frac{\partial u}{\partial e} + \frac{\partial u}{\partial h}\frac{\partial h}{\partial e}\right) + \frac{\partial \Delta y}{\partial e} + \frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e} \notag
\end{align}

- Most research in environment/development economics seeks to identify $\frac{\partial h}{\partial e}$
  - At this point, $\frac{\partial h}{\partial e}>0$ is unambiguous
  - Focus is on **magnitudes**

\vfill
- What about indirect effects via productivity ($\frac{\partial \Delta y}{\partial h}\frac{\partial h}{\partial e}$)? 
  - This is the research frontier
\vfill
- Next: what about $\frac{\partial \Delta y}{\partial e}$, the direct effect of environment on incomes
  - Almost no research on this

## How might the environment **directly** affect incomes?

- Assume landholders vary in distance to a central market
- Each chooses land and labor to maximize farm profits
- New PA imposed that overlaps existing productive land and restricts land use
- Assuming initially efficient allocation, what happens to net welfare?
- What if land rents increase outside the PA?

## Background: Conservation and Development

- On other hand, PAs may generate tourism income and other positive spillovers

- Surprisingly little evidence on income channel

- Main challenge: non-random placement of PAs

## Sims (2010): Conservation and Development: Evidence from Thailand

**Research Question:** How does PA policy affect local poverty outcomes? 
  - What are the mechanisms?

**Setting:** Thailand in 2000

**Data:** Household consumption surveys + PA locations

**Empirical Strategy:** Instrument PA location with priority watershed status

**Results:** PAs increase consumption and lower poverty
  - Higher tourism is the main mechanism

## Data

- Wildlife sanctuaries (N=31) and national parks (N=57)
  - Source: World Database on Protected Areas
  - Includes year of establishment
\vfill
- Poverty Mapping
  - Model household income in 2000 Thai Economic Survey as a function of household characteristics and assets
  - Use coefficient to predict household income in census households
\vfill
- Poverty Measure
$$ FGT(\alpha) = \frac{1}{n}\sum_{y_{i}<p}\left( \frac{p-y_{i}}{p} \right)^{\alpha} $$
where $p$ is poverty line and $y_{i}$ is income

## Empirical Strategy: Ordinary Least Squares

$$y_{ij} = \beta_{1} PA_{ij} + \beta_{2}G_{ij} + \beta_{3}X_{ij} + \alpha_{j} + \epsilon_{j}$$

- $y_{ij}$ is socioeconomic outcome for locality $i$ in district $j$
- $PA_{ij}$ is share of locality land designated as PA prior to 2000
- $G_{ij}$ are fixed geographic controls (slope, elevation, water, etc.)
- $X_{ij}$ is a measure of pre-treatment (historic forest cover, distance to major roads)
- $\alpha_{i}$ are district fixed effects
\vfill

**What variation is exploited?**

**What are the identification assumptions?**

**What are potential omitted characteristics?**

<!-- talk about spatial spillovers -->

## OLS Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Median land share protect is 1/3
- Expected $\uparrow$ in consumption is about 4.5\% (divide by coeff. by 3)

## Instrumental Variables Solution

- Instrument PA placement with **priority watershed status**
\vfill
- Upper watersheds (headwaters) protected to improve **downstream** water quality
  - Watershed protected not driven by local (headwater) economic needs
\vfill
- First Stage:
 $$ PA_{ij} = \pi_{1}Z_{ij} + \pi_{2}G_{ij} + \pi_{3}X_{ij} + \pi_{4}W_{ij} + \delta_{j} + u_{ij} $$

- $Z_{ij}$ includes a) $i$ overlaps major tributary, b) distance to nearest major tributary .
\vfill
- $\pi_{1}$ picks up variation in PA driven by differences in downstream characteristics

## Watershed Instrument
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Discussion
\centering
Is the IV plausible? Why or why not?

## IV Results
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Signs agree with OLS, but magnitudes are more than 2x as large
- Consistent with idea that OLS negatively biased

## Mechanisms: Is tourism responsible?

- PAs restricts land use, so why does income increase?
\vfill
- Test 1: Test whether impacts in national parks $>$ in wildlife sanctuaries
  - Idea: NP's encourage tourism whereas sanctuaries only permit ecotourism/research
\vfill
- Test 2: Back-of envelope calculation
  - if we know costs/benefits of tourism, is it enough to explain net $\uparrow$ in consumption?
\vfill
- Test 3: Are benefits higher where tourism is higher
  - Park statistics show that visitation is highest near major cities

## Test 1: National parks vs. Wildlife Sanctuaries
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Effects larger where tourism is allowed (NPs)
- Difference in coefficients not statistically significant

## Test 2: Back of the envelope calculation

- Assume income is agricultural and reduced proportionally to land lost
- Assume credit constraints (no savings) $\rightarrow$ income losses = consumption losses
- Consumption cost of PAs is 11\% of income (forest cover coeff. - see paper)
- To achieve 5\% *net* $\uparrow$ in consumption, tourism must generate 16\% consumption
  - Equivalent to 233 Baht per month (0.16 $\times$ mean consumption)
- 10.81 mil. tourists in 2020 and 172,978 hh in PA localities = 61 tourists/hh
- Each tourist would need to contribute: (233*12)/61 = 46 Baht (\$USD 1.20) of net income to generate observed increases in consumption

**Is this reasonable?**

## Test 3: Heterogeneity by Distance to City
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig5a.png",
                        auto_pdf = TRUE
                        )

```
- Most tourists near city, but effects concentrated at intermediate distance
- Why?
<!-- Close cities, opportunity cost of agricultural production is higher (urban wage, because migration cost is low). Tourism income may not be enough to compensate for high opportunity cost. At far distances, opportunity costs are lower (wage divided by migration cost), but so are tourism prospects. Sweet spot is in the middle. This is just an ex-post rationalization and would make a good research idea. -->

## Test 3: Heterogeneity by Distance to City
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/sims2010_fig5b.png",
                        auto_pdf = TRUE
                        )

```

## Discussion
- What do you think of the paper?
- Do you believe the instrument?
- What are some alternative explanations?

### 2024-08-29 lectures/week3/inclass.R

@@ -1,58 +1,0 @@
-# Purpose: This script is for week 3, class 2
-# Author: Josh Merfeld
-# Date: 2023-10-05
-################
-
-# libraries
-library(tidyverse)
-library(haven)
-library(fixest)
-
-# read in data
-df <- read_dta("shrugdata.dta")
-
-df <- df %>%
-        mutate(prop_female = pc11_pca_tot_f/pc11_pca_tot_p,
-                prop_male = pc11_pca_tot_m/pc11_pca_tot_p,
-                sexratio_f_to_m = pc11_pca_tot_f/pc11_pca_tot_m,
-                state = substr(df$shrid2, 4, 5),
-                district = substr(df$shrid2, 7, 9))
-
-
-reg1 <- feols(sexratio_f_to_m ~ log(pc11_pca_tot_p), data = df)
-reg2 <- feols(sexratio_f_to_m ~ log(pc11_pca_tot_p), data = df, vcov = "HC1")
-reg3 <- feols(sexratio_f_to_m ~ log(pc11_pca_tot_p) + state, data = df, vcov = "HC1")
-reg4 <- feols(sexratio_f_to_m ~ log(pc11_pca_tot_p) | state, data = df, vcov = "HC1")
-
-
-
-table <- etable(reg1, reg2, reg3, reg4)
-
-
-
-feols(pc11_pca_p_lit ~ pc11_pca_tot_m + pc11_pca_tot_f, data = df, vcov = "HC1")
-
-
-
-
-
-
-ggplot() + 
-    geom_point(data = df, aes(x = log(pc11_pca_tot_p), y = sexratio_f_to_m)) +
-    # add regression line
-    geom_smooth(data = df, aes(x = log(pc11_pca_tot_p), y = sexratio_f_to_m), method = "lm", se = FALSE) +
-    theme_minimal()
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2024-08-29 lectures/week10/inclass.R
update slides

@@ -1,51 +1,0 @@
-library(tidyverse)
-library(readxl)
-library(fixest)
-
-#######
-# load the xlsx file
-#######
-df <- read_excel("wdi.xlsx")
-colnames(df)[1:3] <- c("country", "iso", "variable")
-
-tempnames <- names(table(df$variable))
-
-df$variable[df$variable==tempnames[1]] <- "empag"
-df$variable[df$variable==tempnames[2]] <- "empind"
-df$variable[df$variable==tempnames[3]] <- "empservices"
-df$variable[df$variable==tempnames[4]] <- "emptopopratio"
-df$variable[df$variable==tempnames[5]] <- "gdppc"
-df$variable[df$variable==tempnames[6]] <- "gdppcgrowth"
-
-
-df <- pivot_longer(data = df, cols = -c(1:3))
-
-tempnames <- names(table(df$variable))
-
-tempdf <- df %>% filter(variable==tempnames[1]) %>% select(-c("variable"))
-colnames(tempdf)[ncol(tempdf)] <- tempnames[1]
-for (i in 2:length(tempnames)){
-    tempdf <- cbind(tempdf, (df %>% filter(variable==tempnames[i]))[,ncol(df)])
-    colnames(tempdf)[ncol(tempdf)] <- tempnames[i]
-}
-
-tempdf[tempdf==".."] <- NA
-
-for (i in tempnames){
-    tempdf[i] <- as.numeric(as_vector(tempdf[i]))
-}
-
-
-summary(tempdf)
-
-tempdf$name <- substr(tempdf$name, 1, 4)
-tempdf$name <- as.numeric(tempdf$name)
-tempdf <- tempdf %>% rename(year = name)
-
-
-
-mean(tempdf$empag, na.rm = TRUE)
-apply(tempdf[tempnames], 2, mean, na.rm = TRUE)
-apply(tempdf[tempnames], 2, sd, na.rm = TRUE)
-
-

### 2024-08-29 lectures/week4/classcode.R

@@ -1,49 +1,0 @@
-# code for class, 2023-10-12
-
-library(tidyverse)
-library(fixest)
-library(haven)
-
-
-df <- read_dta("deserrannoetal.dta")
-df <- df %>% filter(is.na(vote)==F)
-
-groupstats <- df %>%
-                group_by(group_unique) %>%
-                summarize(groupsize = n(),
-                            vote = mean(vote))
-groupstats <- groupstats %>%
-                group_by(vote) %>%
-                summarize(sizemean = mean(groupsize),
-                            sizesd = sd(groupsize))
-groupstats
-
-indstats <- df %>%
-            group_by(vote) %>%
-            summarize(agemean = mean(age, na.rm = T),
-                        agesd = sd(age, na.rm = T))
-indstats
-
-statsall <- cbind(indstats, groupstats[,2:3])
-statsall
-
-
-
-# regressions
-# does wealth of COMMITTEE MEMBERS differ by treatment status
-feols(wealth ~ vote | branch, 
-        cluster = "group_unique", 
-        data = df %>% filter(CM==1))
-feols(primary_school ~ vote | branch, 
-        cluster = "group_unique", 
-        data = df %>% filter(CM==1))
-
-
-
-
-
-
-
-
-
-

### 2024-08-29 lectures/week7and8/inclass.R

@@ -1,60 +1,0 @@
-# In-class code from week 6
-
-library(tidyverse)
-library(fixest)
-library(did2s)
-
-# Load data
-df <- readRDS("weeks/week5and6/ateretal.rds")
-colnames(df)
-df <- zap_labels(df)
-
-# Plot
-sums <- df %>%
-        group_by(date) %>%
-        summarize(adoption = mean(public_defender))
-ggplot(sums) + 
-    geom_line(aes(x = date, y = adoption)) +
-    xlab("Date") + ylab("Proportion adoption") + theme_minimal()
-
-
-# TWFE
-(reg1 <- feols(log_num_arrests ~ public_defender | district + date,
-                data = df,
-                cluster = "district"))
-
-
-# Wild cluster bootstrap
-library(fwildclusterboot)
-(reg1 <- feols(log_num_arrests ~ public_defender | district + week_year,
-                data = df,
-                cluster = "district"))
-boot_reg <- boottest(
-                    reg1, 
-                    clustid = c("district"), 
-                    param = "public_defender", 
-                    B = 9999,
-                    type = "webb"
-                    )
-boot_reg
-
-
-
-
-# DID2S
-(did2s <- did2s(data = df, yname = "log_num_arrests", first_stage = "1 | district + date", 
-                second_stage = "public_defender", treatment = "public_defender", cluster_var = "district"))
-
-
-
-
-# event study
-df <- df %>%
-        mutate(first_date = ifelse(public_defender==1, week_year, NA)) %>%
-        group_by(district) %>%
-        mutate(first_date = min(first_date, na.rm = TRUE),
-                event_weeks = week_year - first_date) %>%
-        ungroup()
-(did2s2 <- did2s(data = df, yname = "log_num_arrests", first_stage = "1 | district + date", 
-                second_stage = "i(event_weeks, ref = 0)", treatment = "public_defender", cluster_var = "district"))
-

### 2024-08-29 lectures/week7and8/week9practice.R

@@ -1,37 +1,0 @@
-# in class, week 9
-
-library(tidyverse)
-library(fixest)
-
-df <- read_csv("pollution.csv")
-
-head(df)
-
-# first stage: pm25 = wind
-# second stage: yield = pm25hat
-
-(sumstats <- df %>%
-            group_by(year) %>%
-            summarize(mean_pm25 = mean(pm25, na.rm = TRUE),
-                        mean_yield = mean(log(yield), na.rm = TRUE),
-                        mean_rain = mean(rain_z, na.rm = TRUE),
-                        mean_temp = mean(temp_mean, na.rm = TRUE),
-                        mean_wind = mean(wind, na.rm = TRUE)))
-colnames(sumstats) <- c("Year", "PM 2.5.", "Yield (log)", "Rain (z)", "Temp", "Wind")
-
-
-# reduced form
-# log(yield) = rain + temp + wind, add shrid + year FE
-feols(log(yield) ~ rain_z + temp_mean + wind | shrid + year, data = df, cluster = "shrid")
-
-feols(pm25 ~ rain_z + temp_mean + wind | shrid + year, data = df, cluster = "shrid")
-
-# 2SLS
-reg1 <- feols(log(yield) ~ rain_z + temp_mean | shrid + year | pm25 ~ wind,
-                data = df, cluster = "shrid")
-
-etable(reg1,
-        stage = 2,
-        fitstat = )
-
-

### 2024-08-30 assignments/research_proposal/research_proposal.Rmd
update first week

@@
 ---
+
+**Due date: Final draft due Oct. 31st, 2024**
 
@@
 \end{table}
+Note: the first draft is pass/fail. I will check it for completion and award a pass. If you fail to submit a first draft, then five percentage points will be taken off of your final grade.
 
@@
 - Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}
-- Read papers!
 

### 2024-08-30 lectures/module1/module1.Rmd
update first week

--- 2024-08-29 lectures/week1/week1.Rmd
@@
     | Environmental and Development Economics
-    | Week 1 - Introduction
+    | Module 1 - Introduction
 author:
@@
 header-includes:
-  - \AtBeginDocument{\title[Week 1 - Introduction]{\textbf{Environmental and Development Economics} \\ Week 1 - Introduction}}
+  - \AtBeginDocument{\title[Module 1 - Introduction]{\textbf{Environmental and Development Economics} \\ Module 1 - Introduction}}
   - \input{header.tex}
@@
 
+## Housekeeping
+
+- **Class Time/Location:** Tues/Thurs 11:45am-1:25pm, Ruttan 119
+\vfill
+- **Office Hours:** Thursdays 1:30pm-2:30pm, Ruttan 337D
+\vfill
+- **Course website:** https://github.com/rmadhok/enviro-dev-grad
+  - lectures, assignments, syllabus
+  - Try to skim reading(s) beforehand
+\vfill
+- **Assignments**: Upload through Canvas
+\vfill
+
 ## Today
@@
 \vfill
-- Course website: https://github.com/rmadhok/enviro-dev-grad
-  - lectures, assignments, syllabus
-
-\vfill
+- If time: conceptual framework for environment & development econ
 
@@
 
-Instead, I am organizing around **\textcolor{orange}{FOUR}** key questions:
+Instead, I am organizing around **\textcolor{orange}{FIVE}** key questions:
 
-1. Why is environmental quality so bad in developing countries?
+1. How does economic development affect the environment, and vice versa?
 
-2. What are the costs of poor environmental quality in developing countries?
+2. Why is environmental quality so bad in developing countries?
 
-3. Why is WTP for environmental quality low in developing countries?
+3. What are the costs of poor environmental quality in developing countries?
 
-4. What are the political economy barriers to environmental protection?
+4. Why is WTP for environmental quality low in developing countries?
+
+5. What are the political economy barriers to environmental protection?
 
@@
   - research design, identification strategy, estimation techniques
+  - aim for two per class (please skim beforehand)
 
-- Emphasize recent JMPs
+- Emphasize recent papers
 
@@
 
-## Weeks 1 and 2
+## Course Outline
 
-**Week 1: Theory of Environment/Development**
+**Module 1: Introduction**
 
-- How to use theory to ask the right questions (lecture 2)
+- Lecture 1: Course intro + how to use theory to ask the right questions
 
@@
 
-**Week 2: The effect of development on the environment**
+**Module 2: The effect of development on the environment**
 
-- air quality, water quality (lecture 3)
-- forests, biodiversity (lecture 4)
+- Lecture 2: Income effects
+- Lecture 3: Access to capital (technology and infrastructure)
 
-## Weeks 3 and 4
+## Course Outline
 
-**Week 3: The effect of the environment on development**
+**Module 3: The effect of environment on development**
 
-- human capital (lecture 5)
-- firms (lecture 6)
-
+- Lecture 4: Health
+- Lecture 5: Productivity
 \vfill
 
-**Week 4: WTP for Environmental Quality**
+**Module 4: Why is WTP low in developing countries?**
 
-- TBD (lecture 7)
-- TBD (lecture 8)
+- Lecture 6: Revealed preference approaches
+- Lecture 7: Incentive compatible approaches
 
-## Weeks 5, 6, and 7
+## Course Outline
 
-**Week 5: Environmental policy design**
+**Module 5: Environmental Policy Design**
 
-- Spillovers (lecture 9)
-- Pollution markets in developing countries (lecture 10)
-
+- Lecture 8: Monitoring, enforcement
+- Lecture 9: Barriers to optimal design
 \vfill
 
-**Week 6: Political economy of the environment**
+**Module 6: Political Economy of the Environment**
 
-- Corruption and weak enforcement  (lecture 11)
-- Environmental justice (lecture 12)
+- Lecture 10: Electoral cycles, corruption
+- Lecture 11: State Capacity
 
-\vfill
+## Course Outline
 
-**Week 7: Research presentations**
+**Module 7: Research Proposal Presentations**
 
-- I will provide more details throughout the semester
+- Lecture 12: TBD
+- Lecture 13: Presentations
+- Lecture 14: Presentations
 
@@
 In-class presentations      & 10\% &  &  &  \\
-Problem Set           & 20\% &  &  &  \\
+Replication Assignment           & 20\% &  &  &  \\
 Research Proposal & 60 \% &  &  &  \\
@@
 
-- At start of **first** lecture each week, you'll give a 15 min paper presentation
+- I want you to become expert conference presenters after taking this class
 
-- Each student submits **seven** summary slides (5% of grade)
+- At start of **each** class, you'll give a 10 min paper presentation
+  - The paper for presentation is on the syllabus
+
+- Each student submits **10** summary slides (5% of grade)
   - motivation, research question, methods, results
@@
   - You will also **extend** the results
+  - Many papers on the syllabus have replication files
+
+- You will submit a write-up explaining what you did
 
@@
 \begin{tabular}{lllll}
-Written Proposal      & 30\% &  &  &  \\
-Peer Review           & 20\% &  &  &  \\
-Proposal Presentation & 10\% &  &  &  \\
+Written Proposal      & 30\% & Oct. 31  &  &  \\
+First Draft      & pass/fail \% & Oct. 3rd  &  &  \\
+Peer Review           & 20\% & Oct. 10th  &  &  \\
+Proposal Presentation & 10\% & Oct 15/17  &  &  \\
                       &      &  &  & 
@@
   - You are NOT expected to actually do the analysis  
-  - I will provide small deadlines (outline, first draft, etc.) along the way
+  - I will provide small deadlines (first draft, etc.) along the way
+
+- Come to office hours to pitch your idea
 
@@
 
-## Next week
+## Next Class
 
@@
 
-- Impact of development on the environment (air, water)
+- Impact of economic development on the environment (income effects)
 
 - Impact of development on the environment (forests, biodiversity)
+
+##
+
+\huge
+Aside: Best Practice for Short Presentations
+
+## Best Practices: Structure
+
+- You cannot present a paper in 10 minutes
+  - Do not give detailed lit review or go into extreme detail
+
+\vfill
+- Instead, you are giving a trailer for the movie
+  - Convince the audience that they should read the paper
+
+\vfill
+- Your goal is only to state why the paper is important, and what you did
+\vfill
+
+- Convey paper's importance in first slide
+
+## Best Practices: Slides
+
+- Motivation + broad research question (1 slide)
+
+- Full paper overview (1 slide)
+
+- Lit review (optional, 1 slide)
+
+- Background (1 slide)
+
+- Data (2 slides)
+
+- Empirical Strategy (2 slides)
+
+- Results (2 slides)
+
+- Summary (1 slide)
+
+## Example of Preview Slide
+
+- **Question:** How do firms react to tribal forest policy?
+
+- **Idea:** Model aggregate economic response and changes in firm composition
+
+- **Setting:** India Forest Rights Act (2008)
+  - Imposes transaction cost on firms
+
+- **Data:** Manufacturing census (2001-2015); Deforestation permits (2001-2021)
+
+- **Empirical Strategy:** Diff-in-diff using policy shift in tribal and non-tribal district
+
+\vfill
+
+### Results Preview
+1) decline in firm activity, 2) less forest encroachment by industry\
+3) larger, but less productive firms survive
+
+## Best Practices: Slide Format
+
+- 1 minute per slide
+
+- Avoid chart junk
+
+- One line bullets
+  - No need for full sentences
+
+- Vary text slides and text + image slides
+
+- Don't put too many equations
+  - Save details for speaking, or talk ``about'' equation
+
+- Summarize findings again at the end
+
+## Best Practices: Presentation
+
+- Speak clearly and loudly
+
+- Speak slowly
+
+- Look at audience; Do not show your back
+
+- Do not stand in front of slides
+
+- Avoid jargon
+
+- Avoid pacing around room
+
+- Stick to your time limit

### 2024-08-30 lectures/module2/module2.Rmd
update first week

--- 2024-08-29 lectures/week2/week2.Rmd
@@
 header-includes:
-- "\\AtBeginDocument{\\title[Week 2 - development and environment]{Environmental and Development
-  Economics \\\\ Week 2 - How does development affect the environment?}}"
+- "\\AtBeginDocument{\\title[Module 2 - development and environment]{Environmental and Development
+  Economics \\\\ Module 2 - How does development affect the environment?}}"
 - \input{header.tex}

### 2024-09-02 lectures/module4/module4.Rmd
before first lecture

---
title: "| Environmental and Development Economics\n| Module 4 - WTP for Environmental Quality
  \n"
author: |
  | Raahil Madhok
  | UMN Apllied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 4 - WTP for environment]{Environmental and Development
  Economics \\\\ Module 4 - WTP for Environmental Quality in LMICs}}"
- \input{header.tex}
- \usepackage{gensymb}
---

## 
\centering
\Huge Lecture 6

\large WTP for Environmental Quality in Developing Countries

## Housekeeping

- State things that are due.

## Today

- **Guiding question:** what is the WTP for environmental quality in LMICs?
\vfill
- Today's focus: How do we even measure this?
  - Behavioral models
  - Revealed preference approaches

\vfill
- Next time: why is WTP so low in developing countries?

## Measuring Willingness To Pay

- Slides on Logit and Mixed Logit

## 
\huge How is this implemented in practice?

## Kremer et al. (2011): WTP for Clean Water in Kenya

- **Research Question:** What is household WTP for clean water in LICs?
  - What are health impacts of water quality improvements? (we will skip this)
\vfill
- **Setting:** 1,354 households in Kenya
\vfill
- **Intervention:** Improve water quality at source for 184 natural springs
\vfill
- **Empirical Strategy:** RCT + Revealed preference valuation
\vfill
- **Results:** Clean water valued at \$USD 2.96 per hh (\$800 VSL)

## Spring Protection

- Spring water is crucial drinking water source in Africa
  - But vulnerable to contamination when people dip vessels
  - Runoff introduces animal/human waste
\vfill
- Spring protection common in Africa to improve water quality at source
\vfill
- Seal off spring at source by encasing in concrete
  - Water flows out of pipe rather than ground, avoiding contamination
\vfill
- Nothing can ``break'' $\rightarrow$ less maintenance than wells/pumps 
\vfill
- How much are hh WTP for spring protection?

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "68%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "75%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/spring_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Spring Protection
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center", fig.cap="Source: USAID"}

knitr::include_graphics(
                        path = "assets/spring_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Study Sample
- Sample frame: list of all unprotected springs in Busia and Butere-Mumias districts
\vfill
- Treatment: 184 randomly selected for protection
  - 1,345 households selected that use springs
\vfill
- Phased design: all springs eventually protected
  - Round I (Jan-Apr 2005); 47 springs, 350 households
  - Round II (Apr-Nov 2005); 46 springs; 349 households
  - Round III (2005-2006); 91 springs; 685 households
\vfill
- 7-8 households from each spring select from ``user list''

## Data Collection

- Water quality measured at each spring based on EPA protocols
  - Measure: E.Coli (log of \# bacteria per 100ml)
\vfill
- Household survey
  - Diarrhea
  - Anthopometrics
  - Water collection behvavior
  - Socioeconomic data
\vfill
- No difference in baseline water quality between treatment/control springs

## Water Quality and Health
- Impact of spring protection on water quality:
$$W_{jt'}^{SP} = \alpha_{t} + \phi_{1}T_{jt} + X_{j}^{SP'}\phi_{2} + (T_{jt} \cdot X_{j}^{SP})'\phi_{3} + \epsilon_{jt}$$
  - where $W_{jt}$ is water quality for spring $j$ at time $t\in {0,1,2,3}$
  - $T_{jt}$ is treatment status (=1 after protection)
\vfill
- Impact on health
$$Y_{ijt} = \alpha_{i} + \alpha_{t} + \phi_{1} T_{jt} + X_{ij}'\phi_{2} + (T_{jt} \cdot X_{ij})'\phi_{3} + u_{ij} + \epsilon_{ijt}$$

  -  where $Y_{ijt}$ is whether child $i$ had diarrhea in the past week

## Results: Impact of Spring Protection on Water Quality and Health

**Water Quality**

- Huge improvement in water quality: 66\% $\downarrow$ in E.coli
- No heterogeneity by baseline hygiene knowledge or education
  - Hypothesis: educated hh may maintain spring better
- Improvement in home water quality: 24\% $\downarrow$ in E.coli 
  - Why is the effect smaller?

\vfill
**Health**

- Diarrhea incidence $\downarrow$ by 4.7pp (20\% reduction compared to control)
- No impact on child weight
- Positive effect on BMI
\vfill

**What if control households choose to use treatment springs?**

## Revealed Preference: Valuing Clean Water in Kenya

- Let valuation of water from source $j$ be $Z_{j}$
  - reflects health and non-health attributes (e.g. ease of collection)
\vfill
- $\beta_{i}$: indirect utility from protection ($T_{jt}$) at $j$ at time $t$
\vfill
- $C_{i}>0$ is time cost of household $i$
\vfill
- $C_{i}D_{ij}$: Travel cost for additional trip to $j$ where $D_{i}$ is round-trip distance
\vfill
- Each trip affected by unobserved factors (weather, queue, errands, mood)

## Revealed Preference Approach
- Household indirect utility from one trip to $j$ at time $t$ is:
$$ u_{ijt} = \beta_{i}T_{jt}+Z_{j}-C_{i}D_{ij} + e_{ijt}$$
  - where $e_{ijt}$ is iid type I extreme value distributed
\vfill
- $i$ chooses $j$ over $k$ if: 
$$\beta_{i}(T_{jt}-T_{kt})+(Z_{j}-Z_{k})-C_{i}(D_{ij}-D_{ik})+(e_{ijt}-e_{ikt}) \geq 0$$
- Using CDF of EV Type I error, prob. of choosing $j$ over alternatives $h$:
$$P(y_{ijt}|X) = \frac{exp(X_{ijt}^{'}\beta)}{\sum_{h}exp(X_{iht}^{'}\beta)} \equiv \rho_{ijt}$$

## Mixed Logit
- Follow-up survey gives list of alternative water sources
- What is ratio of coeff. on $T_{jt}$ to coeff. on walking time?
  <!-- - Value of spring protection in terms of minutes spent walking -->

- Add preference heterogeneity with mixed logit:
$$P(y_{ijt}|X) = \int_{\beta} \rho_{ijt}f(\beta)d\beta$$
  - where $f(\cdot)$ is the mixing distribution
- Maximize log-likelihood to estimate mean and s.d. of these distributions

## WTP results
- Conditional logit estimates: how do we interpret these?
  - Coefficient on walking distance ($C_{i}D_{ij}$) = -0.055
  - Coefficient on spring protection ($\beta_{i}$) = 0.51

- Can $\beta_{i}$ entangle other amenity values?
  <!-- - reduced need to bend down, lower collection times -->
  <!-- - yes, but when include perceived water quality, $\beta_{i}=0$, meaning the greater value placed on protected springs is due to impact on water quality, not other factors -->

- Households with kids under 3 find walking more costly
  - Coeff. on interaction b/w $T_{jt}$ and child is positive (insignificant)
  - Households with kids value water quality more

## WTP for spring protection

- Ratio of mixed logit coefficients: 2.95/0.21

- Households value clean water by willing to walk 14 extra minutes

- Average hh takes 48 trips per week (see Descriptive Stats)
- Total extra minutes per year = 14 x 48 x 52 = 35,078 minutes
- Assuming 8-hour workday $\rightarrow$ **mean value of clean water is 73 workdays**
  - Note: in paper, it is 32.4 workdays!

## Comparison with Stated Preference Values

## Converting WTP to dollars

**Survey evidence**

- Subset of 104 respondents
- Contingent valuation: are you WTP \$5 to walk 1 less minute? \$10?, etc 
- Divide WTP (in terms of time) by money value of time
- Regress time value on household characteristics, and predict on full sample
- Mean value of time = \$USD 0.088 per 8-hour day (7\% of ag wages)

\vfill

**Wages**

- \$USD 1.26 for 8-hour day (Suri, 2009)
- Time cost = 25\% of average workday wage

## Health Valuation and VSL

## Discussion
<!-- - what do you think about approach? -->
<!-- - what do you think about time valuation? -->
<!-- - Why not include income directly in indirect utility fxn? -->

## Distribution of Household Values of Clean Water (Mixed Logit)
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/kremer2011_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Why is stated preference values larger? Why is CV distribution wider?

### 2024-10-01 assignments/research_proposal/research_proposal.Rmd
update proposal assignment
lecture 9

@@
 # Assignment Overview
-The goal of this project is to take you through a mock process of publishing in economics, including writing a paper (proposal, in this case) as well as the revise and resubmit process.
-You will: 1) draft a research proposal, 2) send your proposal to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, and 4) submit the final draft for ``publication'' (i.e.
-a final grade...).
+The goal of this project is to take you through a mock process of publishing in economics, including writing a paper (proposal, in this case), the revise and resubmit process, and presenting it.
+You will: 1) draft a research proposal, 2) send it to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, 4) present your proposal to the class, and 5) submit the final draft for "publication" (i.e.
+a final grade...) at the end of the semester.
 
-You **are not** allowed to ``double dip''. This means the idea must be original and you cannot use your second-year paper, or a proposal from another class.
+You **are not** allowed to ``double dip''. The idea must be original and you cannot use your second-year paper, or a proposal from another class.
 
@@
 
-Broadly, you will identify an original and specific research question, explain why it is important, and outline the research design that you will use to address it.
-After receiving an ``R\&R'' on the paper, you will improve it and submit a final draft at the end of the semester.
+Broadly, you will identify an original research question, explain why it is important, and outline the research design that you will use to address it.
+After receiving an ``R\&R'', you will improve it and submit a final draft.
 
@@
 Referee report    & October 10th     \\ \hline
+Presentations    & October 15/17th     \\ \hline
 Final draft    & October 31st     \\ \hline
@@
 \end{table}
-Note: the first draft is pass/fail. I will check it for completion and award a pass. If you fail to submit a first draft, then five percentage points will be taken off of your final grade.
+Note: the first draft is pass/fail. I will check it for completion and award a pass. If you fail to submit a first draft, then three percentage points will be taken off of your final grade.
 
@@
 \clearpage
-# Grading Rubric
+# Grading Rubric: Written Proposal
 
@@
 \end{table}
+
+\clearpage
+
+# Grading Rubric: Proposal Presentation
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|p{4cm}|p{8cm}|r|}
+\hline
+\textbf{Category} & \textbf{Description} & \textbf{Points} \\ \hline
+\textbf{Content} &  &  \\ \hline
+Introduction \& Research Question & Clear presentation of the research question, its importance, and its relevance to environmental and development economics. & /10 \\ \hline
+Theoretical Framework & Presents a coherent theoretical framework. & /15 \\ \hline
+Data \& Measurement & Explanation of the data sources, measurements, and units of analysis. Discusses the suitability of data for the research. & /10 \\ \hline
+Empirical Strategy & Explains the research design and methodology, including causality and identification strategies. & /15 \\ \hline
+\textbf{Delivery} &  &  \\ \hline
+Clarity \& Organization & Well-organized presentation, logical flow, and easy to follow. & 10 \\ \hline
+Speaking Style \& Engagement & Speaker presents confidently, engages the audience, and maintains good eye contact. & /10 \\ \hline
+\textbf{Visual Aids} &  &  \\ \hline
+Use of Slides/Visuals & Visual aids are clear, well-designed, and enhance the presentation. & /10 \\ \hline
+\textbf{Q\&A Performance} &  &  \\ \hline
+Understanding \& Response to Questions & Demonstrates understanding by responding thoughtfully and clearly to questions.
+Engages respectfully with questions, explaining complex concepts well & /10 \\ \hline
+\textbf{Time Management} &  &  \\ \hline
+Adheres to Time Limits & Presentation fits within the 20-minute timeframe without rushing. & /10 \\ \hline
+\textbf{Total} &  & /100  \\ \hline
+\end{tabular}
+\end{table}

### 2024-10-01 assignments/research_proposal/research_proposal.tex
update first week
update proposal assignment
lecture 9

@@
 
+\textbf{Due date: Final draft due Oct.~31st, 2024}
+
 \hypertarget{assignment-overview}{%
@@
 
-The goal of this project is to take you through a mock process of publishing in economics, including writing a paper (proposal, in this case) as well as the revise and resubmit process.
-You will: 1) draft a research proposal, 2) send your proposal to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, and 4) submit the final draft for ``publication'\,' (i.e.~a final grade\ldots).
+The goal of this project is to take you through a mock process of publishing in economics, including writing a paper (proposal, in this case), the revise and resubmit process, and presenting it.
+You will: 1) draft a research proposal, 2) send it to a classmate for peer review, and review one of their proposals, 3) integrate your reviewer's comments into your proposal, 4) present your proposal to the class, and 5) submit the final draft for ``publication'' (i.e.~a final grade\ldots) at the end of the semester.
 
-You \textbf{are not} allowed to ``double dip'\,'. This means the idea must be original and you cannot use your second-year paper, or a proposal from another class.
+You \textbf{are not} allowed to ``double dip'\,'. The idea must be original and you cannot use your second-year paper, or a proposal from another class.
 
@@
 
-Broadly, you will identify an original and specific research question, explain why it is important, and outline the research design that you will use to address it.
-After receiving an ``R\&R'\,' on the paper, you will improve it and submit a final draft at the end of the semester.
+Broadly, you will identify an original research question, explain why it is important, and outline the research design that you will use to address it.
+After receiving an ``R\&R'\,', you will improve it and submit a final draft.
 
@@
 \end{table}
+
+Note: the first draft is pass/fail. I will check it for completion and award a pass. If you fail to submit a first draft, then three percentage points will be taken off of your final grade.
 
@@
   Plamen Nikolov's \href{https://docs.iza.org/dp15057.pdf}{writing tips for economics research papers}
-\item
-  Read papers!
 \end{itemize}
@@
 
-\hypertarget{grading-rubric}{%
-\section{Grading Rubric}\label{grading-rubric}}
+\hypertarget{grading-rubric-written-proposal}{%
+\section{Grading Rubric: Written Proposal}\label{grading-rubric-written-proposal}}
 
@@
 
+\clearpage
+
+\hypertarget{grading-rubric-proposal-presentation}{%
+\section{Grading Rubric: Proposal Presentation}\label{grading-rubric-proposal-presentation}}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{|p{4cm}|p{8cm}|r|}
+\hline
+\textbf{Category} & \textbf{Description} & \textbf{Points} \\ \hline
+\textbf{Content} &  &  \\ \hline
+Introduction \& Research Question & Clear presentation of the research question, its importance, and its relevance to environmental and development economics. & /10 \\ \hline
+Theoretical Framework & Presents a coherent theoretical framework. & /15 \\ \hline
+Data \& Measurement & Explanation of the data sources, measurements, and units of analysis. Discusses the suitability of data for the research. & /10 \\ \hline
+Empirical Strategy & Explains the research design and methodology, including causality and identification strategies. & /15 \\ \hline
+\textbf{Delivery} &  &  \\ \hline
+Clarity \& Organization & Well-organized presentation, logical flow, and easy to follow. & 10 \\ \hline
+Speaking Style \& Engagement & Speaker presents confidently, engages the audience, and maintains good eye contact. & /10 \\ \hline
+\textbf{Visual Aids} &  &  \\ \hline
+Use of Slides/Visuals & Visual aids are clear, well-designed, and enhance the presentation. & /10 \\ \hline
+\textbf{Q\&A Performance} &  &  \\ \hline
+Understanding \& Response to Questions & Demonstrates understanding by responding thoughtfully and clearly to questions.
+Engages respectfully with questions, explaining complex concepts well & /10 \\ \hline
+\textbf{Time Management} &  &  \\ \hline
+Adheres to Time Limits & Presentation fits within the 20-minute timeframe without rushing. & /10 \\ \hline
+\textbf{Total} &  & /100  \\ \hline
+\end{tabular}
+\end{table}
+
 \end{document}

### 2024-10-08 assignments/research_proposal/peer_review.Rmd
lecture 11

@@
 
-- Clarity of the writing (grammar, organization, and structure),
-- Minor inconsistencies in the theoretical argument,
-- Specific technical clarifications needed in the research design or data analysis,
-- Suggestions for additional literature that could strengthen the theoretical framework,
-- Formatting issues or citation inaccuracies.
-- For each minor point, include:
+- Clarity of the writing (grammar, organization, and structure)
+- Minor inconsistencies in the theoretical argument
+- Specific technical clarifications needed in the research design or data analysis
+- Suggestions for additional literature that could strengthen the theoretical framework
+- Formatting issues or citation inaccuracies
 

### 2024-10-08 assignments/research_proposal/peer_review.tex
lecture 11

@@
 \item
-  Clarity of the writing (grammar, organization, and structure),
+  Clarity of the writing (grammar, organization, and structure)
 \item
-  Minor inconsistencies in the theoretical argument,
+  Minor inconsistencies in the theoretical argument
 \item
-  Specific technical clarifications needed in the research design or data analysis,
+  Specific technical clarifications needed in the research design or data analysis
 \item
-  Suggestions for additional literature that could strengthen the theoretical framework,
+  Suggestions for additional literature that could strengthen the theoretical framework
 \item
-  Formatting issues or citation inaccuracies.
-\item
-  For each minor point, include:
+  Formatting issues or citation inaccuracies
 \end{itemize}

### 2024-10-10 assignments/replication_exercise/replication_exercise.Rmd
lecture 12

@@
 - Some ideas for an extension are:
-    - **Enhance the research question**: this could mean studying heterogeneity or 
+    - **Enhance the research question**: this could mean studying heterogeneity or non-linearities.
     - **Robustness checks**: Verify the original results under different assumptions or specifications.

### 2024-11-11 lectures/module1/module1.Rmd
added .Rmd files

@@
 - **Class Time/Location:** Tues/Thurs 11:45am-1:25pm, Ruttan 119
+  - Last class: Oct 17th
+
 \vfill
@@
 - $n$ identical agents with utility from consumption, EQ, and health
-- Initial wealth $y_{0}$, health $h_{0}$, environmental equality $e_{0}$
+- Initial individual wealth $y_{0}$, health $h_{0}$, environmental equality $e_{0}$
 - health depends on self-protection, $s$, and $e$
-- Assume perfect markets (i.e. no externalities)
+- Assume perfect markets (i.e. no externalities) to benchmark first-best
 
@@
 
-## The Social Planner
+## The Fist Best Outcome
 
-- In first best, social planner sets MB = MC
-  - where $MC_{e} = \frac{\partial c_{e}}{\partial \Delta e}$ and $MC_{s} = \frac{\partial c_{s}}{\partial \Delta s}$
+- Agent sets ratio of marginal costs = ratio of marginal benefits
+$$\frac{MWTP_{e}}{MWTP_{s}}=\frac{\frac{\partial c_{e}}{\partial \Delta e}}{\frac{\partial c_{s}}{\partial \Delta s}}$$
+
 \vfill
-- But to aggregate over $n$, we must assume:
+- For social planner to aggregate over $n$, we must assume:
   - No preferences of her own
-  - No market failures
   - Can observe true MWTP
@@
 - Set the stage: 
-  - how does environment affect development ($\frac{\partial h}{\partial e}$) (week 2)
-  - how does development affect the environment (week 3)
+  - how does environment affect development ($\frac{\partial h}{\partial e}$) (module 2)
+  - how does development affect the environment (module 3)
 
@@
 - Goal: where can you make a contribution?
+  - E.g. Lots of work on $\frac{\partial h}{\partial e}$, but very little on $\frac{\partial \Delta y}{\partial e}$
 
@@
   - pollution (Duflo et al., 2013)
-  - deforestation (Burgess et al., 2012; Viera et al., 2024)
+  - deforestation (Burgess et al., 2012)
   - human-wildlife conflict (Madhok et al., 2024)
@@
 
-- This is partially a couse on development economics
+- This is partially a course on development economics
   - About market failures: land, labor, credit, etc.

### 2024-11-11 lectures/module2/module2.Rmd
added .Rmd files

--- 2024-08-29 lectures/week2/week2.Rmd
@@
 ---
-title: "| Environmental and Development Economics\n| Week 2 - How does development affect the environment?
+title: "| Environmental and Development Economics\n| Module 2 - How does development affect the environment?
   \n"
@@
   | Raahil Madhok
-  | UMN Apllied Economics
+  | UMN Applied Economics
 date: "`r Sys.Date()`"
@@
 
-- State things that are due.
+- Replication project (Ryan announcement)
+  - Due Oct 17th
+\vfill
+- Research proposal
+  - Start thinking about idea + come talk to me
+  - First draft: Oct 3rd (pass/fail)
+\vfill
+- Lecture slides will be posted before each class
+\vfill
+- Reading list is a work in progress
+\vfill
+- 5 min break mid-way through lectures
 
@@
 \vfill
-- Channel I: Income effects
-  - Changes in consumption
-  - Energy
+- Channel I: Income effects (today)
+  - **Changes in consumption**
+  - **Energy**
   - Diet
 \vfill
-- Channel II: Technology and Infrastructure
+- Channel II: Technology and Infrastructure (next week)
   - Agricultural productivity
   - Infrastructure
+
 \vfill
@@
 \vfill
+
+## GDP and Lead in Gasoline
+```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "80%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/ekc_lead.png",
+                        auto_pdf = TRUE
+                        )
+
+```
 
@@
 - What variation in "development" can approximate the experiment?
-  - what proxies development?
+  - what proxies development? (see \href{https://github.com/teamookla/ookla-open-data}{ookla speed test})
   - what does the proxy miss?
   - what (non-random) variation isolates causal relationship?
-  - what does that variation miss?
 
@@
 - Or, it may $\uparrow$ demand for conservation
-
-\vfill
-- Empirical challenge: household income is endogenous
 \vfill
@@
 - If we control for OVB, will panel fixed effects work? 
-  - No: small year-to-year income changes do not reflect ``development''
-  - Predictable short-term income fluctuations different than permanent change
+  <!-- - No: small year-to-year income changes do not reflect ``development'' -->
+  <!-- - Predictable short-term income fluctuations different than permanent change -->
 
@@
 
-- **Main variables:** village lat/lon, eligibility index, forest cover
+- **Main variables:** village lat/lon, eligibility index, forest cover (satellite)
+  - Enrollment by village until 2003
 \vfill
@@
 
-## Empirical Strategy
-
-- OLS:
-$$\Delta f_{i} = \alpha + \delta E_{i} + \beta'X_{i} + \epsilon_{i}$$
-where $E_{i}=1$ if index $I_{i} > -1.2$
-
-\vfill
+## Empirical Strategy: OLS Equation
+
+$$\Delta f_{i} = \alpha + \delta T_{i} + \beta'X_{i} + \epsilon_{i}$$
+
+  - where $\Delta f_{i}$ is \% deforestation in polygon i over 2000-2003
+  - $T_{i}=1$ if the locality enrolled by 2003
+  - $X_{i}$ includes poverty
+
+- But $T_{i}$ is not random!
+
+\vfill
+
+- If discontinuity is sharp, replace $T_{i}$ with $1[E_{i}>-1.2]$
 - RDD: Sample window $-2 < I < -0.2$
-\vfill
+
+## RDD Equation
+
+- Discontinuity is fuzzy, not sharp
+  - How do we know? Two reasons
+
+- Solution: Use cutoff to instrument treatment probability
+
 - Fuzzy RDD/IV:
@@
 - Approach 1: Use experimental sample (Progresa) with household survey data
+  - Progresa had randomized rollout at village level for first 3 yrs
+
 \vfill
@@
 - Before (1997-1998) and after (2000) data on consumption
-- 506 villages, 320 treated, 186 control
+- 506 villages, 320 treated, 186 control (not-yet-treated)
 - Treatment at the village level
@@
 - $y_{it}$ = consumption
-- cluster at locality level 
+- $Treat_{i}=1$ if household $i$ in treated locality
+- cluster at locality level bc randomization was at locality level
+- How do we interpret $\gamma_{3}$?
 \vfill
@@
 \begin{align*} 
-y_{it} & = \beta_{0} + \beta_{1}Treat_{i} + \beta_{2}Post_{t} + \beta_{3}Treat_{i} \times Post_{i} \\ & + \beta_{4}Road_{i} + \beta_{5} Road_{i} \times Treat_{i} + \beta_{6} Road_{i} \times Post_{t} \\ & + \beta_{7} Road_{i} \times Treat_{i} \times Post_{i} + \epsilon_{it}
+y_{it} & = \beta_{0} + \beta_{1}Treat_{i} + \beta_{2}Post_{t} + \beta_{3}(Treat_{i} \times Post_{t}) \\ & + \beta_{4}Road_{i} + \beta_{5} (Road_{i} \times Treat_{i}) + \beta_{6} (Road_{i} \times Post_{t}) \\ & + \beta_{7} (Road_{i} \times Treat_{i} \times Post_{t}) + \epsilon_{it}
 \end{align*}
@@
 \vfill
-Bigger impact where supply response is localized
+More deforestation in places with less market access (low road density)
 
@@
 ## Thoughts?
-- Form discussion group to think of alternate story
 
 - Do we believe the story about increased demand for land-intensive goods?
-  - Increased income relieves credit constraints
-  - Expands capital, increases ag productivity on extensive margin b/c malfunctioning factor markets.
-  - Market access may improve access to inputs, and reduce deforestation. 
-  - Consistent with papers results.
-
-- Only focused on one mechanism
-  - returns to off-farm labor
-  - income may increase aggregate production and lead to deforestation?
-  - market access may lead to migration, less population pressure and deforestation
-  - also consistent story
+  - Less market access means less access to inputs
+  - Underinvest in technology, leading to higher deforestation
+  - Maybe this is a story about technology, not consumption
+
+- Can you think of an alternate explanation?
+
+- Other concerns?
 
@@
   - we cannot answer why
-
-- But where is supply response coming from?
-  - Need much more details spatial analysis
-  - SLX model
-
-- Road density can be capturing other dimensions of underdevelopment, not only market access
-
-- How do we know that number of plots, and farm size increases are encroaching into forest? what if they displace other land types
-
-- Policy implication limited by thiessen polygon
-  - What is the natural unit for policy implementation?
 
@@
 
+##
+
+\centering
+\huge Take a 5 minute break
+
 ## Back to the question
@@
 - Energy ladder: as income $\uparrow$, move from solid fuel $\rightarrow$ gas $\rightarrow$ electricity
-  - Electricity may displace dirty energy (wood)
+  - Electricity may displace dirty energy (wood), which $\downarrow$ pollution
 
 - But, electricity also increases total energy (fridge, AC)
+  - ...which $\uparrow$ pollution
   - energy footprint increases with income
@@
 ## GDP and Car Ownership
-```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "45%", fig.align = "center"}
+```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "40%", fig.align = "center"}
 
@@
 ```
+Of course, these are correlations
+
+## But first, some theory
+
+- Imagine we can exogenously vary income to poor households
+\vfill
+- Will transfers to initially poor households yield **lumpy** investments?
+  - Movements up the ladder involve big purchases
+  - May explain why 1.3 billion people lack electricity (Gertler et. al, 2016) 
+\vfill
+- Credit constraints play an important role
+  - Without credit markets, buy now but have nothing in next period
+  - With savings, can save now, buy later (smooth consumption)
+  - With borrowing, buy now, smooth consumption over time
+\vfill
+- Lets formalize this
+
+## Model Set up
+
+- Two periods with no discounting
+- Agent $i$ can consume two goods
+  - non-durable gives per period utility: $u'(\cdot)>0$, $u''(\cdot)<0$
+  - lumpy durable that gives static per period utility R, if owned
+
+- Durable price = P; non-durable is numeraire
+
+- Let $Y_{1}$, $Y_{2}$ be per period 1, 2 income
+
+- Define $\bar{Y} = 1/2(Y_{1}+Y_{2})$ = average per period income
+
+## No credit constraints
+
+Without credit constraints, if don't buy, total utility is:
+$$u(\bar{Y}) + u(\bar{Y}) = 2u(\bar{Y})$$
+If buy durable, spend $P$ and spread cost equally across periods:
+$$2u(\bar{Y}-\frac{P}{2}) + 2R$$
+Purchase durable iff:
+$$ u(\bar{Y}) - u(\bar{Y}-\frac{P}{2}) \leq R$$
+**$u''(\cdot)<0$ implies acquisition increases in income**
+
+## 1) With credit constraints: Buy in period 1
+
+If don't purchase, total utility is:
+$$ u(Y_{1}) + u(Y_{2}) $$
+
+If purchase in period 1: 
+$$ u(Y_{1} - P) + 2R + u(Y_{2}) $$
+
+Purchase in period 1 iff:
+
+$$ \frac{u(Y_{1}) - u(Y_{1}-P)}{2} \leq R $$
+
+## 2) With credit constraints: Wait to buy in period 2
+
+If don't purchase, total utility is: 
+$$ u(Y_{1}) + u(Y_{2}) $$
+If save in period 1, and buy in period 2:
+$$ 2u(\bar{Y} - \frac{P}{2}) + R$$
+Wait to buy in period 2 if:
+$$u(Y_{1})+u(Y_{2}) - 2u(\bar{Y} - \frac{P}{2}) \leq R $$
+
+## Intuition of Model
+
+- $u''(\cdot)<0$ means consumers gain from smoothing consumption
+
+- Unconstrained household will buy in period 1
+  - Use period 2 income (loan) to smooth consumption
+
+- Credit constrained households cannot do this
+  - If buy now, magnify consumption inequality across periods
+  - Or, wait and buy in period 2, but delay utility gain
+
+- Higher income in any period leads savers to buy more
+
+- But, first period buyers respond only to first period income (lemma 1)
+
+- But increase in *cumulative* income increases buying through delay/saving (lemma 2)
+
+## Testable Predictions
+
+- Prediction 0: Acquisition increasing in income
+- Prediction 1: S-shaped curve in acquisition
+<!-- - Prediction 2: Faster income growth leads to more period 2 adoption -->
 
@@
 
-## How can we explain the S-shape?
-
-**Set up:**
-
-- Two periods with no discounting
-- Agent $i$ can consume two goods
-  - non-durable gives per period utility: $u'(\cdot)>0$, $u''(\cdot)<0$
-  - lumpy durable that gives static per period utility R, if owned
-
-- Durable price = P; non-durable is numeraire
-
-- Let $Y_{1}$, $Y_{2}$ be per period 1, 2 income
-
-- Let $\bar{Y} = 1/2(Y_{1}+Y_{2})$ be average income
-
-## No credit constraints
-
-Without credit constraints, if don't buy, total utility is:
-$$u(\bar{Y}) + u(\bar{Y}) = 2u(\bar{Y})$$
-If buy durable, spend $P$ and spread cost equally across periods:
-$$2u(\bar{Y}-\frac{P}{2}) + 2R$$
-Purchase durable iff:
-$$ u(\bar{Y}) - u(\bar{Y}-\frac{P}{2}) \leq R$$
-**$u''(\cdot)<0$ implies acquisition increases in income**
-
-## 1) With credit constraints: Buy in period 1
-
-If don't purchase, total utility is:
-$$ u(Y_{1}) + u(Y_{2}) $$
-
-If purchase in period 1: 
-$$ u(Y_{1} - P) + 2R + u(Y_{2}) $$
-
-Purchase in period 1 iff:
-
-$$ \frac{u(Y_{1}) - u(Y_{1}-P)}{2} \leq R $$
-
-## 2) With credit constraints: Wait to buy in period 2
-
-If don't purchase, total utility is: 
-$$ u(Y_{1}) + u(Y_{2}) $$
-If save in period 1, and buy in period 2:
-$$ 2u(\bar{Y} - \frac{P}{2}) + R$$
-Wait to buy in period 2 if:
-$$u(Y_{1})+u(Y_{2}) - 2u(\bar{Y} - \frac{P}{2}) \leq R $$
-
-## Intuition of Model
-
-- $u''(\cdot)<0$ means consumers gain from smoothing consumption
-
-- Unconstrained household will buy in period 1
-  - Use period 2 income (loan) to smooth consumption
-
-- Credit constrained households cannot do this
-  - If buy now, magnify consumption inequality across periods
-  - Or, wait and buy in period 2, but delay utility gain
-
-- First period buyers respond only to first period income (lemma 1)
-
-- But increase in *cumulative* income increases buying through delay/saving (lemma 2)
-
-## Testable Predictions
-
-- Prediction 0: Acquisition increasing in income
-- Prediction 1: S-shaped curve in acquisition
-- Prediction 2: Faster income growth leads to more period 2 adoption
-- Prediction 3: Period 1 ownership depends on interaction of income and growth
-
-## Empirical Setting: Oportunidades (again)
+## Empirical Setting: Mexico Oportunidades (again)
 
 - 320 randomly selected communities given early treatment (April 1998)
+  - 52\% were ``eligible'' for cash transfer
 \vfill
@@
 - This paper: N=506 communities, 10,000 households surveyed (1997-2007)
-\vfill
-- Compare early and late households
+  - Surveys done in  1998, 1999, 2000, 2003, and 2007
+\vfill
+- Compare early and late treatment households (**not treatment vs. control**)
   - Similar b/c both groups selected on ``vulnerability'' characteristics
   - Strongly balanced on covariates
+  - Assumes timing is random
 
@@
 $$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \beta X_{i} + \beta_{t}F_{i} + R_{rt} + \epsilon_{it}$$
-where h(a) is the prob. that $i$ buys $a$ in time $t$ conditional on not having it in $t-1$. 
-\vfill
+
+  - where h(a) is the prob. that $i$ buys $a$ in time $t$ conditional on not having it in $t-1$. 
+  - $CI_{it}$ is cumulative income
+  - $X_{i}$ includes baseline household controls
+  - $F_{i}$ are family structure variables interacted w/ survey round
+  - $R_{rt}$ are state-year fixed effects
+
+Hypothesis: $\alpha_{1} >0$
+
+## Empirical Design
 
@@
 $$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{2} CI^{2}_{it} + \beta X_{i} + R_{rt} + \epsilon_{it}$$
-Hypothesis: $\alpha_{2} >0$
-
-## Empirical Design
-
-- **Prediction 2 and 3:** conditional on having same level of cumulative income, households which accumulated income slower are less likely to acquire asset; this effect is increasing in cumulative income
-
-$$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{3} Early_{i} + \alpha_{4} (Early_{i} \times CI_{it}) + \beta X_{i} + R_{rt} + \epsilon_{it}$$
-- where $Early_{i}=1$ if $i$ began receiving transfers 18 months before control households
+
+Hypothesis: $\alpha_{2} \lessgtr 0?$
+
+<!-- ## Empirical Design -->
+
+<!-- - **Prediction 2:** conditional on having same level of cumulative income, households which accumulated income slower are less likely to acquire asset -->
+
+<!-- $$h(a_{it}) = Pr(a_{it}=1 | a_{it-1}=0) = \gamma_{0} + \alpha_{1} CI_{it} + \alpha_{3} Early_{i} + \alpha_{4} (Early_{i} \times CI_{it}) + \beta X_{i} + R_{rt} + \epsilon_{it}$$ -->
+
+<!-- - where $Early_{i}=1$ if $i$ began receiving transfers 18 months before control households -->
+<!-- - $Early_{i}$ is synonymous with treated, and $Late_{i}$ is control (not-yet-treated) -->
 
@@
 - Cumulative Income is sum of wage, farm, business income, CCT transfers
-  - measurement error in wage data
-  - cumulation requires interpolating between survey waves
+  - Reporting biases in wage data
+  - Income reported in survey year, not cumulative
+  - Authors interpolate across years, adding noise
 \vfill
@@
 \vfill
-- **Solution:** use cumulative CCT **transfers** to instrument cumulative **income**
-  - Assumes transfers affect consumption only through income
-  - Is this reasonable?
+- **Solution:** use potential CCT **transfers** to instrument cumulative **income**
 
@@
 Plus variation from early/late enrollment
+
+## Identification Assumptions
+
+- What is the exclusion restriction?
+- Is it reasonable?
 
@@
 
-## Results: Prediction 2 and 3
-- include only if time
-
 ## Thoughts on paper?
 
-<!-- We learned about effect of income on energy. What about environment? -->
-
-## Lessons from income/consumption channel?
-
-- What do we learn about how development affects the environment?
-
-- What are shortcomings of research designs? Measurement?
-
-- External validity?
-
 ## Research gaps
@@
 \vfill
+- Can we trace out the full S-curve?
+\vfill
+- Is there an S-curve for other technologies?
+\vfill
 - Which externalities have highest income elasticities?
 \vfill
-- Can the effect of income on $MWTP_{e}$ be separated from changes in consumption?
 
@@
 ## Housekeeping
+
+- Research proposal first draft: Oct 3rd
+  - If you find idea w/n 10 days, gives 2 weeks to write
+\vfill
+- Class presentations: separate Canvas upload each week
+\vfill
+- Replication project?
+
+## Recap: Alix-Garcia et al. (2013)
+
+- Do we believe the story about increased demand for land-intensive goods?
+  - Less market access means less access to inputs
+  - Underinvest in technology, leading to higher deforestation
+  - Maybe this is a story about technology, not consumption
+
+- Can you think of an alternate explanation?
+  - Talk to your neighbor for five minutes
+
+- Other concerns?
 
@@
 - Case II: Development spurs infrastructure development
-  - Asher et al. (2020)
+  - Garg et al. (2023)
 
@@
 
+## Existing Work
+
+- Theory: (Angelsen and Kaimowitz,2001; Balsdon, 2007; Goldstein et al., 2012; Green et al., 2005; Phalan et al.,2016; Takasaki,2006). 
+  - Most models yield ambiguity
+\vfill
+- Most work on agricultural productivity and deforestation
+  - Intensive margin dominates: Abman et al. (2024), Abman and Carney (2020a), Abman and Carney (2020b), Assuncao et al. (2023)
+\vfill
+- New work: market access and deforestation 
+  - Abman and Lundberg (2024)
+\vfill
+- Open Topics:
+  - Effect of agriculture on deforestation in the first place? (Madhok et al. 2024)
+  - Role of factor markets? Labor vs. input markets
+  - Mechanisms behind market access effects
+  - Agriculture, migration and deforestation
+
 ## Conceptual framework
@@
 - Continuum of agents with heterogeneous outside options
-  - Changes in returns to ag affect number of agents who select into farming
+  <!-- - Changes in returns to ag affect number of agents who select into farming -->
 
@@
 - Differentiating with respect to $\Omega$ gives:
-$$\frac{\partial D^{*}}{\partial \Omega} = \underbrace{n^{*}\frac{\partial \gamma^{*}}{\partial \Omega}}_\text{$\leq 0$} + \underbrace{\gamma^{*}\frac{\partial n^{*}}{\partial \Omega}}_\text{$\geq 0$}$$
+$$\frac{\partial D^{*}}{\partial \Omega} = \underbrace{n^{*}\frac{\partial \gamma^{*}}{\partial \Omega}}_\text{$\lessgtr 0$} + \underbrace{\gamma^{*}\frac{\partial n^{*}}{\partial \Omega}}_\text{$\lessgtr 0$}$$
 
@@
 
-- Agricultural survey on 7,781 households (451 villages) in 2011
+- BRAC agricultural survey on 7,781 households (451 villages) in 2011
   - Detailed cropping practices from last two seasons
@@
 - Alternative explanations
-- External validity (recall forest is private in Uganda)
+- External validity
 
@@
 - Roads may facilitate agricultural labor loss
-  - Will farmers turn to crop burning, despit the costs?
+  - Will farmers turn to crop burning, despite the costs?
 
@@
 ## Data
-- Roads: SHRUG database (N=11,151 villages w/o paved road at baseline)
+- Roads: SHRUG database 
+  - N=11,151 villages w/o paved road at baseline (2001)
+  - Main variable: road construction date in village
 \vfill
 - Fires: NASA EODIS
-  -\# fires w/n 10km of village polygon
+  - \# fires w/n 10km of village polygon
 \vfill
 - Pollution: Van donkelaar et al. (2016) measure PM2.5 at 10km resolution
+  - Black Carbon and Organic Carbon from biomass/non-biomass sources
 \vfill
@@
 ```
+- Note: optimal bandwidth = +/- 84 around population threshold
 - Road access increases annual crop fires by 60\%
@@
 ## Mechanism: Roads induce labor exit
-Back to RDD design:
+- Back to RDD design
+- Data from Socioeconomic and Caste Census (2013), provided in SHRUG
+- Outcome: ag. labor share in districts with high/low ag wage
 ```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.show='hold',  out.width = "85%", fig.align = "center"}
@@
 
-## Next week
+## Next Time
 

### 2024-11-11 lectures/module4/module4.Rmd
added .Rmd files

@@
 - \usepackage{gensymb}
+- \usepackage{cancel}
 ---
@@
 
-\large WTP for Environmental Quality in Developing Countries
+\large WTP Estimation: Revealed Preference Models
 
@@
 
-- State things that are due.
+- Research Proposal: you do NOT have to carry out analysis
+\vfill
+- First draft: Oct 3rd
+\vfill
+- Replication: Oct 21st
+  - Start soon
 
@@
   - Behavioral models
-  - Revealed preference approaches
+  - Choice probabilities
+  - Estimation
+  - Application to WTP
 
@@
 
-## Measuring Willingness To Pay
+## Estimating Non-market Value of Goods and Services
 
-- Slides on Logit and Mixed Logit
+- **Stated Preference:**
+  - Hypothetical data to estimate **ex-ante** WTP
+  - Contingent valuation, contingent behavior
+  - Respondents directly asked about WTP (phone, mail, etc.)
+  - Issue: hypothetical bias and strategic bias
+\vfill
+
+- **Revealed preferences**
+  - Behavioral data to estimate **ex-post** WTP
+  - Travel cost, averting behavior, hedonic price
+  - Pro: based on **actual** choices
+  - Con (?): need a **behavioral model** in which to analyze choices
+
+## Behavioral Models
+
+- Our goal is to define a behavioral process generating agent's choice
+
+- Let agent face choice among set of options (products, actions, etc.)
+  - What leads agent to choose one?
+
+- Let some factors ($x$) be observed by researcher, and some not ($\epsilon$)
+
+- Factors relate to agent choice through a **behavioral process**:
+
+$$y = h(x, \epsilon)$$
+
+- \(h(x, \epsilon)\) determines the selected choice based on these factors
+
+## Probabilistic Choice
+- Since \(\epsilon\) unobserved, the outcome is probabilistic
+
+- **Choice probability** of choosing $y$ derived by assuming $\epsilon \sim f(\epsilon)$
+$$P(y | x) = \text{Prob}(\epsilon \quad \text{s.t.} \quad h(x, \epsilon) = y)$$
+
+- This probability is much more tractable i.e., fully characterized by $f(\epsilon)$
+
+- We can rewrite this as the expected value of an indicator function:
+$$P(y | x) = \int I[h(x, \epsilon) = y] f(\epsilon) d\epsilon$$
+
+- where $I[h(x, \epsilon) = y] =1$ when statement in brackets is true
+- Integral of behavioral process indicator over all possible values of unobservables
+
+## Evaluating the Choice Probability
+
+- Method 1: Closed-form solution
+  - For certain specifications of $h$ and $f$, $P(y|x)$ calculated via a formula
+  - Main example is Logit
+
+\vfill
+- Method 2: Simulation
+  - Some specifications of $h$ and $f$ have no closed form solution
+  - We simulate $P(y|x)$ by taking random draws from $f$ and taking the average
+
+## Evaluating the Choice Probability: Closed Form
+
+**First, define the behavioral model**
+
+- Consider a binary model where agent considers whether to take an action
+- Agent gets utility (+/-) from the action:
+
+$$U= \beta'x + \epsilon$$
+
+  - $U$: Utility of taking an action; $\beta' x$: Observed component of utility.
+  - $\epsilon$: Unobserved component; lets assume **logistic distribution**.
+
+- Take action if utility is positive:
+$$P = \int I[\beta' x + \epsilon > 0] f(\epsilon) d\epsilon$$
+
+## Closed Form Solution
+\begin{align*}
+P & = \int I[\beta' x + \epsilon > 0] f(\epsilon) d\epsilon \\
+ & = \int_{\epsilon > -\beta' x} f(\epsilon) d\epsilon \\
+ & = 1 - F(-\beta' x) \\
+ & = 1 - \frac{1}{1 + e^{\beta' x}} \\
+ & = \frac{e^{\beta' x}}{1 + e^{\beta' x}} \\
+\end{align*}
+
+## Evaluating the Choice Probability: Simulation
+
+Recall: $P(y | x) = \int I[h(x, \epsilon) = y] f(\epsilon) d\epsilon$
+
+\vfill
+
+- **Step 1:** Take draw of $\epsilon$ from $f(\epsilon)$
+  - Label it $\epsilon^{1}$, denoting first draw
+\vfill
+- **Step 2:** Check whether $h(x, \epsilon^1)=y$
+  - If so, create $I^{1}=1$, otherwise $I^{1}=0$
+\vfill
+- **Step 3:** Repeat this R times and collect $I^{r}$ for $r=1,...,R$
+\vfill
+- **Step 4:** Average **simulated** probability: $P(i | x) = \frac{1}{R}\sum_{r=1}^{R} I^{r}$
+  - Proportion of times that draws of $\epsilon$, when combined with $x$, yield $y$
 
 ## 
-\huge How is this implemented in practice?
+\huge Lets define a realistic behavioural model
+
+\large Grounded in economic theory
+
+## Random Utility Model
+
+**Behavioural model:**
+
+- Agent, $n$ faces choice among $J$ alternatives
+  - Choice set must be 1) mutually exclusive, 2) exhaustive, 3) finite
+
+- Utility from $j$ is: $U_{nj}$ for $j=1,...,J$
+
+- Utility known to agent but not the researcher
+
+- Agent chooses $j$ that maximizes utility:
+
+$$ \text{Choose i iff} \quad U_{ni} > U_{nj} \quad \forall j\neq i$$
+
+## Random Utility Model
+
+**The Researcher**
+
+- Observes attributes of alternatives, $x_{nj} \quad \forall j$
+
+- Observed attributes of agent, $s_{n}$
+
+- Let $V_{nj} = V(x_{nj}, s_{n}) \quad \forall j$ (representative utility)
+
+- **V can depend on parameters unknown to researcher**
+  - But can be structurally estimated...
+
+## Random Utility Model
+
+- Since $V_{nj} \neq U_{nj}$, we can write:
+$$U_{nj} = V_{nj} + \epsilon_{nj}$$
+
+  - where joint density of random vector $\epsilon_{n} = (\epsilon_{n1},...,\epsilon_{nJ}) = f(\epsilon_{n})$
+
+- Then the probability than agent $n$ chooses alternative $i$ is:
+\begin{align*}
+P_{ni} & = P(U_{ni} > U_{nj}, \forall j \neq i) \\
+ & = P(V_{ni} + \epsilon_{ni} > V_{nj} + \epsilon_{nj}, \forall j \neq i) \\
+& = P(\epsilon_{nj} - \epsilon_{ni} < V_{ni} - V_{nj}, \forall j \neq i) \leftarrow \text{{\color{blue} CDF!}} \\
+ & = \int I(\epsilon_{nj} - \epsilon_{ni} < V_{ni} - V_{nj}) f(\epsilon) d\epsilon
+\end{align*}
+- Different DCMs obtained from different choices of $f(\epsilon)$
+
+## Logit
+
+- Assumes $\epsilon_{nj}$ distributed iid extreme value
+$$ \text{PDF:} \quad f(\epsilon_{nj}) = e^{-\epsilon_{nj}} e^{-e^{-\epsilon_{nj}}}, \quad \text{CDF:} \quad F(\epsilon_{nj}) = 1 - e^{-e^{\epsilon_{nj}}}$$
+- Note: if $\epsilon$'s are iid EV, then $\epsilon_{nji}^{*} = \epsilon_{nj} - \epsilon_{ni}$ is distributed logistic:
+$$F(\epsilon_{nji}^{*}) = \frac{e^{\epsilon_{nji}^{*}}}{1+{\epsilon_{nji}^{*}}}$$
+
+## Aside: Extreme Value Distribution
+
+:::::::::::::: {.columns}
+::: {.column width="45%"}
+
+**Useful Properties**
+
+\vspace{0.3cm}
+
+1. Mathematical tractability
+    - Closed form choice probability
+
+\vspace{0.3cm}
+2. Long right tail
+    - Unobservables can drive unexpected choices
+
+\vspace{0.3cm}
+
+3. IIA property
+
+\vspace{0.3cm}
+
+4. Connection to logistic regression
+
+\vspace{0.3cm}
+
+5. Easy to (structurally) estimate
+
+:::
+::: {.column width="45%"}
+```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/ev_pdf.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+:::
+::::::::::::::
+
+## Logit
+
+- The probability that agent $n$ chooses alternative $i$:
+\begin{align*}
+P_{ni} & = P(V_{ni} + \epsilon_{ni} > V_{nj} + \epsilon_{nj} \quad \forall j\neq i) \\
+&= P(\epsilon_{nj} < \epsilon_{ni} + V_{ni} - V_{nj} \quad \forall j\neq i) \leftarrow \text{{\color{blue} invoke iid assumption!}} \\
+ &= \int \left( \prod_{j \neq i} e^{-e^{-(\epsilon_{ni} + V_{ni} - V_{nj})}} \right) e^{-\epsilon_{ni}}e^{-e^{-\epsilon_{ni}}}d\epsilon_{ni} \\
+&= \frac{e^{V_{ni}}}{\sum_{j}e^{V_{nj}}}
+\end{align*}
+\vfill
+- **Preview:** If V is linear in parameters, $V_{nj} = \beta'x_{nj}$, then: $P_{ni} = \frac{e^{\beta'x_{ni}}}{\sum_{j}e^{\beta'x_{nj}}}$
+
+## Independence of Irrelevant Alternatives
+
+- Crucial assumption in logit model
+
+- Relative odds of choosing b/w 2 $i$ and $k$ does not change when $j$ introduced
+
+- For any two alternatives $i$ and $k$, the ratio of logit probabilities is:
+\begin{align*}
+\frac{P_{ni}}{P_{nk}} &= \frac{e^{V_{ni}}/ \sum_{j}e^{V_{nj}}}{e^{V_{nk}}/ \sum_{j}e^{V_{nj}}} \\
+&= \frac{e^{V_{ni}}}{e^{V_{nk}}} = e^{V_{ni}-V_{nk}}
+\end{align*}
+
+- Ratio does not depend on any alternatives other than $i$ and $k$!
+  - Denominators cancel
+
+## Discussion of IIA Assumption
+
+- Advantages
+\vfill
+
+- Disadvantages
+
+##
+\centering
+\huge 5 minute break
+
+## Estimation: Maximum Likelihood Structural Estimation
+
+- Suppose we have have data on $N$ agent's choices
+- The likelihood of observing the choices made by each individual is:
+
+$$L(\beta) = \prod_{n=1}^{N} \prod_{i} (P_{ni})^{y_{ni}}$$
+
+- where \(y_{ni} = 1\) if alternative \(i\) is chosen, and 0 otherwise
+- Log-likelihood is easier to maximize:
+$$LL(\beta) = \sum_{n=1}^{N} \sum_{i} y_{ni} \log(P_{ni})$$
+
+- ML estimator is the $\beta$ that maximizes this function
+
+## Method of Moments
+
+- FOC of $LL(\beta)$ $\leftrightarrow$ $\frac{\partial LL(\beta)}{\partial \beta}=0$
+- This can be rewritten as:
+$$\frac{1}{N} \sum_{n=1}^{N} \sum_{i} y_{ni} x_{ni} = \frac{1}{N} \sum_{n=1}^{N} \sum_{i} P_{ni} x_{ni}$$
+
+- LHS: mean $x$ over choices by sampled agents
+- RHS: mean $x$ over **predicted** choices by sampled agents
+
+- $\beta$ found s.t. predicted mean of each explanatory variable equals observed mean
+  - Model reproduces observed averages in the data
+  - Also known as **matching on the first moment**
+
+## Limitations
+
+- Cannot represent random taste variation
+
+- Implies proportional substitution across alternatives (IIA)
+
+##
+\Large What if we want to estimate more moments of $\beta$?
+
+## Specifying Taste Variation
+
+- Supposed agent now has the utility: $U_{nj} = {\color{red} \beta_{n}}'x_{nj} + \epsilon_{nj}$
+  - where $\beta_{n}\sim f_{n}(\beta|\Theta)$ and  $\epsilon_{nj}$ still iid EV
+
+- If we observed $\beta_{n}$, then we are back to logit: $P_{ni}(\beta_{n})=\frac{e^{\beta_{n}'x_{ni}}}{\sum_{j}e^{\beta_{n}'x_{nj}}}$
+
+- Instead, we integrate logit over all possible parameters of $\beta_{n}$:
+
+$$P_{ni} = \int \underbrace{\left(\frac{e^{\beta_{n}'x_{ni}}}{\sum_{j}e^{\beta_{n}'x_{nj}}}\right)}_\text{logit probability} \underbrace{f(\beta | \Theta)d(\beta)}_\text{mixing distribution} \quad \leftarrow \text{{\color{blue} no closed form solution!}}$$
+
+- Mixing distribution usually normal: $f(\beta|\Theta) = f(\beta|\mu,\sigma)$
+- Note: $P_{ni}$ is a function of $\Theta$, not $\beta$! ($\beta$ is integrated out)
+
+## Aside: Properties of Mixed Logit
+
+- Relaxes IIA assumption!
+\vfill
+- Ratio of mixed logit probabilities, $P_{ni}/P_{nj}$ depends on all data
+  - Denominators are inside the integral and do not cancel
+\vfill
+- Enables correlation in unobserved factors across alternatives
+  - Flexible substitution patterns
+
+## Estimation: Simulation
+
+**Steps for Simulating Mixed Logit Probability**
+
+1. Draw a value $\beta^r$ from $f(\beta | \Theta)$, where $r$ is which draw
+2. Calculate the logit formula $P_{ni}(\beta^r)$ using the drawn value
+3. Repeat steps 1 and 2 for *R* draws.
+4. Average the results to obtain the simulated probability:
+
+$$ \hat{P}_{ni}= \frac{1}{R} \sum_{r=1}^{R} P_{ni}(\beta^r)$$
+
+## Simulated Log-Likelihood Estimation
+
+- As before, insert simulated $\hat{P}_{ni}$ into log-likelihood:
+
+$$SLL = \sum_{n=1}^{N} \sum_{j=1}^{J} y_{nj} \ln \hat{P}_{ni}$$
+
+- where $y_{nj} = 1$ if $n$ chose $j$ and zero otherwise
+
+- Maximum simulated likelihood estimator (MSLE) is the $\Theta$ that maximizes SLL
+\vfill
+**This recovers taste distribution $(\mu, \sigma)$ that matches observed choices to predicted choice probabilities as close as possible**
+
+##
+\centering
+\huge Lets do a simple application
+
+\large What is the WTP for biodiversity?
+
+## WTP for Species Diversity in India
+
+- **Context:** people travel to ``hotspots'' to see pretty birds
+  - Birds are a good proxy for overall biodiversity
+\vfill
+- **Data:** individual level data from citizen science app (eBird)
+  - location, alternative sites, income, **travel cost**, **species richness**
+\vfill
+- **Behavioral model:** choose site that maximizes utility
+\vfill
+- **Question:** what cost will you incur to $\uparrow$ utility from biodiversity?
+  - How does your ``price'' change for an additional unit of biodiversity? (MWTP)
+
+## Deriving WTP in Practice
+
+- Agent $i$'s utility from site $j$ at time $t$ is:
+\begin{align*}
+U_{ijt} & = V_{ijt} + \epsilon_{ijt} \\
+& = \underbrace{\beta_{1}}_{\text{MU Income}}(y_{i} - c_{ijt}) + \underbrace{(\beta_{2} + \eta_{i})}_{\text{MU EQ}} e_{jt} + ... + \epsilon_{ijt}
+\end{align*}
+
+- If stay home: $U_{i0t}=\beta_{1}y_{i} + \epsilon_{i0t}$
+- Cost s.t. $i$ indifferent b/w travelling and avoiding cost:
+$$c_{ijt}* = \frac{1}{\beta_{1}} [(\beta_{2} + \eta_{i})e_{jt} + \beta_{3} x_{ijt} + ... + (\epsilon_{ijt}-\epsilon_{i0t})]$$
+- MWTP for one more species, assuming $\eta_{i}\sim N(0,\sigma)$: 
+$$MWTP=\boxed{\frac{\partial c_{ijt}*}{\partial d_{jt}} = \frac{\hat{\beta_{2}}}{\hat{\beta_{1}}} + \xcancel{\frac{\hat{\eta_{i}}}{\hat{\beta_{1}}}}}$$
+
+## Implementation
+
+- Data
+  - attributes of choice and alternatives for each **choice occasion**
+  - Variable *choice=1* if chosen, otherwise 0
+  - Data can get huge, and simulation takes forever
+  - solution: restrict choice set
+\vfill
+- Stata
+  - clogit: conditional logit (fast)
+  - mixlogit: mixed logit estimation (slow)
+\vfill
+- R
+  - *mlogit* library
+  - Exercises and applications: https://cran.r-project.org/web/packages/mlogit/vignettes/c5.mxl.html
+
+## Aside
+
+- **If you want to get involved in this project, let me know**
+
+- Goal: estimate WTP for species diversity in India
+
+##
+\centering
+\huge Lecture 7
+
+\large Why is WTP for Environmental Quality Low in Developing Countries?
+
+## Housekeeping
+
+- Free lecture (Oct 10)
+  - Choose one: 1) firms, 2) migration, 3) policy (guest lecture)
+\vfill
+- Research Proposal: first draft due one week from Thursday
+\vfill
+- All assignments have detailed instructions and rubric
+\vfill
+- Start selecting replication paper
+
+## WTP for Environment Quality in LMICs
+
+- Crucial for policy design (Greenstone and Jack, 2013)
+\vfill
+- Most papers estimate $MWTP_{s}$
+  - and assume market failures do not bias valuation
+\vfill
+- Examples of $MWTP_{s}$
+  - MWTP for water filter in Ghana: \$1.80-2.40 (Berry et al. 2011)
+  - MWTP for clean cookstoves in Kenya: \$12 (Berkouwer and Dean, 2022)
+  - MWTP for sanitation in Cambodia: \$30 (Ben-Yishay et al., 2017)
+\vfill
+- Revealed Preferences: very few!
+  - WTP for clean water: Kremer et al. (2011) ${\color{red} \leftarrow today}$
+  - WTP for clean air: Ito and Zhang (2020)
+  - \textcolor{red}{WTP for biodiversity: ???}
 
@@
 - **Research Question:** What is household WTP for clean water in LICs?
-  - What are health impacts of water quality improvements? (we will skip this)
+  - What are health impacts of water quality improvements?
 \vfill
@@
 \vfill
+- 7-8 households from each spring select from ``user list''
+\vfill
 - Phased design: all springs eventually protected
@@
   - Round III (2005-2006); 91 springs; 685 households
-\vfill
-- 7-8 households from each spring select from ``user list''
 
@@
   - $T_{jt}$ is treatment status (=1 after protection)
+  - $\phi$ = impact of spring protection on water quality
 \vfill
@@
 \vfill
-- $C_{i}>0$ is time cost of household $i$
+- $C_{i}>0$ is time cost per minute of household $i$
 \vfill
@@
 ## Mixed Logit
-- Follow-up survey gives list of alternative water sources
-- What is ratio of coeff. on $T_{jt}$ to coeff. on walking time?
-  <!-- - Value of spring protection in terms of minutes spent walking -->
 
@@
   - where $f(\cdot)$ is the mixing distribution
-- Maximize log-likelihood to estimate mean and s.d. of these distributions
+- Maximize SLL to estimate mean and s.d. of these distributions
+
+- **What is the object of interest?**
+  <!-- - ratio of$\beta_{i}$ to coeff. on travel cost -->
+  <!-- - Value of spring protection in terms of minutes spent walking -->
 
@@
   - Coefficient on spring protection ($\beta_{i}$) = 0.51
+\vfill
+- **Can $\beta_{i}$ entangle other amenity values?**
 
-- Can $\beta_{i}$ entangle other amenity values?
   <!-- - reduced need to bend down, lower collection times -->
@@
 
-- Households with kids under 3 find walking more costly
-  - Coeff. on interaction b/w $T_{jt}$ and child is positive (insignificant)
+\vfill
+- Coeff. on ($T_{jt}$ $\times$ under-3 child) is positive (insignificant)
   - Households with kids value water quality more
@@
 
-- Ratio of mixed logit coefficients: 2.95/0.21
+- Ratio of **mixed logit** coefficients: 2.95/0.21=14 minutes
 
@@
   - Note: in paper, it is 32.4 workdays!
-
-## Comparison with Stated Preference Values
 
@@
 - Regress time value on household characteristics, and predict on full sample
-- Mean value of time = \$USD 0.088 per 8-hour day (7\% of ag wages)
+- Mean value of time = \$USD 0.088 for 8-hr workday
 
@@
 
+## Comparison with Contingent Valuation
+
+- Households in protected spring communities asked:
+  - **How much are you WTP per year to keep the spring protected?**
+\vfill
+- In follow up survey:
+  - First: are you willing to pay 500 shillings (\$7.14)?
+  - Second: are you willing to pay 1000 shillings (\$14.29)?
+  - Then emphasize tradeoff (opportunity cost) and ask again
+\vfill
+- Almost all WTP \$7.14
+  - Majority WTP \$14.29 even after learning about tradeoff
+  - Starting value matters!
+
+<!-- ## Distribution of Household Values of Clean Water (Mixed Logit) -->
+<!-- ```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"} -->
+
+<!-- knitr::include_graphics( -->
+<!--                         path = "assets/kremer2011_fig1.png", -->
+<!--                         auto_pdf = TRUE -->
+<!--                         ) -->
+
+<!-- ``` -->
+<!-- - Why is stated preference values larger? Why is CV distribution wider? -->
+
 ## Health Valuation and VSL
 
+- Spring protection averts **3.2 diarrhea cases per household-year**
+  - (0.047 cases/child-week) * (1.3 children under 3 per hh) * (52 weeks/year)
+\vfill
+- Annual WTP for spring protection = 32.4 workdays/year
+  - Assume this value captures benefits from averted diarrhea
+\vfill
+- If protection prevents 3.2 cases/yr, valued at 32.4 workdays/yr, then
+\vfill
+$\text{WTP/case averted} = \frac{\text{Total Workdays}}{\text{Diarrhea Cases Averted}} = \frac{\text{32.4 workdays}}{\text{3.2 cases}} = \text{10.1 workdays/case}$
+
+## Health Valuation and VSL
+
+- Assume spring protection $\downarrow$ diarrhea morbidity \& mortality proportionally
+
+- Using fact that 862 diarrhea cases $\rightarrow$ 1 death (literature):
+
+- WTP to prevent 862 diarrhea cases (1 death) i.e. VSL:
+$$ \text{Total WTP/death averted}= \text{10.1 workdays} \times \text{862 cases} \approx \text{8,700 workdays}$$
+
+**VSL = \$0.088 $\times$ 8700 $\approx$ \$USD 770**
+
+**In USA 2013: VSL $\approx$ \$USD 8.5 million (EPA)**
+
 ## Discussion
-<!-- - what do you think about approach? -->
-<!-- - what do you think about time valuation? -->
-<!-- - Why not include income directly in indirect utility fxn? -->
 
-## Distribution of Household Values of Clean Water (Mixed Logit)
-```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}
+- What do you think about approach?
+- What do you think about time valuation?
+
+## 
+\centering
+\Huge Lecture 8
+
+\large Why is WTP for Environmental Quality Low in LMICs?
+
+## Housekeeping
+
+- First draft due one week from today
+
+- Select your replication paper
+
+<!-- ## Today's lecture -->
+
+<!-- - Why is willingness to pay so low in developing countries? -->
+
+<!-- - One reason: **Revealed $MWTP_{e} \neq MWTP_{e}$ from first best** -->
+
+<!-- - Market failures create a wedge -->
+
+<!-- - Today: experiments to alleviate frictions -->
+
+## Energy efficiency uptake in LMICs
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}
 
 knitr::include_graphics(
-                        path = "assets/kremer2011_fig1.png",
+                        path = "assets/bd2022_fig1.png",
                         auto_pdf = TRUE
@@
 ```
-- Why is stated preference values larger? Why is CV distribution wider?
+
+## Market Failures in Development Economics
+
+- Market failures can shape observed $MWTP_{e}$
+  - And, therefore, environmental policy
+
+\vfill
+- Main frictions: information, credit, risk, or property right imperfections
+  - Frictions warp $MWTP_{e}$, resulting in lower EQ
+
+\vfill
+- **Revealed preference** $MWTP_{e}$ assumes agent knows payoffs
+  - Misinformation common in developing countries
+  - Due to literacy, education, information access, state capacity, etc.
+
+## Credit Market Failures
+
+- Stem from difficulty writing/enforcing contracts in developing countries
+\vfill
+- Costs of monitoring borrowers + enforcing repayment leads to high interest rates and credit rationing (Conning and Udry, 2007)
+\vfill
+- Agents cannot pay upfront for clean technology that has future payoffs
+  - Forego future improvements in EQ, income, and health
+\vfill
+- **Determinants of liquidity therefore confound measured $MWTP_{e}$**
+
+## Missing Risk Markets
+
+- Lower $MWTP_{e}$ if payoffs are uncertain and insurance unavailable
+\vfill
+- Worse EQ can increase var(Y)
+\vfill
+- Missing risk markets $\uparrow$ exposure to risk and $\downarrow$ measured $MWTP_{e}$
+  - Liquidity: prioritize income for emergencies, instead of clean water
+  - Discounting: undervalue investment in EQ because uncertain about whether will be able to enjoy benefits (discount future more heavily)
+
+## Evidence on the effect of market failures on MWTP
+
+- What is the ideal experiment?
+
+- How do you distinguish between measured and actual $MWTP_{e}$?
+
+- Would random variation help?
+
+##
+
+\begin{quote}
+`` there is hardly a more important topic
+for future study than developing revealed
+preference measures of $MWTP_{e}$ that capture
+the aesthetic, health, and/or income gains
+from environmental quality''
+\vspace{0.5cm}\\
+- Envirodevonomics (Greenstone and Jack, 2013)
+\end{quote}
+
+## Berkouwer and Dean (2022): Frictions in WTP for Clean Air
+
+- **Research Question:** What are the barriers to adopting clean cookstoves?
+
+- **Setting:** Nairobi, Kenya
+  - Test whether credit constraints and inattention to cost-savings lead to underadoption
+
+- **Data:** 1000 households
+  - Measure WTP with BDM mechanism
+
+- **Research Design:** RCT with 1000 households to quantify WTP drivers
+  - Three-by-three experimental design
+  - Cross-randomize two credit and two attention treatments
+
+- **Results:** 1) Stove generates \$237 in charcoal savings, 2) yet only \$12 WTP, 3) credit constraints are the main friction preventing takeup
+
+## Key Idea
+
+WTP for enviromental quality $\rightarrow$ WTP for improved stove
+
+## Experimental Setting
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig2.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+
+## Model of Adoption
+- Consider agent evaluating an energy efficient technology (costs $P_{E}$)
+- Adopt iff sum of savings ${\color{blue} S_{t}}$ exceed costs today:
+$$\underbrace{u(c_0) - u(c_0 - P_E)}_{\text{cost of adoption today}} < \underbrace{\sum_{t=1}^{T} \delta^t [u(c_t + {\color{blue} S_t}) - u(c_t)]}_{\text{future benefits}}$$
+- Agent can take loan ${\color{red} L}$ and pay back in payments ${\color{red} r_{t}}$:
+$$\underbrace{u(c_0) - u(c_0 - P_E+{\color{red} L})}_{\text{cost of adoption today}} < \underbrace{\sum_{t=1}^{T} \delta^t [u(c_t + {\color{blue} S_t} - {\color{red} r_{t}}) - u(c_t)]}_{\text{future benefits}}$$
+- WTP $P^{*}$ is when agent is indifferent
+$$u(c_0) - u(c_0 - P^{*}+{\color{red} L}) = \sum_{t=1}^{T} \delta^t [u(c_t + {\color{blue} S_t} - {\color{red} r_{t}}) - u(c_t)]$$
+
+## Model of Adoption
+
+**Market Failure Tests**
+
+- Agent may be **inattentive** to energy savings by ${\color{blue} \theta_{s}} \in [0,1]$
+  - concentration bias: less attentive to small payments ${\color{red} r_{t}}$
+  - Time inconsistency: inattentive to future loan payments ${\color{red} \theta_{c}} \in [0,1]$
+\vfill
+- Agent may be **credit constrained** ${\color{red} L \leq \bar{C}}$
+
+## Empirical Tests
+
+Main outcome variable: WTP $P^{*}$
+$$u(c_0) - u(c_0 - P^* + {\color{red} L}) = \sum_{t=1}^{T} \delta^t \left[ u(c_t + {\color{blue} \theta_s S_t} - {\color{red} \theta_c r_t}) - u(c_t) \right] \quad \text{s.t.} \quad {\color{red} L \leq \overline{C}}$$
+
+1. Randomize attention to energy savings $\theta_s$ \hfill $\frac{\partial P^*}{\partial \theta_s} > 0$
+
+2. Randomize access to credit $\overline{C}$ \hfill $\frac{\partial P^*}{\partial \overline{C}} > 0$
+    - Randomize into \( N \) payments of size \( r_t \) \hfill $\frac{\partial^2 P^*}{\partial \overline{C} \partial r_t} < 0$
+
+    - Randomize attention to costs \( \theta_c \) \hfill $\frac{\partial^2 P^*}{\partial \overline{C} \partial \theta_c} < 0$
+
+    - Measure time-inconsistency (TI) \hfill $\frac{\partial^2 P^*}{\partial \overline{C} \partial TI} > 0$
+
+## Research Design: Cross-Randomize Credit and Attention
+- Each cell contains number of respondents
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig3.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+
+## Credit: Control and Treatment Groups
+
+- **Control:** pay (subsidized) price today, receive stove today
+
+- **Treatment:** receive stove today, pay w/n 3 months
+  1. Pay by 12 weekly deadlines
+  2. Pay by 3 monthly deadlines
+
+- SMS reminders if miss deadlines
+
+- Interest rate of 1.16\% per month, 14.8\% per year
+
+## Attention: Control and Treatment
+
+- **Control:** Informed that stove reduces charcoal use by 50\%
+  - Informed of \$ equivalent of savings
+\vfill
+- **Treatment 1:** Attention to energy savings
+  - Respondent writes how much they think they'll save
+  - Writes down how they will use the savings
+\vfill
+- **Treatment 2:** Attention to energy savings minus costs
+
+## Attention to Savings
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig4.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+
+## Attention to Savings minus Costs
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig5.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+
+## Elicit WTP in Each Group
+
+- **Becker-DeGroot-Marschak (BDM) Mechanism**
+  - Field worker puts random price $P_{i} \in (0.01-29.99)$ in envelope
+  - Elicit max $WTP_{i}$ from respondent using binary search
+  - Open envelope
+  - WTP is binding: pay $P_{i}$ iff $P_{i} \leq WTP_{i}$
+\vfill
+- Benefits
+  - Generates **incentive-compatible** behavior! Why?
+  - Adoption is random, conditional on WTP
+
+\vfill
+**BDM mechanism elecits WTP AND generates random stove adoption**
+
+## Stove reduces charcoal spending by \$2.28 per week!
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig6.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+
+- Large effects: \$120 savings per year
+- Given \$40 cost of stove, this is **300\% rate of return!**
+
+## Yet, demand is low
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig7.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+
+## No evidence of inattention to energy savings!
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig8.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+- Information frictions may not be attenuating WTP
+
+## Access to credit doubles WTP!
+```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}
+
+knitr::include_graphics(
+                        path = "assets/bd2022_fig9.png",
+                        auto_pdf = TRUE
+                        )
+
+```
+- credit increases WTP by \$13 (104\% relative to control)
+<!-- - credit push demand curve to breakeven point (i.e. use savings to pay for stove) -->
+
+## Other Reasons why WTP is low
+
+- Information access: Baylis et al. (2024), Jalan and Somanathan (2008), Bennear et al. (2013)
+\vfill
+- Credit: Ben-Yishay et al. (2017)
+\vfill
+- Property rights: Ali et al. (2014), Abman and Carney (2020), Bellemare et al. (2018), Noack and Costello (2024)
+
+## Discussion
+
+- Market frictions affects many things $\rightarrow$ endogeneity
+
+- Partial vs. general equilibrium
+
+- Cannot separate measurement from actual distortions
+  - Compare across methods more/less affected by frictions?

### 2024-11-11 lectures/module5/module5.Rmd
added .Rmd files

---
title: "| Environmental and Development Economics\n| Module 5 - Environmental Policy Design
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 5 - Environmental Policy Design]{Environmental and Development
  Economics \\\\ Module 5 - Environmental Policy Design in LMICs}}"
- \input{header.tex}
- \usepackage{gensymb}
---

## 
\centering
\Huge Lecture 8 (continued)

\large Environmental Policy Design in LMICs: Monitoring and Enforcement

## Housekeeping

## Rich countries have resolved many environmental problems
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/regulation_fig1a.png",
                        auto_pdf = TRUE
                        )

```

## Within-country: Richer countries effectively use environmental regulation
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "50%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/regulation_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Role of the state in improving environmental quality

- With perfectly functioning markets, agent sets:

$$MWTP_{e} = \frac{\partial c_{e}}{\partial \Delta e}$$

- This yields **first-best outcome**
\vfill
- For policy design, social planner sets optimal $e$
  - **Assumption 1:** Assume they know true $MWTP_{e}$ (last week)
  - **Assumption 2:** Costless technology to raise revenue
\vfill
- **Assumption 2** fails in LMICs $\longleftrightarrow$ $\frac{\partial c_{e}}{\partial \Delta e}$ too high!

## Optimal Environmental Policy Design 101

:::::::::::::: {.columns}
::: {.column width="45%"}

- LMICs on left of equilibrium
\vspace{0.3cm}
- Optimal to abate until MAC=MB
\vspace{0.3cm}
- Should be easy since MAC low
\vspace{0.5cm}

**puzzle:  $\frac{\partial c_{e}}{\partial \Delta e}$ is high!**

- why?

:::
::: {.column width="45%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/mac.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Today

- **Guiding question:** How do we design optimal environmental policy in LMICs?

\vfill
- Today's focus: challenges for designing low MAC policies
  - Perverse incentives (Duflo et al., 2013)
  - Spillovers and unintended consequences (Englander 2023)

\vfill
- Next time: other barriers to efficient policy design
  - Trust
  - Knowledge

## Background: Poor countries DO have environmental policies

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/regulation_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Background: India

- Command-and-control regulation is the norm
\vfill
- "Command" is easy, "control" is costly
\vfill
- State Pollution Control Board (SPCB) enforces regulatons
  - State capacity is weak
  - Enforcement is very costly $\rightarrow$ less monitoring
\vfill
- **Role of third-party auditors**

## Duflo et al. (2013): Truth-Telling by Third-party Auditors

- **RQuestion:** How much do firms skirt environmental policy in developing countries?
  - Can realigning incentives of auditors $\uparrow$ compliance with environmental norms?

- **Setting:** Pollution audit system in Gujarat, India

- **Data:** 473 firms; audits, backchecks, and pollution monitoring

- **Research Design:** RCT
  - Treatment: random auditor assignment with fixed payment

- **Results:** 1) widespread corruption in status quo, 2) more truth-telling in treatment, 3) firms $\downarrow$ pollution

<!-- ## Background -->

<!-- - Third party auditing used to monitor firm regulatory compliance -->
<!--   - Financial sector: credit ratings -->
<!--   - Auto: compliance with vehicle standards -->
<!--   - Food retail: health inspector -->
<!--   - **Environmental:** compliance with national/state laws -->

<!-- - Feature: auditor chosen by, paid by, and reports to audited firm -->

<!-- - **Incentive to shade/falsify reports!** -->
<!--   - Corrupt information provision and undermine regulation -->

## Context: Gujarat, India

- One of India's fastest growing states
  - Home to the two post polluted industrial clusters

- Pollution standards set nationally, enforcement is state-level
  - GPCB responsible for enforcing Air Act

- GPCB uses regulates 20,000 plants

- Main regulatory instrument: plant-level inspections + third-party audits

## Environmental Audits in Gujarat

- Polluting firms submit yearly environmental audit

- \textcolor{red}{Audit conducted by audit firm hired and paid by the plant}

- Auditors visit plant three times/year to measure pollution
  - Submit audit report to GPCB

- (On paper) checks and balances on auditors
  - Max 15 plants per year
  - Can audit a plant at most 3 years in a row
  - Auditors with inaccurate reports can be decertified

## Environmental Audits in Gujarat

- (On paper) checks and balances on plants
  - Failure to submit report punishable by closure/disconnection
  - Noncompliance audit report punishable by closure/disconnection
\vfill  
- Despite safeguards, system functions poorly
  - GPCB believes inaccurate reporting renders audits useless for enforcement
\vfill
- Evidence of auditor shopping: Plants pay $\approx$ INR 24,000 (\$300)
  - Significantly less than true costs of conducting audit

## Experimental Design: N=473 plants

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig5.png",
                        auto_pdf = TRUE
                        )

```

## 
\centering
\Huge Lecture 9

\large Environmental Policy Design in LMICs

## Housekeeping

- First draft due on Thursday
  - Peer review due one week later (2 pages)
\vfill
- Proposal presentation (Oct 15th and 17th)
  - Consider drafting it as you write your first draft
  - Rubric added in github
\vfill
- Replication (3 pages, due Oct 21st)
\vfill
- Minor syllabus updates
  - State capacity in-class presentation
\vfill
- Migration lecture readings will be posted this week

## Model: Set Up

- **Three Players**: plant, auditor, and  regulator.
\vfill
- Plants face the decision to:
  - **Abate pollution**: Incurs a cost, but avoids penalties.
  - **Falsify reports**: Pay auditor to underreport pollution and avoid abatement.
\vfill
- Auditors considers:
  - Payment from plant
  - Risk of disbarment by regulator if found cheating.
\vfill
- Regulator: reduce pollution by enforcing penalties and disbarring false auditors

## Plant's Decision and Cost Function

- The plant's goal is to minimize total costs:

$$S - 1\{Abate\} \cdot c - w - 1\{\text{Fine Imposed}\} \cdot f $$

- Where:
  - S = Value of production
  - c = Cost of abating pollution
  - w = Payment to auditors for the report
  - f = Fine imposed if non-compliance is detected

- **Abatement**: Reduces emissions from $p_{h}$ to $p_{l}$, at cost $c \sim G(\cdot)$

- **Non-compliance**: Plant chooses $w(\hat{p}|p)$ to minimize costs
  - where $\hat{p}$ is auditor report, and $p$ is true pollution

## Auditor’s Payoff and Reporting Incentives

- Auditor's payoff function is:

$$w - a + E[V(\hat{p}|p)]$$

- Where:
  - $w$ is payment for performing audit
  - a = Cost of performing the audit
  - $E[V(\hat{p}|p)]$ = continuation value of future business
  - Assume leniency increases EV: $V_{lh} = V(p_{l}|p_{h})>V_{hh} =V_{0}$
\vfill
- Auditors may have incentive to falsify reports to maintain business
  - Higher \( V_{lh} \), more attractive for auditors to underreport
  - $E[V(\hat{p}|p)]$ incorporates likelihood of disbarrment

## Regulator Problem

- Reviews reports $\hat{p}$ and may impose $f$ on plant or disbar auditor

- Let $q$ = probability regulator observes true $p$ (backchecks)

- Decide whether to impose $f$

- Continuation value of disbarred auditor is zero

- Regulator only goal: minimize pollution

## Equilibrium: Proposition 1

Unique subgame-perfect equilibrium with regulation:

-  **Plants Abate**: If \( c \leq \min\{ f, qV_{lh} - \Delta V \} \).
-  **Underreporting**: If \( qV_{lh} - \Delta V < \min\{ c, f \} \).
-  **Defy Regulation**: If \( f < \min\{ c, qV_{lh} - \Delta V \} \).

\vfill

**Intuition**

- **Abatement**: when cheaper than paying fines or paying auditors to cheat

- **Underreport**: when paying auditors to cheat cheaper than abatement and fines

- **Defiance** when both abatement and bribing more expensive than risking fines

## Equilibrium: Proposition 2 - Backchecks

**Status Quo (control group)**

- $q \rightarrow 0$ in status quo
- Negative cost of false reports: $qV_{lh} - \Delta V$
- Result: Widespread underreporting and no incentive for abatement

\vfill

**With Backchecks and Incentive Pay (Proposition 2)**

- Backchecks: $\uparrow$ q s.t. $qV_{lh} - \Delta V > f$ induces fraction $G(f)$ of plants to abate
- Incentive pay: Makes underreporting more costly than fine if $q(V_{lh}+B)-\Delta V > f$

\vfill

**Intuition**:

- Backchecks increase $q$, raising expected cost of false reporting
- When $qV_{lh} - \Delta V > f$, some plants prefer abatement over risking detection.
<!-- - Higher $q$ makes underreporting less attractive, shifting plants towards compliance -->

## Random Assignment of Auditors

- **Flat fee:** Payment unconditional on report outcomes: $w({\hat{p}|p})=w(p)$
  - Plant cannot bribe auditor, nor can they refuse to pay

- Auditors only consider reputation $\rightarrow$ increase in $q$ for plants to abate is smaller

- Reduces continuation value of notoreity, $\Delta V$

\vfill

**Proposition 3**: Rise in $q$ s.t. $qV_{lh}\geq \Delta V$ induces share $G(f)$ to abate

\[
\text{If } qV_{lh} \geq \Delta V \implies \text{Plants will abate if } c \leq f
\]
\vfill

- **Intuition**: If treatments are strong enough s.t. $q(V_{lh}+B) > \Delta V$, then $G(f)$ plants will abate and rest will report higher $p$ and risk fine.

<!-- - Incentive Pay increases \( V_{lh} \) for accurate reporting s.t. false reporting less attractive -->

## Model Summary

- **Higher \( q \)** through backchecks increases risk of detection for false reporting

- **Random assignment** and **centralized payment** $\downarrow$ auditors-plant dependency

- **Incentive Pay** boosts auditors' continuation value for truthful reporting

- Treatment makes false reporting costlier for auditors and plants

- Plants either abate pollution or risk fines, depending on abatement costs

- The equilibrium shifts towards increased compliance and truthful reporting

## Data

- **Audit Reports (2009 & 2010)**: water and air pollution
   - Collected from auditors in both treatment and control groups
\vfill
- **Backchecks**: independent *unannounced* verification of pollution levels
   - Conducted randomly on 20% of treatment plants in 2009 and 2010
   - Midline backchecks in 2010 for both treatment and control groups.
\vfill
- **Endline Survey (2011)**:
   - Measured actual pollution outcomes after the audit intervention
\vfill
- **Administrative Data from GPCB**:
   - Regulatory actions, violations, and enforcement measures
   - Fines, closures, and disconnections imposed on plants

## Results: Status Quo Pollution
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig3.png",
                        auto_pdf = TRUE
                        )

```

## In Regression Form
$$ 1\{Compliant\}_{ij} = \beta_{1}1\{AuditReport\} + \alpha_{r} + \epsilon_{ij}$$
\pause
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Treatment Effects: Truth Telling About Plant Compliance

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig1.png",
                        auto_pdf = TRUE
                        )

```

## In Regression Form
\vspace{-0.5cm}
$$ 1\{Compliant\}_{ij} = \beta_{1}1\{AuditReport\} \times T_{j} + \beta_{2}1\{AuditReport\} + \beta_{3}T_{j} + \alpha_{r} + \epsilon_{ij}$$
\pause
\vspace{-0.5cm}
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "65%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Treatment increases truth-telling: 19pp less cheating in treatment

## Which firms are getting told on?

- Convert pollution readings to deviation from regulatory standard
\vfill
- Readings grouped into 40 different 0.05$\sigma$ width bins
\vfill
- 40 regressions: regress dummy for each bin on treatment indicator
  - Negative coefficient: treatment auditors less likely to report values in that bin
\vfill 
- Result: treatment reduces the mass just beneath the standard
  - **More likely** to report values above standard

## Density of audit report pollution distribution
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "55%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig6.png",
                        auto_pdf = TRUE
                        )

```
- Readings by treatment auditors 20\% less likely to be in bin just below standard compared to control readings

## Do plants reduce emissions?
- From model: treatment induces $G(f)$ plants abate by $\Delta$ incentives
- Regress plant pollution on treatment + region FE in endline (cross section)
\pause
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig8.png",
                        auto_pdf = TRUE
                        )

```
- Treatment plants $\downarrow$ pollution by 0.21$\sigma$

## Which plants are reducing pollution?
- Quantile regression: $Q_{y_{ij}|X_{j}}(\tau) = \beta T_{j} + \alpha_{r} + \epsilon_{ij}$
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "45%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig9.png",
                        auto_pdf = TRUE
                        )

```
- Clean-up driven by dirty plants (>80th pctile)
- Driven by monotonic increase in severity of GPCB fines w/ non-compliance

## Policy 
- Eliminating conflict of interest led to policy change
- Scaled up to state level since 2015

```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/duflo2013_fig10.png",
                        auto_pdf = TRUE
                        )

```

## Thoughts?

##

\Large Other challenges for designing low MAC policies?

## Barriers to Optimal Policy Design

- Recall, goal of social planner is to internalize externalities
\vfill
- Objective is to set $MWTP_{e} = \frac{\partial c_{e}}{\partial \Delta e}$ using **regulation**
\vfill
- We saw that low state capacity creates a wedge by keeping $\frac{\partial c_{e}}{\partial \Delta e}$ high
  - This is true even if market failures absent
\vfill
- **Top down:** $\uparrow$ state capacity **directly** has large benefits (more next week)
\vfill
- **Bottom up:** what if we instead change citizens' trust in the state?

## Jack et al. (2013): Money not to Burn

- **Research Question:** Can PES contracts curb crop burning?
\vfill
- **Setting:** North India; weak institutions for contract efficacy
\vfill
- **Data:** 171 villages in North India
\vfill
- **Design:** RCT with two treatments: 1) standard PES 2) partial payment upfront
\vfill
- **Results:** Upfront PES has 10pp higher compliance than standard PES
  - Takeaway: trust is an important barrier for efficient environmental policy design

## Textbook externality problem
```{r, echo = FALSE, message = FALSE, warning = FALSE, out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig5.png",
                        auto_pdf = TRUE
                        )

```
## Background

- Crop  burning major source of pollution in Punjab
\vfill
- Burning rice stalks post-harvest $\rightarrow$ 40% of winter pollution in Delhi
  - Severe health risks
\vfill
- Ban on burning since 2015, with fines for violations, but weak enforcement
  - Fine: Rs. 2,500 to Rs. 15,000 based on landholding
\vfill
- Subsidies for residue management equipment exist
  - insufficient to cover farmers' costs (still cheaper to burn)

## PES Contracts

- PES contracts pay farmers for environmentally friendly practices
\vfill
- Goal: increase the private costs of burning to incentivize ``good'' behavior
\vfill
- Advantage: targets outcome not inputs
  - i.e. agnostic about technology, as long as abatement cost $<$ payment
\vfill
- Challenge: trust and liquidity constrain standard PES contracts in LMICs
  - Treatment: pay 25\% or 50\% upfront

## RCT Design
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "60%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Data

- Baseline survey: 1,668 farmers (N=171 villages) from two districts in Punjab
\vfill
- Data: Demographics, income, trust in institutions, and CRM equipent
  - Farm plot shapefiles for linking with satellite imagery
\vfill
- Monitoring: field visits, random spot checks, satellite imagery

## Estimation

$$ y_{ij} = \alpha + \beta StandardPES_{j} + \gamma UpfrontPES_{j} + \psi X_{j} + \epsilon_{ij}$$

- $y_{ij}$ = outcome for farmer $i$ in village $j$
- $X_{ij}$ are strata fixed effects
- Subtreatments are pooled
- $\beta$: effect of standard PES relative to control
- $\gamma$: effect of upfront PES relative to control

## Results: Contract Compliance
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig2.png",
                        auto_pdf = TRUE
                        )

```
- Standard PES: 8.5\% complied with contract
- Upfront PES: 18\% complied with control --- trust and liquidity important!

## Other Results

- Relative to control, upfront PES $\uparrow$ non-burning by 7pp (85\%)
\vfill
- Most farmers still burned...
  - Cost of alternatives to burning still higher than PES payment
\vfill
- Standard PES insignificant in column 2 (inframarginal)
  - farmers who complied would not have burned anyway

## Mechanisms: Why do upfront PES increase compliance?

- Hypothesis: 1) distrust in conditional payment, 2) limited cash to rent CRM

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Upfront PES no better for farmers w/ liquidity constraints or distrust

## Mechanisms: Why do upfront PES increase compliance?

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/jack2023_fig4.png",
                        auto_pdf = TRUE
                        )

```
- Upfront farmers 7pp more likely to trust that payment will be made
- CRM cash shortages not important for either upfront or standard farmers
<!-- - Upfront amount may have been too low to ease constraint -->

## Discussion

- Limited short-term capital and distrust in institutions can limit effectiveness of standard PES contracts

- May generalize to conditional CCTs more broadly

- Thoughts?

##
\huge Other barriers to efficient policy design

\large Unintended Consequences: Spillovers

## Englander (2023): Information \& Spillovers in Conservation Policy

- **Research Question:** What are the direct and spillover effects of targeted conservation policy?
\vfill
- **Setting:** Temporary closures in Peru's anchoveta fishery
\vfill
- **Data:** location, time, and \# of fish caught per boat
\vfill
- **Research Design:** ``Potential'' closure algorithm
\vfill
- **Results:** direct effect: lower catch in closed areas
  - indirect effect: big positive spatial spillovers
  - net effect: policy backfires and $\uparrow$ catch!

## Targeted policies are common forms of incomplete regulation

- Reduce externality by targeting highest MD places
  - Road congestion pricing
\vfill
- Incomplete regulation $\rightarrow$ spillovers (impacts on non-targeted units)
\vfill
- Direction and magnitude of spillovers varies

## Peru's Anchoveta Fishery

:::::::::::::: {.columns}
::: {.column width="45%"}

\vspace{1cm}
- World's largest (tons caught)
\vspace{0.3cm}
- 8\% of global marine catch
\vspace{0.3cm}
- \$2 billion export revenue per year
\vspace{0.5cm}
- 730 active vessels per season

:::
::: {.column width="50%"}
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig1.png",
                        auto_pdf = TRUE
                        )

```
:::
::::::::::::::

## Catching juveniles is higher marginal damage than adults

```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "40%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Spatial Closures

- Regulator temporarily closes area of ocean when \% of juveniles $>$ 10\%

- Regulator exercises discretion; sometimes no closure

- Vessels report catch with electronic logbooks
  - Time, tons, and \% juvenile from each set
  - **Challenge:** fishermen underreport \% juvenile
  - **Solution:** third-party inspection data

- **Closure rule:** observe \% juvenile
  - Draw rectangle around overfished sets
  - Close area inside box
  - Announce on website, email, whatsapp

## Reported Size Distribution (example set)
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig3.png",
                        auto_pdf = TRUE
                        )

```

## Corrected Size Distribution (example set)
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Counterfactual Closures

- Mimic the closure rule (potential closures)

- Intersect potential closures (N=973) w/ actual declared closures (N=410)

- **Control for length distribution from which \% juveniles drawn**
  - Comparable distribution of fish sizes

- Similar potential closures, regulator believes different

## Treatment and Control Groups
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig5.png",
                        auto_pdf = TRUE
                        )

```

## Identifying variation

- Comes from controlling for length distribution
\vfill
- Compare potential closures that by chance had higher \% juvenile draws (and were thus closed by regulator) to potential closures that by chance had lower \% juvenile draws (and were thus not closed by regulator)
\vfill
- Key to identification: fishers report \% juvenile catch from 200 anchoveta out of several million in a set

## Empirical Strategy: Treatment Window for Spillovers
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig9.png",
                        auto_pdf = TRUE
                        )

```

## Empirical Strategy

$$ JuvenileCatch_{ist} = \alpha_{st} + \beta_{st}TreatFraction_{ist} + \sum_{l} \eta_{l}Prop_{il} + \Gamma X_{i} + \sigma_{wg} + \delta_{d} + \epsilon_{ist}$$

- i = potential closure
- s = spatial unit
- l = half-cm length interval
- w = two weeks 
- g = grid cell
- d = day

## Event Study
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig6.png",
                        auto_pdf = TRUE
                        )

```

## Main Results
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig7.png",
                        auto_pdf = TRUE
                        )

```

## Mechanism: Information Signal

- Game theoretic model
- Vessel chooses where to fish to max. expected profits
- Policy increases catch if:
  - Closures are +'ve signal relative to congestion costs
  - \% juvenile sufficiently high near closures
\vfill

- **Predictions:** 
  - Ex-post value of information competed away in equilibrium
  - Expected profit would be higher near closures if not for signal
  - Vessels receiving bigger info shocks have larger treatment effects

## Results: Value of positive signal is competed away

$$Tons_{vjk} = \beta_{1}Near_{vjk} + \beta_{2}DistToShore_{vjk} + \delta_{vj} + \gamma_{d} + \alpha_{jg} + \epsilon_{vjk}$$

where v=vessel, j=season, k=set, d=day, g=gridcell
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/gabe2023_fig8.png",
                        auto_pdf = TRUE
                        )

```

## Thoughts?

- Targeted policy backfires due to positive spillovers
- Policy implicitly signals info about non-targeted units

### 2024-11-11 lectures/module7/module7.Rmd
added .Rmd files

---
title: "| Environmental and Development Economics\n| Module 7 - Climate and Migration
  \n"
author: |
  | Raahil Madhok
  | UMN Applied Economics
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    theme: Montpellier
  slidy_presentation: default
  ioslides_presentation: default
classoption: aspectratio=169
header-includes:
- "\\AtBeginDocument{\\title[Module 7 - Climate and Migration]{Environmental and Development
  Economics \\\\ Module 7 - Climate and Migration}}"
- \input{header.tex}
- \usepackage{gensymb}
---

## 
\centering
\Huge Lecture 12

\Large Climate and Migration

## Housekeeping

- APEC Seminar Tomorrow: Stephanie Onder
\vfill
- Peer review due at midnight
  - I will return tomorrow
\vfill
- Proposal presentations start next week
  - **Oct 15:** 1) Matt B 2) Jovin L 3) Ryan M
  - **Oct 17:** 1) Gustavo M 2) Giang T 3) Sarah W
\vfill
- Office hours will continue until proposal deadline

## Replication Assignment (Due Oct 21st)

- Choose any envirodevonomics paper with data/code files
  - Must be published in either 1) general econ or 2) enviro/dev field journal

- Replicate at least 2 main tables/figures
  - Output everything to latex

- Write your own code and comment it!

- Extension
  - Must be "carefully thought out" and yield new insight
  - e.g. heterogeneity, non-linearity, robustness check

- Write-up (6 pages doubled spaced) 
  - summary, replication process, results, extension
  - Compare your findings with original paper. Interpret.

##
\centering
\huge Climate and Migration

## Environmental disasters are unequal

- Cross-country heterogeneity
  - Inequality in exposure (rich vs. poor countries)
\vfill
- Within-country heterogeneity
  - Coastal regions hit harder by floods
  - Agricultural regions hit harder by droughts
  - Biodiverse regions hit harder by rainfall shocks
\vfill
- Sectoral heterogeneity
  - Weather-dependent agriculture vs. services
\vfill

**Implication $\longrightarrow$ migration!**

## Unequal Climate Impacts

Historic average temperature (1986-2005)
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "95%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/climate_fig1.png",
                        auto_pdf = TRUE
                        )

```

## How do we study flows of people in response to climate change?

- Impacted region (push factors)
  - Direct effect: environmental extremes push people out
  - Indirect effect: violence, famine, etc. push people out
\vfill
- Receiving region (pull factors)
  - Direct effect: less exposure to environmental extremes draws migrants (wage impacts)
  - Indirect effect: e.g. housing prices
\vfill
- Most research: environment $\leftrightarrow$ migration in partial equilibrium
  - Elasticity of migration w.r.t. environmental quality
  - Elasticity of wages w.r.t environmental quality
\vfill
- **New frontier: general equilibrium!**

## Existing Evidence 
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/litreview.png",
                        auto_pdf = TRUE
                        )

```

## Spatial Models of Migration

- Economists are recently thinking about spatial equilibrium
\vfill
- Migrants choose where to live based on costs and benefits
  - Not only wage considerations
  - Amenity values, **exposure to environmental extremes**
\vfill
- When there is an environmental shock to the system...
  - Who moves?
  - What adjustments take place? Housing, land, wages, etc.
  - What is the new spatial equilibrium?
  - What happens to inequality?
\vfill
- This sets the framework for studying environment and migration

## Model Set Up

- Individual $i$ of wage group $j$ choose location $k$ to maximize utility:

$$ \max_{k} \quad \underbrace{v_{jk}}_\text{observed} + \underbrace{\epsilon_{ijk}}_\text{unobserved}$$

- The econometrician observes representative utility:

$$ v_{jk} = \alpha_{j}p_{k} + \beta e_{k} + x_{k}\gamma + \delta_{jk}$$

- housing price $p_{k}$
- environmental "bad" $e_{k}$ (flood risk, extreme temperature)
- observed amenities $x_{k}$, unobserved amenities $\delta_{jk}$
- price elasticity can differ by wage group, $\alpha_{j}$

## Model Set Up

- Logit probability ($\epsilon_{ijk} \sim EV$) yields migration probability (see module 4 slides):

$$ \pi_{ijk} = \pi_{jk} = \frac{e^{v_{jk}}}{\sum_{l}e^{v_{jl}}}$$

- since individuals w/n wage group $j$ have common choice probabilities

- In equilibrium, prices $p = p_{k}$ clear housing markets

$$ n_{k}^{D}(p^{*}) = n_{k}^{S}(p^{*}) \quad \forall k \quad \text{where} \quad n_{k}^{D} = \sum_{i} \pi_{ijk}, \quad n_{k}^{S} = \bar{n}_{k} $$

\vfill
- **Migration arises from wage-specific price elasticities and endogenous prices**
  - Environmental shocks $\uparrow$ inequality. Why?

## Estimation

- Goal: recover $\alpha_{j}$, $\beta$, $\gamma$
  - Tells us equilibrium prices and choice probabilities for any pattern of $e_{k}$
\vfill
- Inverting migration probabilty yields:
\begin{align*}
ln \pi_{jk} - ln \pi_{jo} & = v_{jk} - v_{j0} \\
  & = \alpha_{j} \Delta p_{k} + \beta \Delta e_{k} + \Delta x_{k} \gamma + \Delta \delta_{jk}
\end{align*}
\vfill

- **Challenge:** prices are endogenous i.e. correlated with unobserved amenities!
  - high-amenity places attract skilled migrants who bid up prices
  - Solution: cost shifting instrument

## Hisao (2024): Sea Level Rise and Urban Inequality in Indonesia

- Flood exposure: grid cells $<$ sea level for 1,3, 5m sea level rise (hydrological model)
\vfill
- Price instrument: ruggedness (housing cost supply shifter)
  - Exclusion restriction: terrain shifts housing supply but not demand
\vfill
- Define object of interest for simulation: flood exposure for wage group $j$: 
$$F_{j} = \sum_{k} f_{k}\pi_{jk}$$
- And price incidence by wage group $j$: 
$$P_{j} = \sum_{k}p_{k}\pi_{jk}$$

## Impact of flooding on inequality
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/hsiang2024_fig1.png",
                        auto_pdf = TRUE
                        )

```

- Current flooding: poor people's flood exposure is 1.4x that of rich people
- 5m sea level rise: inequality of exposure doubles
- Prices: richer people move, lower prices compensate for higher flood exposure

## Is inequality driven by prices or migration?
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/hsiang2024_fig1.png",
                        auto_pdf = TRUE
                        )

```
- Evaluate $F_{j}$ with projected flooding and current $\pi_{jk}$ (no migration)
- Evaluation $P_{j}$ with counterfactual prices and current $\pi_{jk}$ (no migration)
- Reveals direct impact of floods on rich/poor **without migration**
- No $\Delta$ in inequality in this scenario i.e. **migration drives unequal exposure!**

## Reduced Form Literature

- **Impact of Migration on Environment**
  - \textcolor{red}{Brewer et al. (2022, Uganda)}; Hornbeck and Naidu (2014, USA); Rozelle et al. (1991, China); Salemi (2021, Africa)}
\vfill
- **Impact of Environment on Migration**
  - Morten (2019, India); \textcolor{red}{Mueller et al. (many papers)}; Missrian and Schlenker (2017, dev countries); Chen et al. (2022, China)
\vfill
- **Migration as Adaptation to Environmental Extremes**
  - \textcolor{red}{Cecato et al. (2024, Indonesia)}; Blakeslee et al. (2020, India)
\vfill
- **Migration and Environment in General Equilibrium**
  - \textcolor{red}{Madhok et al. (2024, India)}; Asher et al. (2024, India); Blakeslee et al. (2022, India)

## Impact of Migration on the Environment: Brewer et al. (2022)

- **Research Question:** What is the impact of rural-urban migration on land use?
  - What are the intensive vs. extensive margin impacts?
\vfill
- **Setting:** Uganda from 2005-2019 
  - 70\% of land is agricultural
\vfill
- **Data:** Many rounds of the LSMS survey (3200 households)
\vfill
- **Research Design:** SSIV: urban consumption (shift) x migrant network (share)
\vfill
- **Results:** Migration causes agricultural contraction and specialization
  - Leads to a reduction in tree cover loss

## Conceptual Framework
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "70%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Estimation

- 1st stage: $IHS(M_{ijt}) = \underbrace{\beta_{11}Z_{jt}M_{ij0}}_\text{shift-share IV} + \beta_{21}Z_{jt} + \beta_{31}W_{jst}+\alpha_{1i} + \psi_{1w} + \gamma_{1t} + \epsilon_{1ijt}$
\vfill
- 2nd stage: $IHS(Y_{ijt}) = \beta_{12}(\hat{M}_{ijt}) + \beta_{22}Z_{jt} + \beta_{32}W_{jst}+\alpha_{2i} + \psi_{2w} + \gamma_{2t} + \epsilon_{2ijt}$
\vfill
- Outcome: $M_{ijt}$ is number of months that hh member $i$ was absent in year $t$
\vfill
- Shift-share IV: 
  - $Z_{jt} = \sum_{d=1}^{n}(\frac{C_{dt}}{dist_{jd}})$
  - $C_{dt}$ is mean consumption in district d
  - $Z_{jt}M_{ij0}$ is spatially weighted consumption shocks

## Results: Agricultural extent
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig2.png",
                        auto_pdf = TRUE
                        )

```
- rural-urban migration leads to agricultural contraction (extensive margin)
- rural-urban migration has no impact on area planted for annual vs. perennial crops

## Results: Crop Specialization
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "90%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig3.png",
                        auto_pdf = TRUE
                        )

```
- migration causes households to specialize
- 1) reduce number of different crops, 2) invest proportionally more in fewer crops

## Results: higher district level tree cover
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/brewer2022_fig4.png",
                        auto_pdf = TRUE
                        )

```

## Impact of Environment on Migration: Chen and Mueller (2018)

- **Research Question:** How do farmers adapt to sea-level rise? 
  - At what point do farmers migrate?

\vfill
- **Setting:** Coastal Bangladesh, 2003-2011
\vfill
- **Data:** Several sources
  - Migration: Vital registration system
  - Agriculture: Household Income and Expenditure Surveys
  - Floods: NASA MODIS water index
\vfill
- **Research Design:** Fixed effects regression
\vfill
- **Results:** Indundation *does not* trigger migration
  - Migration is driven by adverse effects of soil salinity on agriculture

## Conceptual Framework

- Recent studies show minimal migratory response to flooding
  - Monsoons lead to temporary displacement, but people return
  - Delayed yield benefits
  - Many farmers have adapted ag practices to accommodate inundation
\vfill
- Inundation may not be the right ``push'' factor
  - What exactly is triggering flood-related migration?
\vfill
- Hypothesis: gradual increases in salinity (not flooding) will bear modest effects on
migration until less costly adaptation strategies in the agricultural
sector have been exhausted

## Empirical Strategy
- Impact of salinity and flooding on agricultural incomes:

$$ Y_{hjt} = \theta_{F} F_{jt-1} + \theta_{S}S_{jt-1} + \beta Z_{hjt} + \delta X_{hjt-1} + \beta_{S}^{88}S_{j,1988} + \tau_{t} + \mu_{hjt} $$

- $Y_{hjt}$ = annual crop revenue for household $h$ in subdistrict $j$ at time $t$
- $F_{jt-1}$ = lagged flooding; $S_{jt-1}$ = lagged salinity
- $Z_{hjt}$ = demographic and wealth covariates
- $X_{hjt-1}$ = lagged environmental exposure (rain, temp, etc.)
\vfill
- Impact of salinity and flooding on migration

$$M_{hjt} = \beta_{F} F_{jt-1} + \beta_{S}S_{jt-1} + \alpha Z_{hjt} + \gamma X_{hjt-1} + \beta_{S}^{88}S_{j,1988} + \rho_{t} + \epsilon_{hjt}$$

## Results: Effect of flooding/salinity on farm revenue
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/mueller2018_fig1.png",
                        auto_pdf = TRUE
                        )

```

- 1$\sigma$ increase in salinity reduces crop revenue by 0.1$\sigma$ (2500 taka)
- No impact on total farm revenue
- hh shift to aquaculture to mediate income risk as salinity compromises agriculture

## Results: Effect of flooding/salinity on migration
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/mueller2018_fig2.png",
                        auto_pdf = TRUE
                        )

```
- 1$\sigma$ increase in soil salinity increases internal migration by 0.43pp
- 1$\sigma$ increase in soil salinity decreases international migration by 0.19pp
- International migration and aquaculture may be substitutes

## Migration as Adaptation: Cecato et al. (2024)

- **Research Question:** Can migration mitigate weather damages?

\vfill
- **Setting:** Indonesia over 20 years
\vfill
- **Data:** 
  - Migration: 32,000 individuals from IFLS survey
  - Weather: Gridded rain/temperature 
\vfill
- **Research Design:** Shift-share IV
  - Shift: rainfall shock
  - Share: ethnic networks
\vfill
- **Results:** 1$\sigma$ $\downarrow$ in precipitation at origin $\downarrow$ non-migrant consumption 
  - Ability to migrate leads to 14\% increase in consumption over medium run

## Three facts

- There is high migration in Indonesia
  - 4.3\% percent of population per year migrates internally
\vfill
- Rainfall is an important determinant of migration in Indonesia
  - Kleemans and Magruder (2018)
\vfill
- Mobility is associated with ethnicity 
  - Propensity to migrate persists over time within groups (Auwalin 2020)

## Estimation

- **First stage regressions:**

$$M_{ikt} = a_{0}M_{ikt} + a_{1}(Networks_{ik} \cdot R_{kt}) + a_{2}R_{kt} + \phi_{i} + \phi_{t} + X_{ikt} + \epsilon_{ikt}$$

$$M_{ikt} \times R_{kt} = b_{0}M_{ikt} + b_{1}(Networks_{ik} \cdot R_{kt}) + b_{2}R_{kt} + \phi_{i} + \phi_{t} + X_{ikt} + \epsilon_{ikt}$$

\vfill

- **Second stage:**
$$Y_{ikt} + \beta_{0}\hat{M}_{ikt} + \beta_{1}(\widehat{M_{ikt} \times R_{kt}}) + \alpha_{1}R_{kt} + \phi_{i} + \phi_{t} + X_{ikt} + \epsilon_{ikt}$$

\vfill

- $Y_{ikt}$: log consumption of hh $i$ in county $k$ in year $t$
- $R_{kt}$: rainfall shock (z-score)
- $Networks_{ik}$: share of ppl of same ethnicity in origin at age 12 who moved away

## Results: Migration mitigates weather damage
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/cecato2024_fig1.png",
                        auto_pdf = TRUE
                        )

```
- 1$\sigma$ decrease in rainfall reduces consumption by 1.82\%
- Ability to migrate more than offsets this

## Migration and Agriculture in General Equilibrium: Madhok et al. (2024)

- **Research Question:** How does agriculture unfold when labor exits?
\vfill
- **Challenge:** Two competing channels: labour (direct/PE) + prices (indirect/GE)
\vfill
- **Setting:** India --- very high labour misallocation
\vfill
- **Data:** IHDS-I and IHDS-II panel surveys (42,000 households)
\vfill
- **Research Design:** Shift-share IV
  - Shift: destination income shocks
  - Share: 1) inverse distance, 2) number of males at home
\vfill
- **Results:** migration $\downarrow$ agriculture near cities, but $\uparrow$ in remote areas
  - i.e. spatial reorganization of agriculture

## Partial Equilibrium

- **Shift share IV:** 
\begin{equation*}
	z_{idt} = \underbrace{\varphi_{i}}_\text{productivity (share I)} \times \underbrace{\sum_{d' \in \Theta/ d} \frac{1}{\tau_{dd'}}}_\text{inverse distance (share II)} \times \quad \underbrace{(inc_{d't} \cdot pop_{d'})}_\text{shift}
\end{equation*}

- **2SLS Equations:**

\begin{align*}
   & M^{labor}_{idt}  = \mu_{1}z_{idt} + \mu_{2}s_{dt} + \mu_{3} inc_{dt} + \Gamma X'_{dt} + \alpha_{i} + \gamma_{t} + \varepsilon_{ijdt} \\
    & Y_{idt}  = \beta_{1}\widehat{M^{labor}_{idt}}
+ \beta_{2}s_{dt} + \beta_{3}inc_{dt} + \Gamma X'_{dt} + \alpha_{i} + \gamma_{t} + \eta_{ijdt}
\end{align*}

## PE Results: No labor-capital substitution
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig1.png",
                        auto_pdf = TRUE
                        )

```

## Spatial Equilibrium: Visual Evidence
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "100%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig2.png",
                        auto_pdf = TRUE
                        )

```

## Migration and Market Adjustments
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "80%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig3.png",
                        auto_pdf = TRUE
                        )

```
- Mass migration from the crop market increases crop prices
- Mass migration from the land market lowers land prices

## Migration and the Spatial Reorganization of Agriculture
```{r, echo = FALSE, message = FALSE, warning = FALSE,  out.width = "75%", fig.align = "center"}

knitr::include_graphics(
                        path = "assets/madhok2024_fig4.png",
                        auto_pdf = TRUE
                        )

```
- PE (agricultural contraction) dominates in peri-urban areas
- GE (agricultural expansion) dominates in remote areas
- Result: food shortage induced by direct channel compensated by markets

##
\centering
\huge New Directions for Environment and Development Economics

\Large Open Research Areas

## Exciting Area for Research

- **Main takeaway from course:**
  - Many reasons for poor environmental quality in LMICs
  - Evidence on many topics is very thin!

\vfill
- Barriers to research are falling
  - Remote sensing (e.g. cashews in West Africa)
  - Administrative data (e.g. Annual Survey of Industries)
  - Poverty proxies (e.g. nightlights, other geospatial proxies)
\vfill

## Much Room for Methodological Innovations

- I have tried to emphasize **applied theory**
  - We can take existing models and adapt them to LMIC contexts
  - E.g. add credit constraints, factor market constraints
  - Many new insights we can take to the field!
\vfill
- Other methodological innovations
  - Demand estimation (e.g. Ito and Zhang, 2020; Kremer et al., 2013)
  - Field experiments (e.g. Saavedra, 2024; Duflo et al., 2013)
  - Lab-in-the-field (e.g. Dean and Berkouwer, 2022)
  - Reduced form quasi-experimental

## Innovation in Theory

1. Environmental and development econ study market failures
    - New theory can help explain interactions
    - e.g. Burgess et al. (2012), Jack et al. (2024), Madhok and Gulati (2024), Braganca and Dahis (2022)
\vfill   
2. IO and behavioural economics have clear intersections w/ envirodevonomics
    - WTP elicitation (revealed pref, stated pref, etc.)
    - e.g. Ito and Zhang (2020), Berkhouwer and Dean (2022), Kremer et al. (2013)
    - See reading list for many other WTP for environment papers
\vfill
3. [New] Macro-structural and trade

## Research Frontiers: $MWTP_{e}$

- What is $MWTP_{e}$ in poor countries?
  - Lots of work on pollution (air/water)
  - Less work on biodiversity/forests
\vfill
- What is the role of market failures in explaining low $MWTP_{e}$?
  - Credit markets, labor markets, property rights, information access
\vfill
- Is low state capacity reflected in low MWTP?

## Research Frontiers: Environmental policy design in LMICs

- What determines successful environmental policy in LMICs?
  - What is role of monitoring, enforcement, information asymmetries?
  - New areas: e.g. trust (Jack et al., 2024)
\vfill
- Does information provision to public change exposure to environmental risk?
\vfill
- How can environmental policy be tailored to accommodate informal sector?

## Research Frontiers: Political Economy

- How do political structures lead to rent-seeking and overexploitation?
  - Can conservation policy mitigate rent-seeking behavior?
\vfill
- How does leader identity influence environmental outcomes?
  - Ethnic alignment (e.g. Madhok and Gulati, 2024)
  - Gender (e.g. Jagnani and Mahadevan, 2024)
  - Minority/majority: e.g. (Gulzaar et al, 2023)
\vfill
- Do extractive institutions cast a "long shadow" on the environment?

##
\centering
\huge Thank you!

\large Next week: Proposal Presentations
